### 2015-08-06 Manuscript.tex
Updates from ShareLaTeX

\documentclass{sig-alternate}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{cite}
\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}[leftmargin=0.4cm]}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\S\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
\usepackage{picture}
\usepackage{colortbl}
\usepackage[table]{xcolor}
\usepackage{listings}

\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\usepackage{tabu}

\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

\lstset{
    language=Python,
    basicstyle=\ttfamily\fontsize{2.4mm}{0.8em}\selectfont,
    breaklines=true,
    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=tlrb,
    showtabs=false,
    showspaces=false,
    showstringspaces=false,
    %backgroundcolor=\color{Gray},
    keywordstyle=\bfseries,
    emph={COCONUT,GUESSES,ASSESS,COCOMO2,PEEKING2,SAMPLE,WHERE,RIG}, emphstyle=\bfseries\color{Blue},
    stringstyle=\color{green!50!black},
    commentstyle=\color{red}\itshape,
    %numbers=none,
    captionpos=t,
    numberstyle=\bfseries\color{red},
    escapeinside={\%*}{*)}
}

\definecolor{darkgreen}{rgb}{0,0.3,0}

\usepackage[table]{xcolor}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}

\newcommand{\G}{\cellcolor{green}}
\newcommand{\Y}{\cellcolor{yellow}}

\newcommand{\quart}[4]{\begin{picture}(100,4)%1
{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\newenvironment{changed}{\par\color{MyDarkBlue}}{\par}
%\newenvironment{changed}{\par}{\par}

\newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
%\newcommand{\ADD}[1]{#1}
\usepackage{times}
\def\baselinestretch{1}
\setlist{nosep}
\usepackage[font={small}]{caption, subfig}

\setlength{\abovecaptionskip}{1ex}
\setlength{\belowcaptionskip}{1ex}
\setlength{\floatsep}{1ex}
\setlength{\textfloatsep}{1ex}
\newcommand{\subparagraph}{} % defined before loading titlesec
\usepackage[compact,small]{titlesec}
\DeclareMathSizes{7}{7}{7}{7} 
\bibliographystyle{unsrt}
\pagenumbering{arabic} 
\begin{document}  
\conferenceinfo{ISCE}{'16 Austin, Texas}

\title{Prediction is Hard (Especially About the Future)}
\numberofauthors{2} 
\author{
\alignauthor 
Rahul Krishna \\
       \affaddr{CS, NcState, USA}\\
       {i.m.ralk@gmail.com}
\alignauthor
Tim Menzies\\
       \affaddr{CS, NcState, USA}\\
       {tim.menzies@gmail.com}}
\setlength{\columnsep}{7mm}

\maketitle
\begin{abstract}
This paper provides ...  
\end{abstract}
\section{Introduction}
\section{Cluster and Contrast}
A recurring data analysis pattern in this paper is $cluster+contrast$. The data is distilled into a few clusters using a clustering scheme. Then lessons are inferred by studying the differences between these clusters. These lessons are used to generate rules that can be applied in any context.
\subsection{Clustering}
There are a wide variety of clustering methods to choose from. A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and defect data from the PROMISE repository~\cite{promise}. In that study methods such as K-Means, mini-batch-K-Means, DBScan, EM, Ward, and WHERE were investigated. The results of the study showed that the size and number of clusters is more important that the specifics of the techniques used.

For this purposes of this work, we have chosen WHERE, a clustering scheme which is capable of generating at least $\sqrt{N}$ clusters given $N$ instances. In addition to this, WHERE has been shown to run fast while ignoring spurious dimensions~\cite{menzies2013}. This is particularly useful, for much of the DE data is noisy, in that they contain a information not associated with the target variable. 
\subsection{Finding Contrasts}
All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability, shown as Step-1 in figure \ref{fig:xtrees_bare}. Several algorithms are explored as possible tools used to identify contrast between clusters these include Nearest Neighbors methods, Case Based Reasoning, and Decision Trees.

\subsubsection{Nearest Neighbors}
T
\subsubsection{Case Based Reasoning}

\subsubsection{Decision Trees}

\section{CROSSTREES}
\subsection{Design}
	\begin{figure}[t]
		\small
		~\hrule~

		{\bf Step1: Top down clustering using WHERE}

		The data is recursively divided in clusters using WHERE as follows:
		\begin{itemize}

			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(this requires only $O(2N)$ comparisons
			of $N$ cases).
			\item Project each case $Z$
			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
			rule. 
			\item Split the data at the median $X$ value of all cases and
			recurses on each half  (stopping when
			one half has less  than $\sqrt{N}$ of the original population).
		\end{itemize}
		~\hrule~

		{\bf Step2: Distinguish between clusters using  decision trees}

		Call each leaf from WHERE a  ``class''. Use an entropy-based
		 decision tree (DT) learner to learn what attributes select for each ``class''. To limit tree size:
		 \bi
		 \item Only use the top $\alpha=33$\%  of the features, as determined by their information gain~\cite{Irani1993}. 
		 \item Only build the trees down to  max depth of $\beta=10$.
		 \item Only build subtrees if it contains at least $N^{\gamma=0.5}$ examples (where $N$ is the size of the training set).
\ei
{\em Score}  DT  leaf nodes  via the mean score of its majority cluster.

		~\hrule~

		{\bf Step3: Generating contrast sets from DT branches}
		\begin{itemize}
		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
		\item Find the {\em desired} cluster: 
		\bi
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}. If none found, then repeat for $lvl += 1$
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
		\bi
		\item
		For discrete attributes,  return the value from {\em desired}. 
		\item
		For  numerics, return the numeric difference. 
		\item
		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
	\ei
	\ei
		~\hrule~
		\caption{CROSSTREES. Controlled by the parameters
		$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
		\label{fig:xtrees_bare}
	\end{figure}

\subsection{Assessment}
\section{Experiments}
\subsection{Data}
\subsection{When not to Plan?}
\subsection{Evaluation}
\subsection{Results}
\section{Threats to Validity}
\section{Conclusion}
\section*{Acknowledgements}
\bibliographystyle{plain}

\bibliography{References}
\balanced
\end{document}

\end{document}

### 2015-08-07 Manuscript.tex
Updates from ShareLaTeX

@@
 A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and defect data from the PROMISE repository~\cite{promise}.
-In that study methods such as K-Means, mini-batch-K-Means, DBScan, EM, Ward, and WHERE were investigated.
+In that study methods such as WHERE, K-Means, mini-batch-K-Means, DBScan, EM, and Ward were investigated.
 The results of the study showed that the size and number of clusters is more important that the specifics of the techniques used.
@@
 In addition to this, WHERE has been shown to run fast while ignoring spurious dimensions~\cite{menzies2013}.
-This is particularly useful, for much of the DE data is noisy, in that they contain a information not associated with the target variable.
+This is particularly useful, for much of the SE data is noisy, they contain a information not associated with the target variable.
+
 \subsection{Finding Contrasts}
 All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability, shown as Step-1 in figure \ref{fig:xtrees_bare}.
-Several algorithms are explored as possible tools used to identify contrast between clusters these include Nearest Neighbors methods, Case Based Reasoning, and Decision Trees.
+In this work we have explored several algorithms as possible tools used to identify contrast between clusters.
+These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees.
 
-\subsubsection{Nearest Neighbors}
-T
 \subsubsection{Case Based Reasoning}
+
+Case-based reasoning seeks to find solutions to problems by emulating human recollection and adaptation from past experiences.
+It has found extensive usage in Artificial Intelligence because it offers several advantages.
+However, one of the most important benefits that CBR has to offer is that it works on a ``case-by-case'' basis.
+Therefore it's advise is tailored to be specific to the particular case being considered.
+Several paper in SE have applied this technique, most of all for effort estimation ~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.
+
+A classic example of a CBR is a nearest neighbor approach.
+The nearest neighbor approach is rather straight forward and has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.
+
+Our approach is rather straight forward, in that we divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
+For every test case, we identify a cluster from the training set that most closely resembles it.
+Following this, we find the nearest cluster with a better performance score.
+The differences in the attributes between these two clusters constitute the \textit{"contrast set"}.
+These contrast sets acts as plans that can be used to reflect over the test cases to improve them.
+
+In most SE applications, not all features contribute equally to a problem.
+With this in mind, we opine that it would be beneficial if the above method is extended to include some form of feature weighting, thus enabling the tools to recommend changes to only the most informative features.
+
+\subsubsection{HOW}
+
+HOW is a very different planner to other CBR systems. HOW
 
@@
 \subsection{Data}
+
+\begin{figure*}[!t]
+\scriptsize
+   \begin{center}
+   \begin{minipage}{.46\linewidth}
+    \begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}
+   & \multicolumn{5}{c|}{ }\\
+
+   & \multicolumn{5}{c|}{ Data set  properties}\\ 
+   & \multicolumn{5}{c|}{  }\\ 
+           & \multicolumn{2}{c|}{training}   & \multicolumn{3}{c|}{testing}      \\ \cline{2-6}
+   data set      & versions           & cases & versions     & cases    & \% defective             \\ \hline
+        jedit    & 3.2, 4.0, 4.1, 4.2 & 1257      & 4.3          & 492          & 2 \\
+        ivy      & 1.1, 1.4           & 352       & 2.0          & 352          & 11 \\
+        camel    & 1.0, 1.2, 1.4      & 1819      & 1.6          & 965          & 19 \\
+        ant      & 1.3, 1.4, 1.5, 1.6 & 947       & 1.7          & 745          & 22 \\
+        synapse  & 1.0, 1.1           & 379       & 1.2          & 256          & 34 \\
+        velocity & 1.4, 1.5           & 410       & 1.6          & 229          & 34 \\
+        lucene   & 2.0, 2.2           & 442       & 2.4          & 340          & 59 \\
+        poi      & 1.5, 2, 2.5        & 936       & 3.0          & 442          & 64 \\
+        xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
+        log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
+        xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline
+
+    \end{tabular}\end{minipage}\begin{minipage}{.4\linewidth}
+    \begin{tabular}{|rrr|rrr|rr|l} \cline{1-8}
+      \multicolumn{8}{|c|}{  }\\
+      \multicolumn{8}{|c|}{  Results from learning}\\
+       \multicolumn{8}{|c|}{   }\\
+   \multicolumn{3}{|c|}{untuned} & \multicolumn{3}{c|}{tuned} & \multicolumn{2}{c|}{change}\\
+  \cline{1-8}
+
+  pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
+  55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
+  65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
+  49 & 31 &   & 56 & 37 &   & 5 & 6\\
+  49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
+  45 & 19 &   & 47 & 15 &   & 2 & -4\\
+  78 & 60 &   & 76 & 60 &   & -2 & 0\\
+  56 & 25 &   & 60 & 25 & y & 4 & 0\\
+  56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
+ 30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
+  32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
+  38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
+\hline 
+\end{tabular}
+
+\end{minipage}
+\end{center}
+
+    \caption{Training and test {\em data set properties} for the Jureczko data sets,
+    sorted in ascending order of \% defective examples.
+    On the right-hand-side, we show the {\em results from learning}.
+    Data is ``good'' if it has recall over 60\% and false alarm under 40\%
+(and note that, after tuning, there are more ``good'' than before).
+Data   marked with ``$\star$'' show large improvements in performance, after tuning.
+Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
+}\label{fig:j}
+\end{figure*}
 \subsection{When not to Plan?}

### 2015-08-07 Manuscript.tex
Updates from ShareLaTeX

@@
 All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability, shown as Step-1 in figure \ref{fig:xtrees_bare}.
-In this work we have explored several algorithms as possible tools used to identify contrast between clusters.
+
+In this work we have explored several algorithms as possible tools to identify contrast between clusters.
 These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees.
+Each technique offers a distinct advantage over the other.
 
@@
 
-Our approach is rather straight forward, in that we divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
+Our approach is rather straight forward, we divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
 For every test case, we identify a cluster from the training set that most closely resembles it.
@@
 
-HOW is a very different planner to other CBR systems. HOW
+HOW is very different compared to other CBR planners in that it explores the gradient between pairs of nearby clusters instead of studying the clusters themselves.
+HOW works by clustering the data during training using WHERE and then drawing \texttt{slopes} between the centroids of pairs of nearby clusters.
+Assuming the cluster pairs are labeled X and Y, with X having slightly better performance score than Y, the \texttt{slope} between X and Y acts as an indicator pointing to a direction to displace the data; i.e.
+away from Y and towards X.
+
+While testing, HOW finds the nearest slope to every test case.
+The slope provides the exact magnitude and direction of displacements.
+Contrast sets are derived from these displacements.
+HOW offers a distinct advantage over CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).
 

### 2015-08-07 Manuscript.tex
I added shitloads of stuff, I think.. Maybe not. Whatever!

@@
 \newcommand{\PEEKING}{{\bf PEEKING2}}
+%\usepackage[table]{xcolor}
+\definecolor{darkgreen}{rgb}{0,0.3,0}
+\definecolor{Gray}{rgb}{0.88,1,1}
+\definecolor{Gray}{gray}{0.85}
+\definecolor{Blue}{RGB}{0,29,193}
+\usepackage{colortbl}
 \usepackage{picture}
-\usepackage{colortbl}
-\usepackage[table]{xcolor}
+
 \usepackage{listings}
+\usepackage{tabu}
 
@@
 \definecolor{darkgray}{gray}{0.6}
-\usepackage{tabu}
-
 \definecolor{Gray}{gray}{0.95}
@@
 }
-
-\definecolor{darkgreen}{rgb}{0,0.3,0}
-
-\usepackage[table]{xcolor}
-\definecolor{Gray}{rgb}{0.88,1,1}
-\definecolor{Gray}{gray}{0.85}
-\definecolor{Blue}{RGB}{0,29,193}
 
@@
 
+Clustering is followed generating \textit{contrast sets}.
+These contrast sets represent recommendations on what could be altered to better improve an outcome.
 In this work we have explored several algorithms as possible tools to identify contrast between clusters.
 These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees.
-Each technique offers a distinct advantage over the other.
+These techniques are discussed below.
+It is worth noting that they are organized in such a way that each technique seeks to address certain fallacies in the ones that precede it.
 
@@
 With this in mind, we opine that it would be beneficial if the above method is extended to include some form of feature weighting, thus enabling the tools to recommend changes to only the most informative features.
+
+\texttt{<Demrits of knn here and lead to the next\\ section>}
 
@@
 \bibliography{References}
-\balanced
 \end{document}

### 2015-08-07 Manuscript.tex
Updates from @rahlk's local machine.

@@
 \usepackage{tabu}
-
+\DeclareMathOperator*{\argmin}{arg\,min} 
+\DeclareMathOperator*{\argmax}{arg\,max}
 \definecolor{lightgray}{gray}{0.8}
@@
 These techniques are discussed below.
-It is worth noting that they are organized in such a way that each technique seeks to address certain fallacies in the ones that precede it.
+It is worth noting that the following techniques are organized in such a way that each one seeks to address certain fallacies of the ones that precede it.
 
@@
 
-A classic example of a CBR is a nearest neighbor approach.
-The nearest neighbor approach is rather straight forward and has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.
+A classic example of a CBR is K Nearest Neighbor.
+Our version of nearest neighbor for planning is rather straight forward.
+It has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.
 
-Our approach is rather straight forward, we divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
+We divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
 For every test case, we identify a cluster from the training set that most closely resembles it.
@@
 HOW offers a distinct advantage over CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).
+
+\begin{figure}[t!]
+	\small
+	~\hrule~
+
+	{\bf \fig{where}a: top down clustering:}
+
+	Data can be rapidly and recursively divided   as follows.
+	Find   two   distance cases,  $X,Y$
+	by picking any case $W$ at random, then setting $X$ to its most
+	distant case, then setting $Y$ to the case most distant from
+	$X$~\cite{fastmap}
+	(this requires only $O(2N)$ comparisons
+	of $N$ cases).
+	Next, HOW projects each case $Z$
+	onto a {\tt Slope} that  runs between $X,Y$ using the cosine
+	rule of \fig{where}b. After that,  split the data at the median $x$ value of all cases and
+	recurses on each half  (stopping when
+	one half has less  than $\sqrt{N}$ of the original population).
+
+	~\hrule~
+
+	{\bf \fig{where}b: finding distances:}
+
+	In the \fig{where}a, to compute distances, we use
+	the Euclidean measure recommended for
+	case-based reasoning by Aha et al.~\cite{aha91};
+	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
+	where values are  normalized 0..1 for the range min..max and 
+	$w_i$ is some importance weight given to attribute $i$.
+	Usually, $w_i=1$ but it can be adjusted using, say,
+	the feature weighting methods of \fig{where}d.
+
+	~\hrule~
+
+	{\bf \fig{where}c: 3 point geometry:}
+
+	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
+	A third point, $Z$, has distances  $a=dist(Z,X)$ and
+	$b=dist(Z,Y)$ to $X,Y$, respectively.
+	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
+	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
+	Further, according to Pythagoras' theorem, $Z$ stands at a distance
+	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$.
+
+	~\hrule~
+
+	{\bf \fig{where}d: feature weighting:}
+
+	HERE's feature weighting algorithm
+	comes from the CART regression tree learner~\cite{Breiman1984}.
+	It sorts independent variables
+	according to how well they can reduce the variability
+	of a  numeric objective.
+	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
+	then the expected
+	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
+	where $v$ is the standard deviation of the objective.
+	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
+	then split at $\argmin_i w_i$, then recursing into each half.
+
+	~\hrule~
+
+	\caption{Sub-routines within HOW.}\label{fig:where}
+\end{figure}
 

### 2015-08-07 Manuscript.tex
Updates from @rahlk's local machine.
Updates from ShareLaTeX

@@
 In this work we have explored several algorithms as possible tools to identify contrast between clusters.
-These fall into three broad categories: a) Case Based Reasoning techniques (Nearest Neighbors), b) A Gradient base planner (called HOW), and c) Decision Trees.
-These techniques are discussed below.
-It is worth noting that the following techniques are organized in such a way that each one seeks to address certain fallacies of the ones that precede it.
+These fall into two broad categories:
+\begin{enumerate}
+\item Case Based Reasoning techniques (Nearest Neighbors and a gradient base planner called HOW)
+\item Decision Trees.
+\end{enumerate}
+
+These techniques are discussed below. It is worth noting that the following techniques are organized in such a way that each one seeks to address certain shortcomings in the ones that precede it.
 
@@
 With this in mind, we opine that it would be beneficial if the above method is extended to include some form of feature weighting, thus enabling the tools to recommend changes to only the most informative features.
+This paper uses a feature weighting scheme similar to that used in CART, see \fig{where}d.
 
-\texttt{<Demrits of knn here and lead to the next\\ section>}
+Simple though this method may seem, using nearest neighbors presents a couple of significant challenges: 
+\begin{itemize}
+\item[1.] Displacements are seldom limited to small regions, this implies that large changes may be recommended when not necessary. This may make the solutions less feasible to implement. 
+\item[2.] The nearness of a test case to a cluster in the training data is subjective.
+This affects the locality of the solutions, thus jeopardizing the relevance of these changes to the test case under consideration.
+\end{itemize}
+
+In view of the above issues, we developed HOW, a tool that combines (a) Feature Selection; (b) Centroid generation from clusters; (c) Contrast generation using gradient between clusters.
 
@@
 
-HOW is very different compared to other CBR planners in that it explores the gradient between pairs of nearby clusters instead of studying the clusters themselves.
+HOW is very different compared to conventional CBR planners in that it explores the gradient between pairs of nearby clusters instead of studying the clusters themselves.
 HOW works by clustering the data during training using WHERE and then drawing \texttt{slopes} between the centroids of pairs of nearby clusters.
@@
 Contrast sets are derived from these displacements.
-HOW offers a distinct advantage over CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).
+HOW offers a distinct advantage over the other CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).
+
+Although HOW manages to generate plans by localizing displacements to regions small enough to produce a statistically significant change, it needs to be noted that it fails to provide succinct summaries.
+In addition to this, plans derived from HOW
 
@@
 
-	Data can be rapidly and recursively divided   as follows.
-	Find   two   distance cases,  $X,Y$
+		The data is recursively divided in clusters using WHERE as follows:
+		\begin{itemize}
+
+			\item Find   two   distance cases,  $X,Y$
 			by picking any case $W$ at random, then setting $X$ to its most
 			distant case, then setting $Y$ to the case most distant from
-	$X$~\cite{fastmap}
+			$X$
 			(this requires only $O(2N)$ comparisons
 			of $N$ cases).
-	Next, HOW projects each case $Z$
+			\item Project each case $Z$
 			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
-	rule of \fig{where}b. After that,  split the data at the median $x$ value of all cases and
+			rule. 
+			\item Split the data at the median $X$ value of all cases and
 			recurses on each half  (stopping when
 			one half has less  than $\sqrt{N}$ of the original population).
+		\end{itemize}
 
@@
 
-		The data is recursively divided in clusters using WHERE as follows:
-		\begin{itemize}
+		The data is recursively divided in clusters using WHERE. For more information, see \fig{where}a.
 
-			\item Find   two   distance cases,  $X,Y$
-			by picking any case $W$ at random, then setting $X$ to its most
-			distant case, then setting $Y$ to the case most distant from
-			$X$
-			(this requires only $O(2N)$ comparisons
-			of $N$ cases).
-			\item Project each case $Z$
-			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
-			rule. 
-			\item Split the data at the median $X$ value of all cases and
-			recurses on each half  (stopping when
-			one half has less  than $\sqrt{N}$ of the original population).
-		\end{itemize}
 		~\hrule~

### 2015-08-08 Manuscript.tex
Updates from @rahlk's local machine.

@@
 
-Although HOW manages to generate plans by localizing displacements to regions small enough to produce a statistically significant change, it needs to be noted that it fails to provide succinct summaries.
-In addition to this, plans derived from HOW
+Although HOW manages to generate plans by localizing displacements to regions small enough to produce a statistically significant improvement, it needs to be noted that it fails to provide succinct summaries.
+Lack of succinctness makes it difficult draw generalizable conclusions about the test data.
+
+This is a trend commonly observed in most CBR systems.
+They tend to reason directly from a loaded training data instead of first summarizing the data into a model.
+However, not all domains come with reliable models.
+For instance, a model that encompasses all the intricate issues that may lead to defects in software would be very large and indeed rather complex.
+In addition to this, finding empirical data to validate such models can be hard to come by and also time consuming ~\cite{me09i,me09j}.
+Fortunately, this issue can be circumvented by using decision trees.
 
@@
 \subsubsection{Decision Trees}
+
+Decision tress categorically  sort the training data into a tree structure which can be viewed as model to reason about the data. In theory, this is a very desirable approach as it promises a
 

### 2015-08-08 Manuscript.tex
Updates from @rahlk's local machine.

@@
 
-\begin{figure}[t!]
+%\begin{figure}[t!]
+%	\small
+%	~\hrule~
+%	
+%	{\bf \fig{where}a: top down clustering:}
+%	
+%		The data is recursively divided in clusters using WHERE as follows:
+%		\begin{itemize}
+%			
+%			\item Find   two   distance cases,  $X,Y$
+%			by picking any case $W$ at random, then setting $X$ to its most
+%			distant case, then setting $Y$ to the case most distant from
+%			$X$
+%			(this requires only $O(2N)$ comparisons
+%			of $N$ cases).
+%			\item Project each case $Z$
+%			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
+%			rule. 
+%			\item Split the data at the median $X$ value of all cases and
+%			recurses on each half  (stopping when
+%			one half has less  than $\sqrt{N}$ of the original population).
+%		\end{itemize}
+%	
+%	~\hrule~
+%	
+%	{\bf \fig{where}b: finding distances:}
+%	
+%	In the \fig{where}a, to compute distances, we use
+%	the Euclidean measure recommended for
+%	case-based reasoning by Aha et al.~\cite{aha91};
+%	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
+%	where values are  normalized 0..1 for the range min..max and 
+%	$w_i$ is some importance weight given to attribute $i$.
+%	Usually, $w_i=1$ but it can be adjusted using, say,
+%	the feature weighting methods of \fig{where}d. 
+%	
+%	~\hrule~
+%	
+%	{\bf \fig{where}c: 3 point geometry:}
+%	
+%	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
+%	A third point, $Z$, has distances  $a=dist(Z,X)$ and
+%	$b=dist(Z,Y)$ to $X,Y$, respectively.
+%	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
+%	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
+%	Further, according to Pythagoras' theorem, $Z$ stands at a distance
+%	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. 
+%	
+%	~\hrule~
+%	
+%	{\bf \fig{where}d: feature weighting:}
+%	
+%	HERE's feature weighting algorithm
+%	comes from the CART regression tree learner~\cite{Breiman1984}.
+%	It sorts independent variables
+%	according to how well they can reduce the variability
+%	of a  numeric objective.
+%	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
+%	then the expected
+%	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
+%	where $v$ is the standard deviation of the objective.
+%	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
+%	then split at $\argmin_i w_i$, then recursing into each half.
+%	
+%	~\hrule~
+%	
+%	\caption{Sub-routines within HOW.}\label{fig:where}
+%\end{figure}
+
+\subsubsection{Decision Trees}
+
+\section{CROSSTREES}
+\subsection{Design}
+	\begin{figure}[t]
 		\small
@@
 
-	{\bf \fig{where}a: top down clustering:}
-
+		{\bf Step 1: Top down clustering using WHERE}
 		The data is recursively divided in clusters using WHERE as follows:
@@
 
-	{\bf \fig{where}b: finding distances:}
-
-	In the \fig{where}a, to compute distances, we use
-	the Euclidean measure recommended for
-	case-based reasoning by Aha et al.~\cite{aha91};
-	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
-	where values are  normalized 0..1 for the range min..max and 
-	$w_i$ is some importance weight given to attribute $i$.
-	Usually, $w_i=1$ but it can be adjusted using, say,
-	the feature weighting methods of \fig{where}d.
-
-	~\hrule~
-
-	{\bf \fig{where}c: 3 point geometry:}
-
-	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
-	A third point, $Z$, has distances  $a=dist(Z,X)$ and
-	$b=dist(Z,Y)$ to $X,Y$, respectively.
-	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
-	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
-	Further, according to Pythagoras' theorem, $Z$ stands at a distance
-	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$.
-
-	~\hrule~
-
-	{\bf \fig{where}d: feature weighting:}
-
-	HERE's feature weighting algorithm
-	comes from the CART regression tree learner~\cite{Breiman1984}.
-	It sorts independent variables
-	according to how well they can reduce the variability
-	of a  numeric objective.
-	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
-	then the expected
-	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
-	where $v$ is the standard deviation of the objective.
-	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
-	then split at $\argmin_i w_i$, then recursing into each half.
-
-	~\hrule~
-
-	\caption{Sub-routines within HOW.}\label{fig:where}
-\end{figure}
-
-\subsubsection{Decision Trees}
-
-Decision tress categorically  sort the training data into a tree structure which can be viewed as model to reason about the data. In theory, this is a very desirable approach as it promises a
-
-\section{CROSSTREES}
-\subsection{Design}
-	\begin{figure}[t]
-		\small
-		~\hrule~
-
-		{\bf Step1: Top down clustering using WHERE}
-
-		The data is recursively divided in clusters using WHERE. For more information, see \fig{where}a.
-
-		~\hrule~
-
-		{\bf Step2: Distinguish between clusters using  decision trees}
+		{\bf Step 2: Distinguish between clusters using  decision trees}
 
@@
 
-		{\bf Step3: Generating contrast sets from DT branches}
+		{\bf Step 3: Generating contrast sets from DT branches}
 		\begin{itemize}

### 2015-08-08 Manuscript.tex
Updates from @rahlk's local machine.

@@
 \section{Introduction}
+
 \section{Cluster and Contrast}
+
 A recurring data analysis pattern in this paper is $cluster+contrast$.
@@
 These lessons are used to generate rules that can be applied in any context.
+
 \subsection{Clustering}
@@
 
-Clustering is followed generating \textit{contrast sets}.
+Clustering is followed by generating \textit{contrast sets}.
 These contrast sets represent recommendations on what could be altered to better improve an outcome.
@@
 
-\subsubsection{Case Based Reasoning}
+\subsection{Case Based Reasoning}
 
@@
 \begin{itemize}
-\item[1.] Displacements are seldom limited to small regions, this implies that large changes may be recommended when not necessary. This may make the solutions less feasible to implement. 
+\item[1.] Displacements are seldom limited to small regions, this implies that large changes may be recommended when they are not necessary. This may make the solutions less feasible to implement. 
 \item[2.] The nearness of a test case to a cluster in the training data is subjective.
@@
 
-\subsubsection{HOW}
+\subsection{HOW}
 
@@
 HOW works by clustering the data during training using WHERE and then drawing \texttt{slopes} between the centroids of pairs of nearby clusters.
-Assuming the cluster pairs are labeled X and Y, with X having slightly better performance score than Y, the \texttt{slope} between X and Y acts as an indicator pointing to a direction to displace the data; i.e.
+Assuming the cluster pairs are labeled X and Y, with X having slightly better performance score than Y, the \texttt{slope} between X and Y acts as an indicator pointing to a direction in which to displace the data; i.e.
 away from Y and towards X.
@@
 Lack of succinctness makes it difficult draw generalizable conclusions about the test data.
-
 This is a trend commonly observed in most CBR systems.
 They tend to reason directly from a loaded training data instead of first summarizing the data into a model.
-However, not all domains come with reliable models.
+
+% However, not all domains come with reliable models.
 For instance, a model that encompasses all the intricate issues that may lead to defects in software would be very large and indeed rather complex.
 In addition to this, finding empirical data to validate such models can be hard to come by and also time consuming ~\cite{me09i,me09j}.
-Fortunately, this issue can be circumvented by using decision trees.
 
-%\begin{figure}[t!]
-%	\small
-%	~\hrule~
-%	
-%	{\bf \fig{where}a: top down clustering:}
-%	
-%		The data is recursively divided in clusters using WHERE as follows:
-%		\begin{itemize}
-%			
-%			\item Find   two   distance cases,  $X,Y$
-%			by picking any case $W$ at random, then setting $X$ to its most
-%			distant case, then setting $Y$ to the case most distant from
-%			$X$
-%			(this requires only $O(2N)$ comparisons
-%			of $N$ cases).
-%			\item Project each case $Z$
-%			onto a {\tt Slope} that  runs between $X,Y$ using the cosine
-%			rule. 
-%			\item Split the data at the median $X$ value of all cases and
-%			recurses on each half  (stopping when
-%			one half has less  than $\sqrt{N}$ of the original population).
-%		\end{itemize}
-%	
-%	~\hrule~
-%	
-%	{\bf \fig{where}b: finding distances:}
-%	
-%	In the \fig{where}a, to compute distances, we use
-%	the Euclidean measure recommended for
-%	case-based reasoning by Aha et al.~\cite{aha91};
-%	i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
-%	where values are  normalized 0..1 for the range min..max and 
-%	$w_i$ is some importance weight given to attribute $i$.
-%	Usually, $w_i=1$ but it can be adjusted using, say,
-%	the feature weighting methods of \fig{where}d. 
-%	
-%	~\hrule~
-%	
-%	{\bf \fig{where}c: 3 point geometry:}
-%	
-%	Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
-%	A third point, $Z$, has distances  $a=dist(Z,X)$ and
-%	$b=dist(Z,Y)$ to $X,Y$, respectively.
-%	According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
-%	at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
-%	Further, according to Pythagoras' theorem, $Z$ stands at a distance
-%	$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. 
-%	
-%	~\hrule~
-%	
-%	{\bf \fig{where}d: feature weighting:}
-%	
-%	HERE's feature weighting algorithm
-%	comes from the CART regression tree learner~\cite{Breiman1984}.
-%	It sorts independent variables
-%	according to how well they can reduce the variability
-%	of a  numeric objective.
-%	If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
-%	then the expected
-%	value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
-%	where $v$ is the standard deviation of the objective.
-%	A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
-%	then split at $\argmin_i w_i$, then recursing into each half.
-%	
-%	~\hrule~
-%	
-%	\caption{Sub-routines within HOW.}\label{fig:where}
-%\end{figure}
-
-\subsubsection{Decision Trees}
-
-\section{CROSSTREES}
-\subsection{Design}
 \begin{figure}[t]
@@
 
-\subsection{Assessment}
+\subsection{CROSSTREES}
+
+Like all the previous planners, CROSSTREES starts by clustering the training data.
+Following this, it executes a decision tree algorithm using the clusters as class labels.
+Using a decision tree algorithm allows us to identify the most informative attributes to select in different clusters.
+The decision tree built in this fashion emulates a model built from the training data.
+
+Using the decision tree, the test cases can be categorized into one of the branches of the tree.
+Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}?
+The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}.
+For full algorithmic details, refer to \fig{xtrees_bare}.
+
+CROSSTREES offer a great number of benefits compared to the other methods that were discussed above.
+Firstly, they offer a visual medium for experts to identify and explore solutions spaces that are local to the problem.
+Secondly, they offer solutions that are much more stable than other instance based learners.
+This stability can be attributed to the consistency and general reproducibility of the tree structure.
+Thirdly, and perhaps most importantly, the tree summarizes the training data succinctly, making it easier for a business user to examine the solutions.
+
+A tree structure such as CROSSTREES are characterized by attributes such as size and depth.
+It is worth noting that these attributes have a profound impact on its performance.
+Our initial motivation for using CROSSTREES is that they can serve as a medium for experts to reason about a data.
+The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity.
+We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare})
+
 \section{Experiments}
+
 \subsection{Data}
 
+\begin{figure*}[htbp!]
+	\renewcommand{\baselinestretch}{0.8}\begin{center}
+		{\scriptsize
+			\begin{tabular}{c|l|p{4.7in}}
+				amc & average method complexity & e.g. number of JAVA byte codes\\\hline
+				avg\, cc & average McCabe & average McCabe's cyclomatic complexity seen
+				in class\\\hline
+				ca & afferent couplings & how many other classes use the specific
+				class. \\\hline
+				class. \\\hline
+				cam & cohesion amongst classes & summation of number of different
+				types of method parameters in every method divided by a multiplication
+				of number of different method parameter types in whole class and
+				number of methods. \\\hline
+				cbm &coupling between methods &  total number of new/redefined methods
+				to which all the inherited methods are coupled\\\hline
+				cbo & coupling between objects & increased when the methods of one
+				class access services of another.\\\hline
+				ce & efferent couplings & how many other classes is used by the
+				specific class. \\\hline
+				dam & data access & ratio of the number of private (protected)
+				attributes to the total number of attributes\\\hline
+				dit & depth of inheritance tree &\\\hline
+				ic & inheritance coupling &  number of parent classes to which a given
+				class is coupled (includes counts of methods and variables inherited)
+				\\\hline
+				lcom & lack of cohesion in methods &number of pairs of methods that do
+				not share a reference to an case variable.\\\hline
+				locm3 & another lack of cohesion measure & if $m,a$ are  the number of
+				$methods,attributes$
+				in a class number and $\mu(a)$  is the number of methods accessing an
+				attribute, 
+				then
+				$lcom3=((\frac{1}{a} \sum, j^a \mu(a, j)) - m)/ (1-m)$.
+				\\\hline
+				loc & lines of code &\\\hline
+				max\, cc & maximum McCabe & maximum McCabe's cyclomatic complexity seen
+				in class\\\hline
+				mfa & functional abstraction & number of methods inherited by a class
+				plus number of methods accessible by member methods of the
+				class\\\hline
+				moa &  aggregation &  count of the number of data declarations (class
+				fields) whose types are user defined classes\\\hline
+				noc &  number of children &\\\hline
+				npm & number of public methods & \\\hline
+				rfc & response for a class &number of  methods invoked in response to
+				a message to the object.\\\hline
+				wmc & weighted methods per class &\\\hline
+				\rowcolor{lightgray}
+				defect & defect & Boolean: where defects found in post-release bug-tracking systems.
+			\end{tabular}
+		}
+	\end{center}
+	\caption{OO measures used in our defect data sets.  Last line is
+		the dependent attribute (whether a defect is reported to  a
+		post-release bug-tracking system).}\label{fig:ck}
+\end{figure*}
+
+\subsection{When not to Plan?}
 \begin{figure*}[!t]
@@
 \end{figure*}
-\subsection{When not to Plan?}
-\subsection{Evaluation}
+
+\subsection{Assessment}
 \subsection{Results}

### 2015-08-09 Manuscript.tex
Updates from @rahlk's local machine.

@@
 The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity.
-We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare})
+We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).
 

### 2015-08-09 Manuscript.tex
Updates from ShareLaTeX

--- 2015-08-08 Manuscript.tex
@@
 
+Trustworthy domain models are a popular alternatives to CBR planners.
+In our previous work, we have used executing source code as a "model" to check if our mutations to test suites minimize the test suite size while maximizing the number of statements covered~\cite{me09m,andrews07,andrews10}.
+However, we do not always have access to ready-to-use models.
+Further, finding empirical data to validate existing models can be hard.
+Fortunately, this issue can be circumvented by
+
 Like all the previous planners, CROSSTREES starts by clustering the training data.

### 2015-08-14 Manuscript.tex
Updates from @rahlk's ShareLaTeX account.
Updates from @rahlk's ShareLaTeX Account

@@
 %\usepackage[pdftex]{graphicx}
+\usepackage{program}
 \newcommand{\Sample}{{\bf SAMPLE}}
@@
 \usepackage{picture}
-
+\usepackage{url}
+\usepackage{hyperref}
 \usepackage{listings}
-\usepackage{tabu}
 \DeclareMathOperator*{\argmin}{arg\,min} 
@@
 
+\definecolor{Code}{rgb}{0,0,0}
+\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
+\definecolor{Numbers}{rgb}{0.5,0,0}
+\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
+\definecolor{Keywords}{rgb}{0,0,1}
+\definecolor{self}{rgb}{0,0,0}
+\definecolor{Strings}{rgb}{0,0.63,0}
+\definecolor{Comments}{rgb}{0,0.63,1}
+\definecolor{Comments}{rgb}{0.5,0.5,0.5}
+\definecolor{Backquotes}{rgb}{0,0,0}
+\definecolor{Classname}{rgb}{0,0,0}
+\definecolor{FunctionName}{rgb}{0,0,0}
+\definecolor{Operators}{rgb}{0,0,0}
+\definecolor{Background}{rgb}{1,1,1}
+\lstnewenvironment{code}[1]{
 \lstset{
+mathescape,
+numbers=left,
+numberstyle=\scriptsize,
+stepnumber=1,
+numbersep=0.5em,
+xleftmargin=1em,
+framextopmargin=2em,
+framexbottommargin=2em,
+showspaces=false,
+showtabs=false,
+showstringspaces=false,
+tabsize=2,
+% Basic
+basicstyle=\ttfamily\scriptsize,
+backgroundcolor=\color{Background},
 language=Python,
-    basicstyle=\ttfamily\fontsize{2.4mm}{0.8em}\selectfont,
-    breaklines=true,
-    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
-    frame=tlrb,
-    showtabs=false,
-    showspaces=false,
-    showstringspaces=false,
-    %backgroundcolor=\color{Gray},
-    keywordstyle=\bfseries,
-    emph={COCONUT,GUESSES,ASSESS,COCOMO2,PEEKING2,SAMPLE,WHERE,RIG}, emphstyle=\bfseries\color{Blue},
-    stringstyle=\color{green!50!black},
-    commentstyle=\color{red}\itshape,
-    %numbers=none,
-    captionpos=t,
-    numberstyle=\bfseries\color{red},
-    escapeinside={\%*}{*)}
-}
+% Comments
+commentstyle=\color{Comments}\slshape,
+% Strings
+stringstyle=\color{Strings},
+morecomment=[s][\color{Strings}]{"""}{"""},
+morecomment=[s][\color{Strings}]{'''}{'''},
+% keywords
+morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
+keywordstyle={[1]\color{Code}\bfseries},
+% additional keywords
+morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
+keywordstyle={[3]\color{Keywords}\bfseries},
+morekeywords={[2]@invari},
+keywordstyle={[2]\color{Decorators}\slshape},
+emph={self},
+emphstyle={\color{self}\slshape},
+firstnumber=last
+%
+}}{}
 
@@
 \newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
-%\newcommand{\ADD}[1]{#1}
 \usepackage{times}
-\def\baselinestretch{1}
-\setlist{nosep}
-\usepackage[font={small}]{caption, subfig}
-
-\setlength{\abovecaptionskip}{1ex}
-\setlength{\belowcaptionskip}{1ex}
-\setlength{\floatsep}{1ex}
-\setlength{\textfloatsep}{1ex}
-\newcommand{\subparagraph}{} % defined before loading titlesec
-\usepackage[compact,small]{titlesec}
-\DeclareMathSizes{7}{7}{7}{7} 
-\bibliographystyle{unsrt}
 \pagenumbering{arabic} 
@@
 \begin{abstract}
-This paper provides ...  
+ As business users demand more insightful
+ analytics, data scientists need to change
+ their tools. Instead of merely predicting 
+ some result, they also need tools that generate ``plans'';
+ i.e. specific suggestions on  how to change values in order to
+ improve on some predicted value.
+
+ This paper is aims to study such planners.
+The planners fall into three broad categories (1) Cluster specific planners (K Nearest Neighbors); (2) Instance specific planners (HOW); (3) Decision Tree based planners.
+These planners are employed as tools for proposing changes to independent
+ variables in order to improve on 
+ the predictions of the dependent variables. Unlike other approaches
+ for learning optimizations to software, these planners do not require
+ an underlying model of the domain. Also, these plans
+ are designed to result in {\em minimal displacement}
+ to current practice.
+
+ This paper uses these learners to learn methods
+ to reduce defects and decrease program runtime.
+ For the test data of this paper, those improvements   reduce
+ the expected values of defect counts and  runtimes to    
+ 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values, respectively.cy
 \end{abstract}
 \section{Introduction}
+	How should we handle ``unpopulular'' results
+	for data analytics? For example, if a business manager is presented
+	with a data mining result that troubles them (e.g. an estimate of
+	development time that is distressingly long), what advice
+	can we offer them on management actions to {\em change} that estimate?
+
+	This is an important question, and one with much currency in industry.
+	In the summer of 2011 and 2012, one of us (Menzies) spent two months
+	working on-site at Microsoft Redmond,
+	observing data mining analysts.  He
+	noted how Microsoft's data scientists
+	discussed their data with  business users. 
+	One surprising outcome of that study was just how
+	little time was spent  
+	inspecting  of the output of data miners as compared to another process, which we call {\em peeking}.
+	In {\em peeking}, analysts and users spend much time
+	inspecting and discussing small samples of either raw or exemplary or synthesized project data.  Further, very little of those discussions were  focused on classification
+	(the addition of a labels to some unlabelled data). Rather, much time
+	was spent in those meetings discussing {\em what to do next}; i.e. trying
+	to determine what could be altered to better improve some business outcome.
+
+	That   Microsoft  study found two common ``peeking'' methods.
+	In {\em data engagement meetings},
+	users debated the implications of data
+	displayed on a screen. In this way, users
+	engaged with the data and with each other by
+	monitoring each others' queries and check each others'
+	conclusions.
+
+	Another data analysis pattern observed
+	at Microsoft was  {\em cluster + contrast} in which
+	data is  reduced to a few
+	clusters. Users are then just shown the delta between those
+	clusters. While contrasting, if feature values are
+	the same in both clusters, then these were pruned from
+	the reports to the user. In this way, very large
+	data sets can be shown on one PowerPoint
+	slide. Note that {\em cluster+contrast} is a tool that can be usefully employed within
+	{\em data engagement meetings}.
+
+	Cluster+contrast and engagement
+	meetings are common practices at Microsoft. Yet  these methods had never been rigorously studied or certified.
+	For both those reasons,
+	we reflected over those tools to discover and analyze their
+	underlying process. The result was HOW~\cite{howase}: a tool
+	that combines (a)~feature selection; (b)~centroid generation from   clusters;
+	(c)~contrast methods between centroids.
+	While method (a) is widely used (e.g.~\cite{Menzies2010}),
+	to the best of our knowledge, this combination of (abc) has not been thoroughly explored before.
+
+This paper assesses two methods  for ``peeking'': a model-based method called CROSSTREES and an instance-based method   called HOW. 
+CROSSTREES were first developed by Lekkalapudi and Menzies to  explaining results from multi-objective optimizers~\cite{nva14}. This paper is the first
+to apply CROSSTREES to defect prediction and contrast set learning. Also, that prior work evaluated
+CROSSTREES against multi-objective optimizers and not   instance-based methods.
+
+This paper uses three criteria to assess the value of HOW and CROSSTREES for
+learning actionable analytics:
+\be
+\item
+A planning system needs to be {\em effective}; i.e. if its recommendations
+are applied then some statistically significant change should be observed.
+To predict the number of defects in a data set before and after applying our changes,
+we build a    prediction system (built by data mining; specifically: Random Forest). Note that this
+predictor was built from some hold-out data (i.e. from  different data than that used
+to build the predictor).
+\item
+Any  conclusion made to a business user must be understandable;
+it must generate {\em succinct} changes.  
+\item
+Recommendations should be {\em stable}; i.e.
+they shouldn't widely vary due to minor changes in the data.
+Hence, in our experiments, we will add a little randomness to our analysis then report results across 20 repeated runs.
+\ee
+We will find that
+ {\em effectiveness} of CROSSTREES and HOW are similar, but   CROSSTREES wins on {\em succinctness} and {\em stability}.
 
 \section{Cluster and Contrast}
-
+\label{clust_contrast}
 A recurring data analysis pattern in this paper is $cluster+contrast$.
@@
 With this in mind, we opine that it would be beneficial if the above method is extended to include some form of feature weighting, thus enabling the tools to recommend changes to only the most informative features.
-This paper uses a feature weighting scheme similar to that used in CART, see \fig{where}d.
+This paper uses a feature weighting scheme similar to that used in C4.5, see \fig{where}d.
 
@@
 In view of the above issues, we developed HOW, a tool that combines (a) Feature Selection; (b) Centroid generation from clusters; (c) Contrast generation using gradient between clusters.
+\input{tex/pseudo-code-1.tex}
 
@@
 
-\subsection{CROSSTREES}
+\subsection{DECISION TREES}
 
@@
 Further, finding empirical data to validate existing models can be hard.
-Fortunately, this issue can be circumvented by
 
-Like all the previous planners, CROSSTREES starts by clustering the training data.
-Following this, it executes a decision tree algorithm using the clusters as class labels.
-Using a decision tree algorithm allows us to identify the most informative attributes to select in different clusters.
-The decision tree built in this fashion emulates a model built from the training data.
+Taking into account the wealth of data that is available to us, this issue can fortunately be circumvented by constructing a categorical model based on decision trees.
+Unlike the previous planners which use where, here we construct a decision tree using information gain as a metric to find ideal splits in the data.
+This allows us to identify the most informative attributes to select.
+A decision tree (hereafter referred to as DTREE) built in this fashion emulates a model built from the training data.
 
-Using the decision tree, the test cases can be categorized into one of the branches of the tree.
+Using DTREE, the test cases can be categorized into one of the branches of the tree.
 Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}?
@@
 
-CROSSTREES offer a great number of benefits compared to the other methods that were discussed above.
-Firstly, they offer a visual medium for experts to identify and explore solutions spaces that are local to the problem.
-Secondly, they offer solutions that are much more stable than other instance based learners.
-This stability can be attributed to the consistency and general reproducibility of the tree structure.
-Thirdly, and perhaps most importantly, the tree summarizes the training data succinctly, making it easier for a business user to examine the solutions.
-
-A tree structure such as CROSSTREES are characterized by attributes such as size and depth.
+A tree structure such as DTREE is characterized by attributes such as size and depth.
 It is worth noting that these attributes have a profound impact on its performance.
-Our initial motivation for using CROSSTREES is that they can serve as a medium for experts to reason about a data.
+Our initial motivation for using DTREE was that they could serve as a medium for experts to reason about a data.
 The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity.
@@
 
+DTREE offers a great number of benefits compared to the other methods that were discussed above.
+Firstly, it offers a visual medium for experts to identify and explore solutions spaces that are local to the problem.
+Secondly, its solutions are much more stable than other cluster specific and instance specific learners.
+This stability can be attributed to the consistency and general reproducibility of the tree structure.
+Thirdly, and perhaps most importantly, the tree summarizes the training data succinctly, making it easier for a business user to examine the solution sets.
+
 \section{Experiments}
+
+\begin{figure}[!t]
+\small 
+\[
+\begin{array}{r}
+\mathrm{project}\\
+\mathrm{data}
+\end{array} 
+\left\{\begin{array}{l}\mathit{train}
+        \left\{\begin{array}{l}
+                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
+                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
+              \end{array}\right.
+       \\
+      ~\\
+\mathit{test}  
+    \left\{\begin{array}{l@{~}l}
+           \mathit{actual}& =\mathrm{Defects \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
+           \mathrm{\bf if\;}\mathit{actual} & >  \mathit{0}\\
+           \mathrm{\bf then} &
+           \left\{
+            \begin{array}{l}
+                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
+                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
+                \mathrm{{\bf return}\;} \mathit{actual},\mathit{after}
+            \end{array}
+          \right.
+   \end{array}\right.
+\end{array} \right. 
+\]
+
+\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
+\end{figure}
+
+\subsection{Setup}
+
+In the experiments shown below,  we use the planners to comment on how to reduce
+software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
+tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}.
+
+\fig{work} shows our proposal to verify if the recommendations made by our planners actually work. Accordingly, we divide the
+project data  into two disjoint sets {\em train} and {\em test}
+(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
+Next, from the train set, we build both a {\em planner} as well
+as a {\em  predictor}.
+Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper, our {\em planner} will be from \ref{clust_contrast} and our {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} and Random Forest Regressor taken from the SciKit Learn toolkit~\cite{Pedregosa2012}.
+
+As for the {\em test} data, this is passed to the { predictor}
+to measure the baseline prediction accuracy, false alarm, and recall.
+If our { predictors} fail to perform adequately on the test data,
+then we cannot trust them to comment on our plans. Accordingly,
+if that performance is unsatisfactory, we abort.
+Else, we (1)~apply the { planner} to alter the {\em test} data;
+then (2)~apply the { predictor} to the altered data $test'$;
+then (3)~return data on the {\em before, after} predictions.
 
@@
 
-\begin{figure*}[htbp!]
-	\renewcommand{\baselinestretch}{0.8}\begin{center}
-		{\scriptsize
-			\begin{tabular}{c|l|p{4.7in}}
-				amc & average method complexity & e.g. number of JAVA byte codes\\\hline
-				avg\, cc & average McCabe & average McCabe's cyclomatic complexity seen
-				in class\\\hline
-				ca & afferent couplings & how many other classes use the specific
-				class. \\\hline
-				class. \\\hline
-				cam & cohesion amongst classes & summation of number of different
-				types of method parameters in every method divided by a multiplication
-				of number of different method parameter types in whole class and
-				number of methods. \\\hline
-				cbm &coupling between methods &  total number of new/redefined methods
-				to which all the inherited methods are coupled\\\hline
-				cbo & coupling between objects & increased when the methods of one
-				class access services of another.\\\hline
-				ce & efferent couplings & how many other classes is used by the
-				specific class. \\\hline
-				dam & data access & ratio of the number of private (protected)
-				attributes to the total number of attributes\\\hline
-				dit & depth of inheritance tree &\\\hline
-				ic & inheritance coupling &  number of parent classes to which a given
-				class is coupled (includes counts of methods and variables inherited)
-				\\\hline
-				lcom & lack of cohesion in methods &number of pairs of methods that do
-				not share a reference to an case variable.\\\hline
-				locm3 & another lack of cohesion measure & if $m,a$ are  the number of
-				$methods,attributes$
-				in a class number and $\mu(a)$  is the number of methods accessing an
-				attribute, 
-				then
-				$lcom3=((\frac{1}{a} \sum, j^a \mu(a, j)) - m)/ (1-m)$.
-				\\\hline
-				loc & lines of code &\\\hline
-				max\, cc & maximum McCabe & maximum McCabe's cyclomatic complexity seen
-				in class\\\hline
-				mfa & functional abstraction & number of methods inherited by a class
-				plus number of methods accessible by member methods of the
-				class\\\hline
-				moa &  aggregation &  count of the number of data declarations (class
-				fields) whose types are user defined classes\\\hline
-				noc &  number of children &\\\hline
-				npm & number of public methods & \\\hline
-				rfc & response for a class &number of  methods invoked in response to
-				a message to the object.\\\hline
-				wmc & weighted methods per class &\\\hline
+We applied planners to reduce (a) the number of defects found in Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}; (b) the runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
+\fig{ck} shows Jureczko data attributes and \fig{cpm} shows the  Seigmund data attributes which were more diverse than the Jurecko data since they contain the results of applying many  mutators to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  C versions),  Apache,  SQLite,  LLVM, and x264.
+
+\input{tex/jur_char.tex}
+\begin{figure}[!t]
+\scriptsize
+\begin{tabular}{llllll}
+  \hline
   \rowcolor{lightgray}
-				defect & defect & Boolean: where defects found in post-release bug-tracking systems.
+Project & Domain & Lang. & LOC & Features & Config\\\hline
+Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
+Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
+Apache & Web Server & C & 230,277 & 9 & 192\\
+SQLite & Database & C & 312,625 & 39 & 3,932,160\\
+LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
+x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
 \end{tabular}
-		}
-	\end{center}
-	\caption{OO measures used in our defect data sets.  Last line is
-		the dependent attribute (whether a defect is reported to  a
-		post-release bug-tracking system).}\label{fig:ck}
+
+\caption{Seigmund data sets: details on the data sets.
+For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
+\end{figure}
+
+\input{tex/results_jur.tex}
+\input{tex/results_conf.tex}
+\input{tex/prediction_stats.tex}
+
+\begin{figure*}[t]
+\begin{minipage}{0.5\textwidth}
+\includegraphics[width=0.9\linewidth]{Deltas-Ivy.eps}
+\label{fig:delta_ivy}
+
+\end{minipage}
+\begin{minipage}{0.5\textwidth}
+\includegraphics[width=0.9\linewidth]{Deltas-BDBJ.eps}
+\label{fig:delta_apache}
+\end{minipage}
+\caption{Performance Comparison showing the number of times a certain feature requires to be changed.
+X-axis lists the various features that make up a data set.
+The y-axis contains the frequency of change in \%.}
+
 \end{figure*}
 
-\subsection{When not to Plan?}
-\begin{figure*}[!t]
-	\scriptsize
-	\begin{center}
-		\begin{minipage}{.46\linewidth}
-			\begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}
-				& \multicolumn{5}{c|}{ }\\
+The clear advantage of our current work is that it can be readily and automatically applied to domains with historical data.
+However, the obvious drawback with this approach is that our assessment of the plans is only as good as the predictor.
+To mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for {\em satisfactory} thresholds.
 
-				& \multicolumn{5}{c|}{ Data set  properties}\\ 
-				& \multicolumn{5}{c|}{  }\\ 
-				& \multicolumn{2}{c|}{training}   & \multicolumn{3}{c|}{testing}      \\ \cline{2-6}
-				data set      & versions           & cases & versions     & cases    & \% defective             \\ \hline
-				jedit    & 3.2, 4.0, 4.1, 4.2 & 1257      & 4.3          & 492          & 2 \\
-				ivy      & 1.1, 1.4           & 352       & 2.0          & 352          & 11 \\
-				camel    & 1.0, 1.2, 1.4      & 1819      & 1.6          & 965          & 19 \\
-				ant      & 1.3, 1.4, 1.5, 1.6 & 947       & 1.7          & 745          & 22 \\
-				synapse  & 1.0, 1.1           & 379       & 1.2          & 256          & 34 \\
-				velocity & 1.4, 1.5           & 410       & 1.6          & 229          & 34 \\
-				lucene   & 2.0, 2.2           & 442       & 2.4          & 340          & 59 \\
-				poi      & 1.5, 2, 2.5        & 936       & 3.0          & 442          & 64 \\
-				xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
-				log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
-				xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline
+For the Jureczko OO data sets, we found:
+\bi
+\item
+Most predictors in this domain were far from perfect;
+\item 
+But several of these predictors could
+be salvaged using some sampling and optimization methods (see the notes on SMOTE and differential evolution, below).
+\ei
+As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.}.
+Given access to $V$ released
+versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
+shown in \fig{j}, this means that we are training on hundreds to thousands
+of classes and testing on smaller test suites).
+Note the three bottom rows marked with $\times$: these contain predominately
+defective classes (two-thirds, or more) and in such data sets, it is hard to distinguish
+good from bad (since there are so many bad examples).
 
-			\end{tabular}\end{minipage}\begin{minipage}{.4\linewidth}
-			\begin{tabular}{|rrr|rrr|rr|l} \cline{1-8}
-				\multicolumn{8}{|c|}{  }\\
-				\multicolumn{8}{|c|}{  Results from learning}\\
-				\multicolumn{8}{|c|}{   }\\
-				\multicolumn{3}{|c|}{untuned} & \multicolumn{3}{c|}{tuned} & \multicolumn{2}{c|}{change}\\
-				\cline{1-8}
+For the  Jureczko   data,
+the plans generated by our planners advise on  how to change static code attributes in order to reduce defects in Java classes. Here, prediction performance of Random Forest can be measured using:
+\bi
+\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
+the test data detected
+by the {\em predictor}.
+\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
+classes that are {\em predicted} to be defective.
+\ei
 
-				pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
-				55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
-				65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
-				49 & 31 &   & 56 & 37 &   & 5 & 6\\
-				49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
-				45 & 19 &   & 47 & 15 &   & 2 & -4\\
-				78 & 60 &   & 76 & 60 &   & -2 & 0\\
-				56 & 25 &   & 60 & 25 & y & 4 & 0\\
-				56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
-				30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
-				32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
-				38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
-				\hline 
-			\end{tabular}
+The ``untuned'' columns of \fig{j} shows Random Forest's {\em pd,pf}
+values generated after learning from the training data and then being applied to the test data.
+If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
+then only two of our data sets ({\em ivy,ant}) are ``good'' enough.
 
-		\end{minipage}
-	\end{center}
+However, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples.
+Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
 
-	\caption{Training and test {\em data set properties} for the Jureczko data sets,
-		sorted in ascending order of \% defective examples.
-		On the right-hand-side, we show the {\em results from learning}.
-		Data is ``good'' if it has recall over 60\% and false alarm under 40\%
-		(and note that, after tuning, there are more ``good'' than before).
-		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
-		Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
-	}\label{fig:j}
-\end{figure*}
+Turning now to the Seigmund data, the lessons learnt from the planners advise on how  to change the configurations settings within a Makefile (in order to make resulting compiled code  run faster).
+Here, performance can be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s=\frac{abs(a - p)}{a}$.
+
+This paper  explores six Seigmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
+  x264\footnote{Available from the performance prediction section of PROMISE
+  http://openscience.us/repo/performance-predict.}. 
+  After splitting that data into equal sized train:test groups, a Random Forest
+  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\]
+
+That is, we can be very confident that the predictors from the Seigmund data can assess
+our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)
+
+In summary, while we cannot trust predictors from some of our defect data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.
 
 \subsection{Assessment}
+\fig{work} showed how we build and test our planners.
+The dependent variables of Jureczko and Seigmund data sets are discrete and continuous in nature, respectively.
+Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Seigmund data.
+Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
+\bi
+\item For Seigmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
+\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
+\ei
+The 40 repeats defined above were repeated for each planner.
+A Scott-Knott procedure was used to test for statistical differences in the result.
+The Scott-Knott procedure was recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{mittas13}.
+
+This method
+sorts a list of $l$ treatments with $ls$ measurements by their median
+score. It then
+splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
+ differences  in the observed performances
+before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
+ $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
+Scott-Knott  applies an  hypothesis test $H$ to check
+if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.
+
+The hypothesis test $H$ used here is a conjunction of the A12 effect size test and a non-parametric bootstrap sampling.
+In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
+For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
+These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
+
 \subsection{Results}
+\input{tex/deltas_avg.tex}
+The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
+\subsection{Discussion}
 \section{Threats to Validity}
@@
 \end{document}
-
-\end{document}

### 2015-08-14 tex/jur_char.tex
Updates from @rahlk's ShareLaTeX Account

\begin{figure*}[t!]
	\renewcommand{\baselinestretch}{0.8}\begin{center}
		{\scriptsize
			\begin{tabular}{c|l|p{4.7in}}
				amc & average method complexity & e.g. number of JAVA byte codes\\\hline
				avg\, cc & average McCabe & average McCabe's cyclomatic complexity seen
				in class\\\hline
				ca & afferent couplings & how many other classes use the specific
				class. \\\hline
				class. \\\hline
				cam & cohesion amongst classes & summation of number of different
				types of method parameters in every method divided by a multiplication
				of number of different method parameter types in whole class and
				number of methods. \\\hline
				cbm &coupling between methods &  total number of new/redefined methods
				to which all the inherited methods are coupled\\\hline
				cbo & coupling between objects & increased when the methods of one
				class access services of another.\\\hline
				ce & efferent couplings & how many other classes is used by the
				specific class. \\\hline
				dam & data access & ratio of the number of private (protected)
				attributes to the total number of attributes\\\hline
				dit & depth of inheritance tree &\\\hline
				ic & inheritance coupling &  number of parent classes to which a given
				class is coupled (includes counts of methods and variables inherited)
				\\\hline
				lcom & lack of cohesion in methods &number of pairs of methods that do
				not share a reference to an case variable.\\\hline
				locm3 & another lack of cohesion measure & if $m,a$ are  the number of
				$methods,attributes$
				in a class number and $\mu(a)$  is the number of methods accessing an
				attribute, 
				then
				$lcom3=((\frac{1}{a} \sum, j^a \mu(a, j)) - m)/ (1-m)$.
				\\\hline
				loc & lines of code &\\\hline
				max\, cc & maximum McCabe & maximum McCabe's cyclomatic complexity seen
				in class\\\hline
				mfa & functional abstraction & number of methods inherited by a class
				plus number of methods accessible by member methods of the
				class\\\hline
				moa &  aggregation &  count of the number of data declarations (class
				fields) whose types are user defined classes\\\hline
				noc &  number of children &\\\hline
				npm & number of public methods & \\\hline
				rfc & response for a class &number of  methods invoked in response to
				a message to the object.\\\hline
				wmc & weighted methods per class &\\\hline
				\rowcolor{lightgray}
				defect & defect & Boolean: where defects found in post-release bug-tracking systems.
			\end{tabular}
		}
	\end{center}
	\caption{OO measures used in our defect data sets.  Last line is
		the dependent attribute (whether a defect is reported to  a
		post-release bug-tracking system).}\label{fig:ck}
\end{figure*}

### 2015-08-14 tex/prediction_stats.tex
Updates from @rahlk's ShareLaTeX Account

\begin{figure*}[!t]
	\scriptsize
	\begin{center}
		\begin{minipage}{.46\linewidth}
			\begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}
				& \multicolumn{5}{c|}{ }\\

				& \multicolumn{5}{c|}{ Data set  properties}\\ 
				& \multicolumn{5}{c|}{  }\\ 
				& \multicolumn{2}{c|}{training}   & \multicolumn{3}{c|}{testing}      \\ \cline{2-6}
				data set      & versions           & cases & versions     & cases    & \% defective             \\ \hline
				jedit    & 3.2, 4.0, 4.1, 4.2 & 1257      & 4.3          & 492          & 2 \\
				ivy      & 1.1, 1.4           & 352       & 2.0          & 352          & 11 \\
				camel    & 1.0, 1.2, 1.4      & 1819      & 1.6          & 965          & 19 \\
				ant      & 1.3, 1.4, 1.5, 1.6 & 947       & 1.7          & 745          & 22 \\
				synapse  & 1.0, 1.1           & 379       & 1.2          & 256          & 34 \\
				velocity & 1.4, 1.5           & 410       & 1.6          & 229          & 34 \\
				lucene   & 2.0, 2.2           & 442       & 2.4          & 340          & 59 \\
				poi      & 1.5, 2, 2.5        & 936       & 3.0          & 442          & 64 \\
				xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
				log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
				xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline

			\end{tabular}\end{minipage}\begin{minipage}{.4\linewidth}
			\begin{tabular}{|rrr|rrr|rr|l} \cline{1-8}
				\multicolumn{8}{|c|}{  }\\
				\multicolumn{8}{|c|}{  Results from learning}\\
				\multicolumn{8}{|c|}{   }\\
				\multicolumn{3}{|c|}{untuned} & \multicolumn{3}{c|}{tuned} & \multicolumn{2}{c|}{change}\\
				\cline{1-8}

				pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
				55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
				65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
				49 & 31 &   & 56 & 37 &   & 5 & 6\\
				49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
				45 & 19 &   & 47 & 15 &   & 2 & -4\\
				78 & 60 &   & 76 & 60 &   & -2 & 0\\
				56 & 25 &   & 60 & 25 & y & 4 & 0\\
				56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
				30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
				32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
				38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
				\hline 
			\end{tabular}

		\end{minipage}
	\end{center}

	\caption{Training and test {\em data set properties} for the Jureczko data sets,
		sorted in ascending order of \% defective examples.
		On the right-hand-side, we show the {\em results from learning}.
		Data is ``good'' if it has recall over 60\% and false alarm under 40\%
		(and note that, after tuning, there are more ``good'' than before).
		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
		Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
	}\label{fig:j}
\end{figure*}

### 2015-08-14 Manuscript.tex
Updates from ShareLaTeX

@@
 % additional keywords
-morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
+morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, FSel, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
 keywordstyle={[3]\color{Keywords}\bfseries},
@@
 \subsection{Finding Contrasts}
-All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability, shown as Step-1 in figure \ref{fig:xtrees_bare}.
+All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability.
+It works as follows:
+\bi
+\item Find   two   distance samples from the data, say  $X,Y$.
+This can be done by  picking any case $W$ at random, then setting $X$ to its most distant case, then setting $Y$ to the case most distant from $X$ (this requires only $O(2N)$ comparisons of $N$ cases).
+\item Project each case $Z$ onto a {\tt Slope} that  runs between $X,Y$ using the cosine rule. 
+\item Split the data at the median $X$ value of all cases and recurse on each half  (stopping when one half has less  than $\sqrt{N}$ of the original population).
+\ei
 
@@
 
-\begin{figure}[t]
-	\small
-	~\hrule~
-
-	{\bf Step 1: Top down clustering using WHERE}
-	The data is recursively divided in clusters using WHERE as follows:
-	\begin{itemize}
-
-		\item Find   two   distance cases,  $X,Y$
-		by picking any case $W$ at random, then setting $X$ to its most
-		distant case, then setting $Y$ to the case most distant from
-		$X$
-		(this requires only $O(2N)$ comparisons
-		of $N$ cases).
-		\item Project each case $Z$
-		onto a {\tt Slope} that  runs between $X,Y$ using the cosine
-		rule. 
-		\item Split the data at the median $X$ value of all cases and
-		recurses on each half  (stopping when
-		one half has less  than $\sqrt{N}$ of the original population).
-	\end{itemize}
-
-	~\hrule~
-
-	{\bf Step 2: Distinguish between clusters using  decision trees}
-
-	Call each leaf from WHERE a  ``class''. Use an entropy-based
-	decision tree (DT) learner to learn what attributes select for each ``class''. To limit tree size:
-	\bi
-	\item Only use the top $\alpha=33$\%  of the features, as determined by their information gain~\cite{Irani1993}. 
-	\item Only build the trees down to  max depth of $\beta=10$.
-	\item Only build subtrees if it contains at least $N^{\gamma=0.5}$ examples (where $N$ is the size of the training set).
-	\ei
-	{\em Score}  DT  leaf nodes  via the mean score of its majority cluster.
-
-	~\hrule~
-
-	{\bf Step 3: Generating contrast sets from DT branches}
-	\begin{itemize}
-		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
-		\item Find the {\em desired} cluster: 
-		\bi
-		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
-		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
-		\item Using the {\em score} defined above, find the {\em better} siblings; i.e.
-those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}.
-If none found, then repeat for $lvl += 1$
-		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
-		\ei
-		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
-		\bi
-		\item
-		For discrete attributes,  return the value from {\em desired}. 
-		\item
-		For  numerics, return the numeric difference. 
-		\item
-		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
-		\ei
-		\ei
-		~\hrule~
-		\caption{CROSSTREES. Controlled by the parameters
-			$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
-		\label{fig:xtrees_bare}
-	\end{figure}
-
+\input{tex/dtree.tex}
 \subsection{DECISION TREES}
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{Deltas-Ivy.eps}
+\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
 \label{fig:delta_ivy}
-
 \end{minipage}
 \begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{Deltas-BDBJ.eps}
+\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
 \label{fig:delta_apache}
@@
 The y-axis contains the frequency of change in \%.}
-
-\end{figure*}
+\end{figure}
 
@@
 \ei
-As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.}.
+As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository~\cite{promiserepo}.}.
 Given access to $V$ released

### 2015-08-14 tex/dtree.tex
Updates from ShareLaTeX

\begin{figure}[t]
	\small
	~\hrule~

	{\bf Step 1: Distinguish between clusters using  decision trees}

	Use an entropy-based
	decision tree (DT) learner to learn what attributes select for a given data. To limit tree size:
	\bi
	\item Only use the top $\alpha=33$\%  of the features, as determined by their information gain~\cite{Irani1993}. 
	\item Only build the trees down to  max depth of $\beta=10$.
	\item Only build subtrees if it contains at least $N^{\gamma=0.5}$ examples (where $N$ is the size of the training set).
	\ei
	{\em Score}  DT  leaf nodes  via the mean score of its majority cluster.

	~\hrule~

	{\bf Step 2: Generating contrast sets from DT branches}
	\begin{itemize}
		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
		\item Find the {\em desired} cluster: 
		\bi
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}. If none found, then repeat for $lvl += 1$
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
		\bi
		\item
		For discrete attributes,  return the value from {\em desired}. 
		\item
		For  numerics, return the numeric difference. 
		\item
		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
		\ei
		\ei
		~\hrule~
		\caption{CROSSTREES. Controlled by the parameters
			$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
		\label{fig:xtrees_bare}
	\end{figure}

### 2015-08-20 Manuscript.tex
Updates from ShareLaTeX

@@
 
-\title{Prediction is Hard (Especially About the Future)}
-\numberofauthors{2} 
+\title{Learning Useful Changes\\(to Reduce Runtimes and Software Defects)}
+\numberofauthors{3} 
 \author{
 \alignauthor 
-Rahul Krishna \\
-       \affaddr{CS, NcState, USA}\\
-       {i.m.ralk@gmail.com}
+Rahul~Krishna,~Tim Menzies,~Xipeng~Shen \\
+       \affaddr{Com. Sci., NC State, USA}\\
+       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
 \alignauthor
-Tim Menzies\\
-       \affaddr{CS, NcState, USA}\\
-       {tim.menzies@gmail.com}}
+Naveen   Lekkalapudi\\
+ \affaddr{Software Consultant } \\ 
+       \affaddr{Washington D.C.}\\ 
+       {navek91@gmail.com}
+\alignauthor
+Lucas Layman \\
+       \affaddr{Fraunhofer CESE  } \\ 
+       \affaddr{College Park, USA}\\ 
+       {llayman@fc-md.umd.edu}
 \setlength{\columnsep}{7mm}
-
+}
 \maketitle
@@
  the expected values of defect counts and  runtimes to    
- 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values, respectively.cy
+ 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values, respectively.
 \end{abstract}
 \section{Introduction}
-	How should we handle ``unpopulular'' results
-	for data analytics? For example, if a business manager is presented
-	with a data mining result that troubles them (e.g. an estimate of
-	development time that is distressingly long), what advice
-	can we offer them on management actions to {\em change} that estimate?
+We propose a novel {\em planning} approach called CROSSTREES for learning changes to a software system
+such that its performance ``improvess'', according to some measure. For  casew
+stuides, we apply CROSSTREES to   (1)~reducethe predicted number
+of defects and (2)~reduce the runtimes of five system.
 
-	This is an important question, and one with much currency in industry.
+The approach discussed here is novel since most data miners used in software analytics
+comment on ``what is'' and not ``what to do''. For example, a clustering
+algorithm might find clumps of similar examples, but does not comment on the minimal
+change(s) required to move from clump to clump.
+
+Our approach is important for two reasons.
+Firstly, business users now demanding analytics tools that support a business-level interpretation of their data.
+At a panel on software analytics at ICSE’12, industrial practitioners lamented the state of the art in data mining and software engineering~\cite{menzies12a}.
+Panelists commented that ``prediction is all well and good, but what about decision making?''.
+That is, these panelists are more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
+For example:
+\bi
+\item
+Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
+\item
+Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
+\ei
+A second reason for the importance of our approach is, as shown below,
+when asked how to change a project in order to achieve a result, it is surprisingly
+difficult to use standard methods (e.g. clustering) to find those changes.
+In the case study that motivated this paper, we how  standard data mining methods
+can generate  seemingly
+useful insights, but which are in fact completely ineffectual. On the other hand,
+using CROSSTREES, we can propose changes that significantly reduce:
+\bi
+\item The  defects found  in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
+\item The runtime of the  software    configured by  Seigmund et al.~\cite{sven12}.
+\ei 
+Since CROSSTREES is a somewhat novel algorithm, we present it in tutorial form.
+The next section walks the reader through a sequence of approaches that start
+with a standard clustering algorithm, and progresses on to CROSSTREES. The second
+half of the paper then presents an extensive evaluation of CROSSTREES on 
+the Jureczko  and Seigmund data sets. This is followed by notes on related work
+and validity.
+
+\section{From Clustering to  CROSSTREES}
+
 In the summer of 2011 and 2012, one of us (Menzies) spent two months
 	working on-site at Microsoft Redmond,
-	observing data mining analysts.  He
-	noted how Microsoft's data scientists
-	discussed their data with  business users. 
-	One surprising outcome of that study was just how
-	little time was spent  
-	inspecting  of the output of data miners as compared to another process, which we call {\em peeking}.
+	observing data mining analysts.  In that study, he took special
+	note about how Microsoft's data scientists
+	discussed the results of their data mining sessions with  business users.
+
+	One surprising observation was how  
+	little time was spent by business users 
+	inspecting  of the output of standard data miners. Prior to that visit,
+	we had the mistaken impression that   decision trees,
+	clustering algorithms, etc were useful ``off-the-shelf''; i.e    business
+	users would inspect and understand the output of those tools.
+
+Standard data mining tools are not necessarily the best tool for supporting that dialogue.
+Menzies found that he had to do considerable work pre-processing data mining output
+prior to the weekly briefing meetings for the business users. Initially,
+that pre-processing was just clustering. This evolved into feature selection and finally
+the a case-based reasoning tool called HOW.
+
+	as compared to another process, which we call {\em peeking}.
 	In {\em peeking}, analysts and users spend much time
@@
 
-This paper assesses two methods  for ``peeking'': a model-based method called CROSSTREES and an instance-based method   called HOW. 
-CROSSTREES were first developed by Lekkalapudi and Menzies to  explaining results from multi-objective optimizers~\cite{nva14}. This paper is the first
-to apply CROSSTREES to defect prediction and contrast set learning. Also, that prior work evaluated
-CROSSTREES against multi-objective optimizers and not   instance-based methods.
+This paper assesses two methods  for ``peeking'': a model-based method called DTREE and an instance-based method   called HOW. 
+DTREE were first developed by Lekkalapudi and Menzies to  explaining results from multi-objective optimizers~\cite{nva14}. This paper is the first
+to apply DTREE to defect prediction and contrast set learning. Also, that prior work evaluated
+DTREE against multi-objective optimizers and not   instance-based methods.
 
-This paper uses three criteria to assess the value of HOW and CROSSTREES for
+This paper uses three criteria to assess the value of HOW and DTREE for
 learning actionable analytics:
@@
 We will find that
- {\em effectiveness} of CROSSTREES and HOW are similar, but   CROSSTREES wins on {\em succinctness} and {\em stability}.
+ {\em effectiveness} of DTREE and HOW are similar, but   DTREE wins on {\em succinctness} and {\em stability}.
 
@@
 
+\begin{figure}[!t]
+\includegraphics[width=0.9\linewidth]{_figs/BDBC.eps}
+\caption{Feature Model}\label{fig:bdbc}
+\end{figure}
+
 \input{tex/results_jur.tex}
@@
 \section{Threats to Validity}
+\input{tex/results_models.tex}
 \section{Conclusion}

### 2015-08-20 tex/results_models.tex
Updates from ShareLaTeX

\begin{figure}[htbp]
\begin{center}
\begin{minipage}{\linewidth}
%\noindent\begin{minipage}{0.5\textwidth}
{\scriptsize \textbf{Cost}\\[0.1cm]}
{\scriptsize \begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
  1 &        DTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
\hline  2 &      kNN+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
  2 &          HOW &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
  2 &      kNN &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
\hline \end{tabular}}
\end{minipage}\\[0.1cm]
\begin{minipage}{\linewidth}
{\scriptsize \textbf{Completion}\\[0.1cm]}
{\scriptsize \begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
  1 &      kNN+FS &    0.98  &  0.01 & \quart{10}{12}{22}{22} \\
\hline  2 &          HOW &    1.0  &  0.02 & \quart{32}{22}{44}{22} \\
  2 &      kNN &    1.01  &  0.07 & \quart{0}{76}{54}{22} \\
\hline  3 &        DTREE &    1.04  &  0.03 & \quart{66}{64}{88}{22} \\
\hline \end{tabular}}
\end{minipage}
\end{center}
\caption{Results: }\label{fig:conf}
\end{figure}

### 2015-08-20 Manuscript.tex
Updates from ShareLaTeX

@@
  the expected values of defect counts and  runtimes to    
- 72\%,75\%  (median) and 66\%,20\% (best case) of the initial values, respectively.
+ 45\%,\%  (median) and 66\%,20\% (best case) of the initial values, respectively.
 \end{abstract}
@@
 We propose a novel {\em planning} approach called CROSSTREES for learning changes to a software system
-such that its performance ``improvess'', according to some measure. For  casew
-stuides, we apply CROSSTREES to   (1)~reducethe predicted number
+such that its performance ``improves'', according to some measure. For  case 
+studies, we apply CROSSTREES to   (1)~reduce the predicted number
 of defects and (2)~reduce the runtimes of five system.

### 2015-08-20 Manuscript.tex
Updates from ShareLaTeX

@@
  the expected values of defect counts and  runtimes to    
- 45\%,\%  (median) and 66\%,20\% (best case) of the initial values, respectively.
+ 45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
 \end{abstract}
@@
 Firstly, business users now demanding analytics tools that support a business-level interpretation of their data.
-At a panel on software analytics at ICSE’12, industrial practitioners lamented the state of the art in data mining and software engineering~\cite{menzies12a}.
+At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in data mining and software engineering~\cite{menzies12a}.
 Panelists commented that ``prediction is all well and good, but what about decision making?''.
@@
 
-\section{From Clustering to  CROSSTREES}
+\section{ Methods for Learning Changes}
+ This paper experimentally compares four methods for inputting software project data and outputting
+ recommendations on what to change.
+ For case studies that illustrate how this method works, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10} (shown in \fig{ck} and \fig{jd}) and
+  software system   configuration data from by  Seigmund et al.~\cite{sven12} (see \fig{cpm}).
+
+  The Jurecko data records number of known defects for each class, where the classes are described in terms of
+  nearly two dozen metrics such as number of children, lines of code, etc.
+
+  The Seigmund data records the runtime of compiled systems. To generate that data, Seigmund et al. perturbed
+  the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
+  Berkeley database (one written in ``C'' and one in Java). To ensure that the perturbations
+  were valid, the option space was modeled as a tree that included knowledge of what
+  options were mandatory, optional, or were alternatives (see \fig{bdbc}).
+  Given those perturbations, the systems were then compiled and 
+  Seigmund et al. recorded how long each perturbation took to complete a test suite.
+
+\input{tex/jur_char.tex}
+
+ \begin{figure}
+ \scriptsize
+ \begin{center}
+ \begin{tabular}{r|rr}
+ data set & cases & \% defective\\\hline
+  ant &947& 22\\
+  camel& 1819& 19\\
+ jedit& 1257& 2\\
+ ivy &352& 11\\
+ log4j& 244 &92\\
+ lucebe &442 &59\\
+ poi& 936 &64\\
+ synapse &379 &34\\
+ velocity& 410& 34\\
+ xalan& 2411& 99\\
+ xerces &1055& 74
+ \end{tabular}
+ \end{center}
+ \caption{The Jureczko data sets (columns in the format of \fig{ck}).}\label{fig:jd}
+ \end{figure}
+
+ \begin{figure}[!t]
+\scriptsize
+\begin{tabular}{llllll}
+  \hline
+  \rowcolor{lightgray}
+Project & Domain & Lang. & LOC & Features & Config\\\hline
+Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
+Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
+Apache & Web Server & C & 230,277 & 9 & 192\\
+SQLite & Database & C & 312,625 & 39 & 3,932,160\\
+LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
+x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
+\end{tabular}
+
+\caption{Seigmund data sets: details on the data sets.
+For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
+\end{figure}
+
+\begin{figure}[!t]
+\includegraphics[width=1\linewidth]{_figs/BDBC.eps}
+\caption{Feature model for the Berkeley database (``C'' version). }\label{fig:bdbc}
+\end{figure}
+
+Research has shown that given the above data, it is possible to build
+predictors for defective classes (in the Jureczko data) or for the runtime
+of a compiled system (in the Seigmund data) (e.g. see ~\cite{sven12,nam13,he13,peters13,Menzies2013:local,me09b,me11f,me07b}).
+This paper is about the next step {\em after} prediction. We envisonage as scenario
+where business users are presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. Such users will
+then ask a {\em planning}  question; i.e. ``what can we change to do better than that?'' The rest of this paper
+defines and explores four methods for generating those plans.
+
+\subsection{Method1: Centroid Deltas}
 
@@
 
-We applied planners to reduce (a) the number of defects found in Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}; (b) the runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
-\fig{ck} shows Jureczko data attributes and \fig{cpm} shows the  Seigmund data attributes which were more diverse than the Jurecko data since they contain the results of applying many  mutators to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  C versions),  Apache,  SQLite,  LLVM, and x264.
-
-\input{tex/jur_char.tex}
-\begin{figure}[!t]
-\scriptsize
-\begin{tabular}{llllll}
-  \hline
-  \rowcolor{lightgray}
-Project & Domain & Lang. & LOC & Features & Config\\\hline
-Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
-Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
-Apache & Web Server & C & 230,277 & 9 & 192\\
-SQLite & Database & C & 312,625 & 39 & 3,932,160\\
-LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
-x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
-\end{tabular}
-
-\caption{Seigmund data sets: details on the data sets.
-For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
-\end{figure}
-
-\begin{figure}[!t]
-\includegraphics[width=0.9\linewidth]{_figs/BDBC.eps}
-\caption{Feature Model}\label{fig:bdbc}
-\end{figure}
-
 \input{tex/results_jur.tex}

### 2015-08-22 Manuscript.tex
Updates from ShareLaTeX

@@
 \usepackage{balance}
-\newcommand{\bi}{\begin{itemize}[leftmargin=0.4cm]}
+\newcommand{\bi}{\begin{itemize}}
 \newcommand{\ei}{\end{itemize}}
@@
 \newcommand{\eq}[1]{Equation~\ref{eq:#1}}
-\setlist{nolistsep,leftmargin=5mm}
+%\setlist{nolistsep,leftmargin=5mm}
 %\usepackage[pdftex]{graphicx}
@@
 \alignauthor 
-Rahul~Krishna,~Tim Menzies,~Xipeng~Shen \\
+Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
        \affaddr{Com. Sci., NC State, USA}\\
@@
 Naveen   Lekkalapudi\\
- \affaddr{Software Consultant } \\ 
+ \affaddr{Bellhops } \\ 
        \affaddr{Washington D.C.}\\ 
@@
  analytics, data scientists need to change
- their tools. Instead of merely predicting 
+ their tools. Instead of just predicting 
  some result, they also need tools that generate ``plans'';
- i.e. specific suggestions on  how to change values in order to
- improve on some predicted value.
+ i.e. specific suggestions on  what to change  in order to
+ improve on the predicted values.
 
 This paper is aims to study such planners.
-The planners fall into three broad categories (1) Cluster specific planners (K Nearest Neighbors); (2) Instance specific planners (HOW); (3) Decision Tree based planners.
-These planners are employed as tools for proposing changes to independent
- variables in order to improve on 
- the predictions of the dependent variables. Unlike other approaches
- for learning optimizations to software, these planners do not require
- an underlying model of the domain. Also, these plans
- are designed to result in {\em minimal displacement}
- to current practice.
-
- This paper uses these learners to learn methods
- to reduce defects and decrease program runtime.
- For the test data of this paper, those improvements   reduce
+The planners fall into three broad categories (1) Cluster-specific planners (K Nearest Neighbors); (2) Instance specific planners (HOW); (3) Decision Tree based planners.
+Our experiments show that the latter (decision tree planners)
+works best, measured both in terms of the {\em simplicity} of the plan
+and the {\em size} of the improvements.
+For example, in the   test data of this paper, the  improvements   reduce
  the expected values of defect counts and  runtimes to    
@@
 \item The  defects found  in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
-\item The runtime of the  software    configured by  Seigmund et al.~\cite{sven12}.
+\item The runtime of the  software    configured by  Siegmund et al.~\cite{sven12}.
 \ei 
 Since CROSSTREES is a somewhat novel algorithm, we present it in tutorial form.
-The next section walks the reader through a sequence of approaches that start
+After describing our test data,
+this paper  walks the reader through a sequence of approaches that start
 with a standard clustering algorithm, and progresses on to CROSSTREES. The second
 half of the paper then presents an extensive evaluation of CROSSTREES on 
-the Jureczko  and Seigmund data sets. This is followed by notes on related work
+the Jureczko  and Siegmund data sets. This is followed by notes on related work
 and validity.
 
-\section{ Methods for Learning Changes}
+\input{tex/jur_char.tex}
+
+\section{Test Domains}
  This paper experimentally compares four methods for inputting software project data and outputting
- recommendations on what to change.
- For case studies that illustrate how this method works, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10} (shown in \fig{ck} and \fig{jd}) and
-  software system   configuration data from by  Seigmund et al.~\cite{sven12} (see \fig{cpm}).
+ recommendations on what to change. For case studies that illustrate how this method works, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
+  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
 
@@
   nearly two dozen metrics such as number of children, lines of code, etc.
+For details on the Jurecko data, see  \fig{ck} and \fig{jd}.
 
-  The Seigmund data records the runtime of compiled systems. To generate that data, Seigmund et al. perturbed
+  The Siegmund data records  runtimes of compiled systems. To make that data, Siegmund et al. perturbed
   the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
-  Berkeley database (one written in ``C'' and one in Java). To ensure that the perturbations
+  Berkeley database (one written in ``C'' and one in Java). 
+  Given those perturbations, the systems were then compiled and 
+  Siegmund et al. recorded how long each perturbation took to run a test suite. For details on the Siegmund  data, see \fig{cpm}.
+  To ensure that the perturbations
   were valid, the option space was modeled as a tree that included knowledge of what
   options were mandatory, optional, or were alternatives (see \fig{bdbc}).
-  Given those perturbations, the systems were then compiled and 
-  Seigmund et al. recorded how long each perturbation took to complete a test suite.
-
-\input{tex/jur_char.tex}
 
@@
  \end{center}
- \caption{The Jureczko data sets (columns in the format of \fig{ck}).}\label{fig:jd}
+ \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
  \end{figure}
@@
 
-\caption{Seigmund data sets: details on the data sets.
+\caption{Siegmund data.
 For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
@@
 \includegraphics[width=1\linewidth]{_figs/BDBC.eps}
-\caption{Feature model for the Berkeley database (``C'' version). }\label{fig:bdbc}
+\caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
 \end{figure}
 
+\section{Methods for Learning Changes}
 Research has shown that given the above data, it is possible to build
 predictors for defective classes (in the Jureczko data) or for the runtime
-of a compiled system (in the Seigmund data) (e.g. see ~\cite{sven12,nam13,he13,peters13,Menzies2013:local,me09b,me11f,me07b}).
-This paper is about the next step {\em after} prediction. We envisonage as scenario
-where business users are presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. Such users will
+of a compiled system (in the Siegmund data); e.g. see ~\cite{sven12,nam13,he13,Menzies2013:local,me09b,me11f,me07b}.
+
+This paper is about the next step {\em after} prediction. Suppose
+a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
 then ask a {\em planning}  question; i.e. ``what can we change to do better than that?'' The rest of this paper
-defines and explores four methods for generating those plans.
+defines and evaluates different methods for answering that question.
 
-\subsection{Method1: Centroid Deltas}
+\subsection{Method0: Non-automatic Approaches}
 
+For the purposes of completeness, we offer first some notes
+on how  {\em not} to generate recommendations for changes.
+
+One way to change projects for the better is to ask smart experienced
+people for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
+such advice 
+is an effective strategy and sometimes it is not.
+According to Passos et al.~\cite{passos11},  warn that developers
+may  assume that the lessons they learn from a few past
+projects are general to 
+all their future projects. They comment ``past experiences were taken into account without 
+much consideration for their context~\cite{passos11}.  
+ J{\o}rgensen \& Gruschke~\cite{jorgensen09} concur with Passos et al. They report that 
+  supposed software engineering    ``gurus'' rarely use lessons
+  from past projects to improve their future reasoning and that such poor
+  past advice can be detrimental to new projects.~\cite{jorgensen09}.
+  Based on that research, we proposed a ``trust, but verify'' approach.
+  If a software guru offers some sage wisdom, then it is wise to (a)~follow that
+  advice but (b)~also ask some other oracle if there were any better options
+  (just as a sanity check).
+  The rest of this paper discusses some methods to build automatic oracles 
+  to implement that   sanity check.
+
+Another way find good changes to a project
+would be to rely
+on the peer review processes used by the 
+SE research community. This approach would propose changes to software
+projects that concur with internationally accept best practices. There are three
+problems with that approach. Firstly, given the rapid pace of change in software
+engineering, we may be asking questions for which there is no currently accept
+``best practice''. For example, only very recently has there been any work
+of predicting software runtimes based on choices within Makefiles. While the work
+of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
+cannot yet represent ``internationally accept best practices'' since that work is so
+recent.
+
+Secondly, given the diversity of SE products and practices
+and personnel, it may well be that the current project being discussed is 
+substantively different to prior work. 
+Numerous recent {\em local learning} results compare (1) single models
+learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
+A repeated result in those studies in that the local models generated the better effort
+and defect predictions (better median results,
+lower variance in the predictions).
+
+In summary, 
+it is useful to have automatic methods to recommend changes. Such
+methods can fill in for human gurus (if such gurus are absent) or 
+offer a second opinion that (possibly)  a different perspective to the human guru.
+Also, prior to making automatic recommendations, it is wise to first stratify the data
+(clump it into related examples) then generate advice specific to each clump.
+In the following, we will use various methods that generate numerous clumps as follows:
+  $N$ examples will be divided into groups containing at most $\alpha=\sqrt{N}$
+  examples\footnote{We denote variables
+  set via  engineering judgement  with Greek letters; e.g. $\alpha$.
+  In this paper, we show our current settings to these variables produces useful
+  results.  Future papers should check if different
+  settings to these variables are more useful.}.
+
+ This clumping to gether of different examples  can be operationalized in numerous ways. Consider  a table of data
+with many independent features and one dependent feature (where the latter is usually called
+the ``class''):
+\bi
+\item
+Method1, Method2, and Method3
+use ``unsupervised clustering''; i.e. examples are clumped together if they have
+similar values for their independent features. 
+\item
+Method4 uses ``supervised clustering'';
+i.e. examples are clumped together if that  reduces the dependent feature {\em variability}.
+\ei
+For or continuous and discrete values,
+the {\em variability} can be measured using standard deviation\footnote{$\sigma = \sqrt{\sum^n_i \left( (x_i - \bar{x})^2/(n-1)\right)}$} or entropy\footnote{If $n$ discrete values occur at frequency
+$f_1,f_2,.. f_n$, then \mbox{$N = \sum n_i$} and $p_i = f_i/N$ and entropy is 
+$-1 * \sum_i p_ilog_2(p_i)$.}, respectively).  
+\subsection{Method1: Sorted Centroids}
+
+One method for learning changes is as follows.  
+  Cluster project data by reflecting just on the indepedent variables,
+  Next, build one centroid for each cluster (using the median and mode value for continuous
+and discrete values, respectively).
+After that, write a spreadsheet with one column per centroid;
+\bi
+\item
+Sort the columns such that any current projects client fall into the left-most
+columns. That is, make the left-hand-side of the sheet  ``current usual practice''
+and the right-hand side  ``alternatives to current usual practice''.
+\item
+Sort the rows of that spreadsheet such that the rows with
+maximum  variability appear at the top 
+\ei
+Now show the sheet to the business user, encouraging them to make recommendations by
+\bi
+\item
+Comparing the left and right-sided columns;
+\item
+Using the features mentioned in the top-most rows (where changes selects for the most different centroids).
+\ei
+	Recent results suggest that, we clustering on independent variables of software project data,
+	we are free to select from a wide range of 
+	clustering methods.  Ganesan~\cite{div14} explored 
+	different clustering methods for SE data using   effort and defect data from
+	the PROMISE repository~\cite{promiserepo}.
+	That study explored
+	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
+	below.
+	Clusters were assessed via the performance of prediction 
+	learned from the clusters by Random Forests (for defect prediction)
+	and M5prime or linear regression (for effect estimation).  Ganesan found
+	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
+
+	Accordingly, we select a clustering method that generates a large
+	number of small cluster
+	($\sqrt{N}$ clusters
+	from $N$ examples, that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e.
+contains signals not associated with the target variable).
+For the rest of this article,
+	we use  the top-down
+	bi-clustering method described in \fig{contast_trees} which recursively divides the
+	data in two  using a dimension that captures the greatest variability in the data. 
+	\begin{figure}[t]\small
+	\begin{tabular}{|p{.95\linewidth}|}\hline 
+		WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
+		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F \left(\Delta(X_i,Y_i)^2\right)}/\sqrt{F}$
+			(the denominator normalizes the output to $0 \le d \le 1$).
+			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
+			Otherwise, replace any  missing items with values that maximizes the following.
+			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
+			returns 
+			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
+			same,different (respectively).  
+		Using this measure, WHERE runs as follows:
+		\begin{enumerate}[leftmargin=3mm]
+			\item Find   two   distance cases,  $X,Y$
+			by picking any case $W$ at random, then setting $X$ to its most
+			distant case, then setting $Y$ to the case most distant from
+			$X$
+			(which requires only $O(2N)$ comparisons).
+			\item Project each case $Z$
+			to position $x$ along a    lines that  runs from $X$ to $Y$, If $a,b$ re distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
+			\item Split the data at the median $x$ value of all cases.
+			\item For   splits larger than  $\sqrt{N}$, call this algorithm recursively  from step1.
+		\end{enumerate}\\\hline
+		\end{tabular}
+		\caption{Top-down division of data with WHERE.}\label{fig:where}
+		\end{figure}
+
+In theory, One  advantage of this method is that it focuses the attention of the business users
+away from rows that do not select for different centroids and towards differences between
+current usual practice and everything else.  
 In the summer of 2011 and 2012, one of us (Menzies) spent two months
@@
 
-Turning now to the Seigmund data, the lessons learnt from the planners advise on how  to change the configurations settings within a Makefile (in order to make resulting compiled code  run faster).
+Turning now to the Siegmund data, the lessons learnt from the planners advise on how  to change the configurations settings within a Makefile (in order to make resulting compiled code  run faster).
 Here, performance can be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s=\frac{abs(a - p)}{a}$.
 
-This paper  explores six Seigmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
+This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
   x264\footnote{Available from the performance prediction section of PROMISE
@@
 
-That is, we can be very confident that the predictors from the Seigmund data can assess
-our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Seigmund et al.~\cite{sven12}.)
+That is, we can be very confident that the predictors from the Siegmund data can assess
+our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
 
@@
 \fig{work} showed how we build and test our planners.
-The dependent variables of Jureczko and Seigmund data sets are discrete and continuous in nature, respectively.
-Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Seigmund data.
+The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
+Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
 Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
 \bi
-\item For Seigmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
+\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
 \item For the Jureczko data, we used the training and testing sets of \fig{j}. 
@@
 \section{Conclusion}
+
+works for more than defects and tunrimes
+
 \section*{Acknowledgements}
+The work has partially funded by a National Science Foundation CISE CCF award \#1506586.
 \bibliographystyle{plain}

### 2015-08-22 Manuscript.tex
Updates from ShareLaTeX

@@
 
-\title{Learning Useful Changes\\(to Reduce Runtimes and Software Defects)}
+\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
 \numberofauthors{3} 
@@
  improve on the predicted values.
-
-This paper is aims to study such planners.
-The planners fall into three broad categories (1) Cluster-specific planners (K Nearest Neighbors); (2) Instance specific planners (HOW); (3) Decision Tree based planners.
-Our experiments show that the latter (decision tree planners)
-works best, measured both in terms of the {\em simplicity} of the plan
-and the {\em size} of the improvements.
-For example, in the   test data of this paper, the  improvements   reduce
- the expected values of defect counts and  runtimes to    
+This paper defines an evaluation strategy for evaluating such  planners and applies it to four different planning methods.
+It is shown that our  strategy  can (a) recognize   inferior planning methods and  (b) select useful planning methods.
+Using the best method found by this evaluation strategy,  we can
+reduce
+ the expected values of defect counts and  runtimes of software projects to    
  45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
@@
 such that its performance ``improves'', according to some measure. For  case 
-studies, we apply CROSSTREES to   (1)~reduce the predicted number
-of defects and (2)~reduce the runtimes of five system.
-
+studies, we apply CROSSTREES to  reduce the expected value of:
+\bi
+\item  Defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
+\item Runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.
+\ei 
 The approach discussed here is novel since most data miners used in software analytics
@@
 
-Our approach is important for two reasons.
-Firstly, business users now demanding analytics tools that support a business-level interpretation of their data.
-At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in data mining and software engineering~\cite{menzies12a}.
-Panelists commented that ``prediction is all well and good, but what about decision making?''.
-That is, these panelists are more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
-For example:
+Our approach is important since business users are demanding analytics tools that support  a  business-level interpretation of their data.
+At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
+Panelists commented  ``prediction is all well and good, but what about decision making?''.
+Note that these panelists are more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
+So:
 \bi
@@
 \ei
-A second reason for the importance of our approach is, as shown below,
-when asked how to change a project in order to achieve a result, it is surprisingly
-difficult to use standard methods (e.g. clustering) to find those changes.
-In the case study that motivated this paper, we how  standard data mining methods
-can generate  seemingly
-useful insights, but which are in fact completely ineffectual. On the other hand,
-using CROSSTREES, we can propose changes that significantly reduce:
+This main contribution of this paper is to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
+can (a)~select useful planning methods while (b)~culling clearly
+inferior planning methods. 
+For example,
+this paper compares CROSSTREES against three simpler methods that are found to be inferior.
+Consider  a table of data
+with many independent features and one dependent feature (where the latter is usually called
+the ``class''). All our methods find clumps of related
+examples, then generates specific change recommendations
+for each clumps:
 \bi
-\item The  defects found  in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
-\item The runtime of the  software    configured by  Siegmund et al.~\cite{sven12}.
+\item
+Method1, Method2, and Method3
+use ``unsupervised clustering''; i.e. examples are clumped together if they have
+similar values for their independent features. 
+\item
+CROSSTREES is our implementation of Method4.
+This method uses ``supervised clustering'';
+i.e. examples are clumped together if that  reduces the dependent feature {\em variability}. For or continuous and discrete values,
+the {\em variability} can be measured using standard deviation\footnote{$\sigma = \sqrt{\sum^n_i \left( (x_i - \bar{x})^2/(n-1)\right)}$} or entropy\footnote{If $n$ discrete values occur at frequency
+$f_1,f_2,.. f_n$, then \mbox{$N = \sum f_i$} and $p_i = f_i/N$ and entropy is 
+$- \sum_i p_i{\ldot}\mathit{log}_2(p_i)$.}, respectively.
 \ei
+As seen below, Method1 and Method2 are clearly inferior.
+Method3 and Method4 will be shown to be equivalent, measured in terms of change effectiveness. However, Method4 will be recommended since its change recommendations are
+far simpler.
+
+Of course, there are many more ways than these four
+to generate recommendations on how to change projects.
+But the goal of this paper is not to claim that (e.g.) Method4 is some absolute best algorithm. Rather, it is
+to offer an evaluation strategy that  can assess alternate proposals for generating
+plans. Our hope is that other   researchers repeat our experiments in order to repeat and/or improve
+and/or refute our results. To that end, all our materials and data sets are available on-line.
+
+The rest of this paper is structured as follows.
 Since CROSSTREES is a somewhat novel algorithm, we present it in tutorial form.
 After describing our test data,
-this paper  walks the reader through a sequence of approaches that start
-with a standard clustering algorithm, and progresses on to CROSSTREES. The second
-half of the paper then presents an extensive evaluation of CROSSTREES on 
+this paper  walks the reader through four methods that start
+with a standard clustering algorithm, the  progresses on to CROSSTREES. The second
+half of the paper then presents an extensive evaluation of these methods using 
 the Jureczko  and Siegmund data sets. This is followed by notes on related work
@@
 \section{Test Domains}
- This paper experimentally compares four methods for inputting software project data and outputting
- recommendations on what to change. For case studies that illustrate how this method works, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
+For case studies material, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
   software system   configuration data from by  Siegmund et al.~\cite{sven12}.
@@
 
-  The Siegmund data records  runtimes of compiled systems. To make that data, Siegmund et al. perturbed
+  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To make that data, Siegmund et al. perturbed
   the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
   Berkeley database (one written in ``C'' and one in Java). 
-  Given those perturbations, the systems were then compiled and 
-  Siegmund et al. recorded how long each perturbation took to run a test suite. For details on the Siegmund  data, see \fig{cpm}.
-  To ensure that the perturbations
-  were valid, the option space was modeled as a tree that included knowledge of what
-  options were mandatory, optional, or were alternatives (see \fig{bdbc}).
+   \fig{bdbc} shows an example of a feature model defining valid combinations of settings to on the
+   the Siegmund et al. datasets. These feature models were used by Siegmund et al. to ensure all their perturbations are value
+   (we will use the same models to cull invalid plans).
+  Given those valid perturbations, the systems were then compiled and 
+  Siegmund et al. recorded how long each perturbation took to run a test suite.
 
- \begin{figure}
+ \begin{figure}[!b]
  \scriptsize
@@
 
-\subsection{Method0: Non-automatic Approaches}
-
-For the purposes of completeness, we offer first some notes
-on how  {\em not} to generate recommendations for changes.
-
+Before doing that, we digress to comment on manual methods for generating such plans.
 One way to change projects for the better is to ask smart experienced
@@
 is an effective strategy and sometimes it is not.
-According to Passos et al.~\cite{passos11},  warn that developers
+According to Passos et al.~\cite{passos11},  developers
 may  assume that the lessons they learn from a few past
@@
   past advice can be detrimental to new projects.~\cite{jorgensen09}.
-  Based on that research, we proposed a ``trust, but verify'' approach.
-  If a software guru offers some sage wisdom, then it is wise to (a)~follow that
-  advice but (b)~also ask some other oracle if there were any better options
+
+  Accordingly, we   proposed a ``trust, but verify'' approach.
+  If a software guru offers some sage wisdom,  it is wise to ask some other oracle if there were any better options
   (just as a sanity check).
@@
 methods can fill in for human gurus (if such gurus are absent) or 
-offer a second opinion that (possibly)  a different perspective to the human guru.
+to offer a second opinion.
 Also, prior to making automatic recommendations, it is wise to first stratify the data
@@
   $N$ examples will be divided into groups containing at most $\alpha=\sqrt{N}$
-  examples\footnote{We denote variables
-  set via  engineering judgement  with Greek letters; e.g. $\alpha$.
+  examples.
+
+  In the following, note our convention. All variables
+  set via  our engineering judgement  with Greek letters; e.g. $\alpha$.
   In this paper, we show our current settings to these variables produces useful
-  results.  Future papers should check if different
-  settings to these variables are more useful.}.
+  results. In other work~\cite{krall14}, we are exploring tuning methods to automatically
+  find better settings but  we have nothing definitive yet to report
+  on auto-tuning planners.
 
- This clumping to gether of different examples  can be operationalized in numerous ways. Consider  a table of data
-with many independent features and one dependent feature (where the latter is usually called
-the ``class''):
-\bi
-\item
-Method1, Method2, and Method3
-use ``unsupervised clustering''; i.e. examples are clumped together if they have
-similar values for their independent features. 
-\item
-Method4 uses ``supervised clustering'';
-i.e. examples are clumped together if that  reduces the dependent feature {\em variability}.
-\ei
-For or continuous and discrete values,
-the {\em variability} can be measured using standard deviation\footnote{$\sigma = \sqrt{\sum^n_i \left( (x_i - \bar{x})^2/(n-1)\right)}$} or entropy\footnote{If $n$ discrete values occur at frequency
-$f_1,f_2,.. f_n$, then \mbox{$N = \sum n_i$} and $p_i = f_i/N$ and entropy is 
-$-1 * \sum_i p_ilog_2(p_i)$.}, respectively).  
 \subsection{Method1: Sorted Centroids}
@@
 		WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
-		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F \left(\Delta(X_i,Y_i)^2\right)}/\sqrt{F}$
+		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  \Delta(X_i,Y_i)^2}/\sqrt{F}$
 			(the denominator normalizes the output to $0 \le d \le 1$).

### 2015-08-22 Manuscript.tex
Updates from ShareLaTeX

@@
 to offer an evaluation strategy that  can assess alternate proposals for generating
-plans. Our hope is that other   researchers repeat our experiments in order to repeat and/or improve
-and/or refute our results. To that end, all our materials and data sets are available on-line.
+plans. Our hope is that other   researchers apply this strategy  to repeat and/or improve
+and/or refute our results. To that end, all our materials and data sets are available on-line, see
+github.com/ai-se/XTREES.
 
@@
   Siegmund et al. recorded how long each perturbation took to run a test suite.
+
+  See github.com/ai-se/XTREES\#datafor full access to the Jurecko and Siegmund data.
 

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 \usepackage{cite}
+\usepackage{framed,graphicx,xcolor}
+
 \usepackage[shortlabels]{enumitem} 
@@
 \usepackage{hyperref}
-\usepackage{listings}
+%\usepackage{listings}
 \DeclareMathOperator*{\argmin}{arg\,min} 
@@
 \definecolor{Background}{rgb}{1,1,1}
-\lstnewenvironment{code}[1]{
-\lstset{
-mathescape,
-numbers=left,
-numberstyle=\scriptsize,
-stepnumber=1,
-numbersep=0.5em,
-xleftmargin=1em,
-framextopmargin=2em,
-framexbottommargin=2em,
-showspaces=false,
-showtabs=false,
-showstringspaces=false,
-tabsize=2,
-% Basic
-basicstyle=\ttfamily\scriptsize,
-backgroundcolor=\color{Background},
-language=Python,
-% Comments
-commentstyle=\color{Comments}\slshape,
-% Strings
-stringstyle=\color{Strings},
-morecomment=[s][\color{Strings}]{"""}{"""},
-morecomment=[s][\color{Strings}]{'''}{'''},
-% keywords
-morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
-keywordstyle={[1]\color{Code}\bfseries},
-% additional keywords
-morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, FSel, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
-keywordstyle={[3]\color{Keywords}\bfseries},
-morekeywords={[2]@invari},
-keywordstyle={[2]\color{Decorators}\slshape},
-emph={self},
-emphstyle={\color{self}\slshape},
-firstnumber=last
-%
-}}{}
 
@@
 \begin{document}
+
+\definecolor{shadecolor}{gray}{0.9}
 \conferenceinfo{ISCE}{'16 Austin, Texas}
@@
 \section{Introduction}
-We propose a novel {\em planning} approach called CROSSTREES for learning changes to a software system
+We propose a novel {\em planning} approach called XTREES for learning changes to a software system
 such that its performance ``improves'', according to some measure. For  case 
-studies, we apply CROSSTREES to  reduce the expected value of:
+studies, we apply XTREES to  reduce the expected value of:
 \bi
@@
 inferior planning methods. 
-For example,
-this paper compares CROSSTREES against three simpler methods that are found to be inferior.
-Consider  a table of data
-with many independent features and one dependent feature (where the latter is usually called
-the ``class''). All our methods find clumps of related
-examples, then generates specific change recommendations
-for each clumps:
+We apply this strategy to XTREES and three alternate planning methods to find:
 \bi
-\item
-Method1, Method2, and Method3
-use ``unsupervised clustering''; i.e. examples are clumped together if they have
-similar values for their independent features. 
-\item
-CROSSTREES is our implementation of Method4.
-This method uses ``supervised clustering'';
-i.e. examples are clumped together if that  reduces the dependent feature {\em variability}. For or continuous and discrete values,
-the {\em variability} can be measured using standard deviation\footnote{$\sigma = \sqrt{\sum^n_i \left( (x_i - \bar{x})^2/(n-1)\right)}$} or entropy\footnote{If $n$ discrete values occur at frequency
-$f_1,f_2,.. f_n$, then \mbox{$N = \sum f_i$} and $p_i = f_i/N$ and entropy is 
-$- \sum_i p_i{\ldot}\mathit{log}_2(p_i)$.}, respectively.
+\item Two methods that are clearly inferior;
+\item One method that proposes complex changes, which work as well as XTREES;
+\item And XTREES, that proposes very simple changes.
 \ei
-As seen below, Method1 and Method2 are clearly inferior.
-Method3 and Method4 will be shown to be equivalent, measured in terms of change effectiveness. However, Method4 will be recommended since its change recommendations are
-far simpler.
-
-Of course, there are many more ways than these four
-to generate recommendations on how to change projects.
-But the goal of this paper is not to claim that (e.g.) Method4 is some absolute best algorithm. Rather, it is
-to offer an evaluation strategy that  can assess alternate proposals for generating
-plans. Our hope is that other   researchers apply this strategy  to repeat and/or improve
+Of course, there are many more methods for generating plans and
+no   one paper can survey them all.
+But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
+to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
+Our hope is that other   researchers apply this strategy  to repeat and/or improve
 and/or refute our results. To that end, all our materials and data sets are available on-line, see
@@
 The rest of this paper is structured as follows.
-Since CROSSTREES is a somewhat novel algorithm, we present it in tutorial form.
+Since XTREES is a somewhat novel algorithm, we present it in tutorial form.
 After describing our test data,
 this paper  walks the reader through four methods that start
-with a standard clustering algorithm, the  progresses on to CROSSTREES. The second
+with a standard clustering algorithm, the  progresses on to XTREES. The second
 half of the paper then presents an extensive evaluation of these methods using 
@@
 
+\section{From Prediction to Planning}
+
+This paper is about the next step {\em after} prediction. Suppose
+a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
+then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''
+
+Before exploring automatic methods to answer the planning question, we first comment
+on two manual methods.
+
+One way to propose changes to a project would be to   ask some smart experienced
+person for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
+such advice 
+is an effective strategy and sometimes it is not.
+According to Passos et al.~\cite{passos11},  developers
+may  assume that the lessons they learn from a few past
+projects are general to 
+all their future projects. They comment ``past experiences were taken into account without 
+much consideration for their context~\cite{passos11}.  
+ J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that 
+  supposed software engineering    ``gurus'' rarely use lessons
+  from past projects to improve their future reasoning and that such poor
+  past advice can be detrimental to new projects.~\cite{jorgensen09}.
+  Accordingly, we   propose a ``trust, but verify'' approach.
+  After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
+  if there are any better options
+  (just as a sanity check).
+  The rest of this paper discusses some methods to build automatic oracles 
+  to implement that   sanity check.
+
+Another way  to find   changes to a project
+might be to rely
+on the peer review processes used by the 
+SE research community. This approach would propose changes to software
+projects that concur with internationally accept best practices. 
+There are two
+problems with that approach. Firstly, given the rapid pace of change in software
+engineering, we may be asking questions for which there is no currently accept
+``best practice''. For example, only very recently has there been any work
+of predicting software runtimes based on choices within Makefiles. While the work
+of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
+cannot yet represent ``internationally accept best practices'' since that work is so
+recent.
+
+Secondly, given the diversity of SE products and practices
+and personnel, it may well be that the current project being discussed is 
+substantively different to prior work. 
+Numerous recent {\em local learning} results compare (1) single models
+learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
+A repeated result in those studies in that the local models generated the better effort
+and defect predictions (better median results,
+lower variance in the predictions). This paper offers yet another locality result:
+\bi
+\item
+One standard rule in the literature
+is that it is useful to implement modules such that they are internally cohesive (use
+much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
+\item
+While that may be true in general, for particular classes other changes may be more important
+(later in this paper, we show one set of results were that is indeed the case).
+\ei
+In summary, 
+it is useful to have automatic methods to recommend changes. Such
+methods can fill in for human gurus (if such gurus are absent) or 
+to offer a second opinion.
+Also, prior to making automatic recommendations, it is wise to first stratify the data
+(clump it into related examples) then generate advice specific to each clump.
+
+Accordingly, the rest of this paper defines and evaluates
+automatic methods to find plans from
+  $N$ examples divided  into  clumps containing at most \mbox{ $\alpha=\sqrt{N}$}
+  examples.
+
+\input{tex/prediction_stats.tex}
+
 \section{Test Domains}
-For case studies material, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
+To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
   software system   configuration data from by  Siegmund et al.~\cite{sven12}.
+  See github.com/ai-se/XTREES\#data for full access to this data.
 
@@
 
-  See github.com/ai-se/XTREES\#datafor full access to the Jurecko and Siegmund data.
+  Our evaluation strategy (discussed below) divides this data into a training a test set.
+  From the train set we apply a data miner (to learn a quality predictor) and
+  four planning methods (to learn different plans). Next, using the train set,
+  we apply the plans and ask the quality predictor to assess the results.
 
- \begin{figure}[!b]
- \scriptsize
+  Note that this approach depends on having effective predictors for assessing the results.
+  For the Siegmind data, this criteria was   relatively easy to achieve.
+  The data in those data sets have a continuous class (runtime of the compiled system)
+  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
+This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
+  x264. 
+  As a preliminary study, we split that data   into equal sized train:test groups
+  and trained a Random Forest
+  Regressor\footnote{From the SciKit learn kit~\cite{Pedregosa2012}.}   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
+That is, we can be very confident that the predictors from the Siegmund data can assess
+our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
+
+ It proved to be somewhat more complicated to commission the Jureczko data sets for this study.
+ For the Jureczko OO data sets, we found:
+\bi
+\item
+Most quality predictors built from this data were far from perfect;
+\item 
+But several of these predictors could
+be salvaged using some sampling and optimization methods (see the notes on SMOTE and differential evolution, below).
+\ei
+As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository~\cite{promiserepo}.}.
+Given access to $V$ released
+versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
+shown in \fig{j}, this means that we are training on hundreds to thousands
+of classes and testing on smaller test suites).
+Note the three bottom rows marked with $\times$: these contain predominately
+defective classes (two-thirds, or more) and in such data sets, it is hard to distinguish
+good from bad (since there are so many bad examples).
+
+For the  Jureczko   data,
+the plans generated by our planners advise on  how to change static code attributes in order to reduce defects in Java classes. Here, prediction performance of Random Forest can be measured using:
+\bi
+\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
+the test data detected
+by the {\em predictor}.
+\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
+classes that are {\em predicted} to be defective.
+\ei
+
+The ``untuned'' columns of \fig{j} shows Random Forest's {\em pd,pf}
+values generated after learning from the training data and then being applied to the test data.
+If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
+then only two of our data sets ({\em ivy,ant}) are ``good'' enough.
+
+However, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples.
+Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
+
+In summary, while we cannot trust predictors from some of our defect data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.
+
+ \begin{figure}
+ \small
  \begin{center}
@@
 
- \begin{figure}[!t]
+ \begin{figure}
 \scriptsize
@@
 
-\begin{figure}[!t]
+\begin{figure}
 \includegraphics[width=1\linewidth]{_figs/BDBC.eps}
@@
 
-\section{Methods for Learning Changes}
-Research has shown that given the above data, it is possible to build
-predictors for defective classes (in the Jureczko data) or for the runtime
-of a compiled system (in the Siegmund data); e.g. see ~\cite{sven12,nam13,he13,Menzies2013:local,me09b,me11f,me07b}.
+	\begin{figure} 
+	\begin{shaded}
+  ~\hrule~
 
-This paper is about the next step {\em after} prediction. Suppose
-a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
-then ask a {\em planning}  question; i.e. ``what can we change to do better than that?'' The rest of this paper
-defines and evaluates different methods for answering that question.
+ {\bf Figure 5.A: Measuring Variability}
 
-Before doing that, we digress to comment on manual methods for generating such plans.
-One way to change projects for the better is to ask smart experienced
-people for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
-such advice 
-is an effective strategy and sometimes it is not.
-According to Passos et al.~\cite{passos11},  developers
-may  assume that the lessons they learn from a few past
-projects are general to 
-all their future projects. They comment ``past experiences were taken into account without 
-much consideration for their context~\cite{passos11}.  
- J{\o}rgensen \& Gruschke~\cite{jorgensen09} concur with Passos et al. They report that 
-  supposed software engineering    ``gurus'' rarely use lessons
-  from past projects to improve their future reasoning and that such poor
-  past advice can be detrimental to new projects.~\cite{jorgensen09}.
+	For  continuous and discrete values,
+the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
+Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
+where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
+Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
+for  $n$ discrete values occurring at frequency
+$f_1,f_2,.. f_n$. That is, \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.
 
-  Accordingly, we   proposed a ``trust, but verify'' approach.
-  If a software guru offers some sage wisdom,  it is wise to ask some other oracle if there were any better options
-  (just as a sanity check).
-  The rest of this paper discusses some methods to build automatic oracles 
-  to implement that   sanity check.
+  ~\hrule~
 
-Another way find good changes to a project
-would be to rely
-on the peer review processes used by the 
-SE research community. This approach would propose changes to software
-projects that concur with internationally accept best practices. There are three
-problems with that approach. Firstly, given the rapid pace of change in software
-engineering, we may be asking questions for which there is no currently accept
-``best practice''. For example, only very recently has there been any work
-of predicting software runtimes based on choices within Makefiles. While the work
-of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
-cannot yet represent ``internationally accept best practices'' since that work is so
-recent.
+	{\bf Figure 5.B: Measuring distance}
 
-Secondly, given the diversity of SE products and practices
-and personnel, it may well be that the current project being discussed is 
-substantively different to prior work. 
-Numerous recent {\em local learning} results compare (1) single models
-learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
-A repeated result in those studies in that the local models generated the better effort
-and defect predictions (better median results,
-lower variance in the predictions).
+	We use
+		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  \Delta(X_i,Y_i)^2}/\sqrt{F}$
+			(the denominator normalizes the output to $0 \le d \le 1$).
+			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
+			Otherwise, replace any  missing items with values that maximizes the following.
+			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
+			returns 
+			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
+			same,different (respectively).
 
-In summary, 
-it is useful to have automatic methods to recommend changes. Such
-methods can fill in for human gurus (if such gurus are absent) or 
-to offer a second opinion.
-Also, prior to making automatic recommendations, it is wise to first stratify the data
-(clump it into related examples) then generate advice specific to each clump.
-In the following, we will use various methods that generate numerous clumps as follows:
-  $N$ examples will be divided into groups containing at most $\alpha=\sqrt{N}$
-  examples.
+ ~\hrule~
 
+	{\bf Figure 5.C: Top-down Clustering with WHERE}
+
+				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
+		Using this measure, WHERE runs as follows:
+		\begin{enumerate}[leftmargin=3mm]
+			\item Find   two   distance cases,  $X,Y$
+			by picking any case $W$ at random, then setting $X$ to its most
+			distant case, then setting $Y$ to the case most distant from
+			$X$
+			(which requires only $O(2N)$ comparisons).
+			\item Project each case $Z$
+			to position $x$ along a    lines that  runs from $X$ to $Y$, If $a,b$ re distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
+			\item Split the data at the median $x$ value of all cases.
+			\item For   splits larger than  $\alpha=\sqrt{N}$, call this algorithm recursively  from step1.
+		\end{enumerate}
+ In terms of related work,
+	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
+	  at each recursive level to find the PCA principle component. Our method, on the other hand,
+	  performs the same task with only $O(2N)$ distance calculations 	using the 
+	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.
+
+	   ~\hrule~
+
+	{\bf Figure 5.D: Top-down division with Decision Trees}
+
+	Find a split in the values of the independent features that most reduces the variability
+  of the dependent feature\footnote{}. The data is then divided on that feature's split and 
+  the tree learner recurses on each split (stopping when any split is smaller than $\alpha=\sqrt{N}$).
+
+	 ~\hrule~
+	  \end{shaded}
+		\caption{Some algorithms used in this paper.}\label{fig:where}
+		\end{figure}
+
+\section{Four Planning Methods}
+In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
+algorithm should be baselined against simpler alternatives. Accordingly,
+this section described XTREES (which we call Method4) against three  
+alternatives: 
+  \bi
+  \item  
+  Method1, Method2 and Method3 use the   top-down
+	bi-clustering method described in \fig{where}.C  which recursively divides the
+	data in two  using a dimension that captures the greatest variability in the data. 
+	\item
+ Method4 uses the decision tree learner of \fig{where}.D. 
+  \ei
 In the following, note our convention. All variables
-  set via  our engineering judgement  with Greek letters; e.g. $\alpha$.
+  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
   In this paper, we show our current settings to these variables produces useful
@@
 
-\subsection{Method1: Sorted Centroids}
+\subsection{Method1 = CD =   Centroid Deltas}
 
-One method for learning changes is as follows.  
-  Cluster project data by reflecting just on the indepedent variables,
+ {\em Summary:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and
+where you want to be  (which we will call $C_j$).
+
+ {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
+  reporting the delta between the cluster centroids. The advantage of Method1 is its
+  simplicity. However, as shown by our evaluation strategy, it is not an useful
+  method for generating effective plans.
+
+After clustering some training data using the WHERE algorithm of  \fig{where}.C, this method:
+\bi
+\item
+ Replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
+continuous/discrete feature.
+\item Finds the closest centroid $C_j$ that has a better
+performance score. For defect data, ``better'' means fewer defects while for the config data,
+``better'' means faster runtimes.
+\item Caches the  delta in the independent features between $C_i$ and $C_j$. For continuous
+features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
+in $C_j$. 
+\ei
+Next, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
+the nearest centroid $C_i$.  It then proposes a plan for improving that test case
+that is the conjunction of all the deltas between $C_i$ and $C_j$.
+
+\subsection{Method2 = CD+FS =  Method1  + \\Feature Selection }
+
+{\em Summary:} Method2 works like Method1, but this time, it only mentions the most
+informative features. Hence, Method2's plans are simpler.
+
+{\em Details:} In most SE applications, not all features contribute equally to a problem.  Hence, one problem
+with the above method is that its plans comment on all features, even if those values are due
+to noise or other effects not connected to better performance.
+
+This is a standard issue in data mining, with a standard solution.
+A common result is that the signal in a table of data is mostly contained in a handful of features~\cite{hall03,kohavi97}.
+Papakroni~\cite{papa13} has tested for this effect in the Jureczko data sets.
+Papakroni found no loss of   efficacy in defect prediction after
+sorting all features by their information content,
+then making predictions using (a)~all  features or (b)~just using   33\% most informative features.
+
+Based on the above, it might be possible to simplify Method1's plans by pruning back the features used to make
+the plans. Following on from Papakroni, our Method2 returns plans
+containing just the top $\beta=33\%$ most informative features. Here, ``informative'' means
+that the values of a feature are good for selecting a small set of clusters (ideally,
+just one).
+This can be estimated using the Fayyad-Iranni INFOGAIN algorithm~\cite{FayIra93Multi}
+as follows:
+\bi
+\item Ran WHERE to generate   clusters;
+\item Discretize numeric features using the Fayyad-Iranni discretizer
+(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids);
+\item Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows.
+\item Let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
+\item Cull the features as per Papakroni; i.e. just use the $\beta=33\%$ most informative features
+where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).
+\ei
+After that, Method2 works line Method1 but now the   plans  only
+mention the $\beta=33\%$ most informative features.
+\subsection{Method3 = BIC = Method2 + Best-in-cluster}
+
+{\em Summary:} Method3 is like Method2, but it uses more knowledge about the training data.
+
+{\em Details:}
+One problem with Method2 is that it makes little use of the detailed knowledge
+of the examples that are squashed together to form the centroids. Accordingly,
+our Method3 works like Method2 but some important differences:
+\bi
+\item Clusters are summarized in {\em two} examples: (1)~the centroid found in Method1 and
+(2)~the best-in-cluster example; i.e. the  example in that cluster
+with the best performance score.
+\item As before, each centroid $C_i$ is annotated with the nearest centroid $C_j$
+with a better performance score.
+\item When generating a plan for a test instance, Method3:
+\bi
+\item
+Find the test cases's nearest centroid $C_i$;
+\item
+Finds the associated better centroid $C_j$ and its associated best-in-cluster example.
+\item
+Computes a plan to change the test case  from the delta between the test case
+and the best in that $C_j$ cluster.
+\ei
+\ei
+Note that, as with Method2, this method's plans only mentions the  $\beta=33\%$ most informative features. 
+%\input{tex/pseudo-code-1.tex}
+
+\subsection{Method4 = XTREES = Deltas in Decision Tree Branches}
+
+{\em Summary:} Method4 builds a decision tree,  then generates
+plans from the difference between two branches:
+(1)~the branch to where you are and (2)~the branch to where you want to be.
+
+{\em Details:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
+the algorithm used for the initial clustering. Unsupervised methods like WHERE
+execute without knowledge of the target class.
+{\em Supervised} methods, on the other hand, are allowed to reflect on more knowledge including
+the values in the target class. In theory, this extra knowledge can lead to better data mining.
+Accordingly, Method4 uses a supervised method to generate its plans.
+
+In summary, XTREES uses the decision tree algorithm of \fig{where}.D to divide the data.
+Decision trees have branches down to leaves. 
+XTREES builds plans by asking three questions:
+\be
+\item
+What {\em current} branch does a test case fall in?
+\item What {\em desired} branch would the test case want to move to?
+\item What are the {\em deltas} between current and desired? 
+\ee
+The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.
+
+\input{tex/dtree.tex}
+
+\input{tex/results_models.tex}
+
+\section{Experiments}
+
+\begin{figure}[!t]
+\small 
+\[
+\begin{array}{r}
+\mathrm{project}\\
+\mathrm{data}
+\end{array} 
+\left\{\begin{array}{l}\mathit{train}
+        \left\{\begin{array}{l}
+                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
+                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
+              \end{array}\right.
+       \\
+      ~\\
+\mathit{test}  
+    \left\{\begin{array}{l@{~}l}
+           \mathit{actual}& =\mathrm{Defects \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
+           \mathrm{\bf if\;}\mathit{actual} & >  \mathit{0}\\
+           \mathrm{\bf then} &
+           \left\{
+            \begin{array}{l}
+                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
+                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
+                \mathrm{{\bf return}\;} \mathit{actual},\mathit{after}
+            \end{array}
+          \right.
+   \end{array}\right.
+\end{array} \right. 
+\]
+
+\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
+\end{figure}
+
+\subsection{Setup}
+
+In the experiments shown below,  we use the planners to comment on how to reduce
+software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
+tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}.
+
+\fig{work} shows our proposal to verify if the recommendations made by our planners actually work. Accordingly, we divide the
+project data  into two disjoint sets {\em train} and {\em test}
+(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
+Next, from the train set, we build both a {\em planner} as well
+as a {\em  predictor}.
+Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper, our {\em planner} will be from \ref{clust_contrast} and our {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} and Random Forest Regressor taken from the SciKit Learn toolkit~\cite{Pedregosa2012}.
+
+As for the {\em test} data, this is passed to the { predictor}
+to measure the baseline prediction accuracy, false alarm, and recall.
+If our { predictors} fail to perform adequately on the test data,
+then we cannot trust them to comment on our plans. Accordingly,
+if that performance is unsatisfactory, we abort.
+Else, we (1)~apply the { planner} to alter the {\em test} data;
+then (2)~apply the { predictor} to the altered data $test'$;
+then (3)~return data on the {\em before, after} predictions.
+
+\subsection{Data}
+
+\input{tex/results_jur.tex}
+\input{tex/results_conf.tex}
+
+\begin{figure}[t]
+\begin{minipage}{0.5\textwidth}
+\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
+\label{fig:delta_ivy}
+\end{minipage}
+\begin{minipage}{0.5\textwidth}
+\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
+\label{fig:delta_apache}
+\end{minipage}
+\caption{Performance Comparison showing the number of times a certain feature requires to be changed.
+X-axis lists the various features that make up a data set.
+The y-axis contains the frequency of change in \%.}
+\end{figure}
+
+The clear advantage of our current work is that it can be readily and automatically applied to domains with historical data.
+However, the obvious drawback with this approach is that our assessment of the plans is only as good as the predictor.
+To mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for {\em satisfactory} thresholds.
+
+\subsection{Assessment}
+\fig{work} showed how we build and test our planners.
+The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
+Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
+Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
+\bi
+\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
+\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
+\ei
+The 40 repeats defined above were repeated for each planner.
+A Scott-Knott procedure was used to test for statistical differences in the result.
+The Scott-Knott procedure was recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{mittas13}.
+
+This method
+sorts a list of $l$ treatments with $ls$ measurements by their median
+score. It then
+splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
+ differences  in the observed performances
+before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
+ $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
+Scott-Knott  applies an  hypothesis test $H$ to check
+if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.
+
+The hypothesis test $H$ used here is a conjunction of the A12 effect size test and a non-parametric bootstrap sampling.
+In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
+For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
+These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
+
+\subsection{Results}
+\input{tex/deltas_avg.tex}
+The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
+\subsection{Discussion}	whye where?
+Recent results suggest that, when clustering on independent variables of software project data,
+	we are free to select from a wide range of 
+	clustering methods.  Ganesan~\cite{div14} explored 
+	different clustering methods for SE data using   effort and defect data from
+	the PROMISE repository~\cite{promiserepo}.
+	That study explored
+	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
+	below.
+	Clusters were assessed via the performance of prediction 
+	learned from the clusters by Random Forests (for defect prediction)
+	and M5prime or linear regression (for effect estimation).  Ganesan found
+	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
+	Accordingly, for clustering, we select a clustering method that generates a large
+	number of small cluster
+	($\sqrt{N}$ clusters
+	from $N$ examples) that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e.
+contains signals not associated with the target variable.
+
+\section{Threats to Validity}
+\section{Conclusion}
+
+works for more than defects and tunrimes
+
+\section*{Acknowledgements}
+The work has partially funded by a National Science Foundation CISE CCF award \#1506586.
+\bibliographystyle{plain}
+
+\bibliography{References}
+\end{document}
+
   Next, build one centroid for each cluster (using the median and mode value for continuous
@@
 \ei
-	Recent results suggest that, we clustering on independent variables of software project data,
-	we are free to select from a wide range of 
-	clustering methods.  Ganesan~\cite{div14} explored 
-	different clustering methods for SE data using   effort and defect data from
-	the PROMISE repository~\cite{promiserepo}.
-	That study explored
-	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
-	below.
-	Clusters were assessed via the performance of prediction 
-	learned from the clusters by Random Forests (for defect prediction)
-	and M5prime or linear regression (for effect estimation).  Ganesan found
-	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
-
-	Accordingly, we select a clustering method that generates a large
-	number of small cluster
-	($\sqrt{N}$ clusters
-	from $N$ examples, that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e.
-contains signals not associated with the target variable).
-For the rest of this article,
-	we use  the top-down
-	bi-clustering method described in \fig{contast_trees} which recursively divides the
-	data in two  using a dimension that captures the greatest variability in the data. 
-	\begin{figure}[t]\small
-	\begin{tabular}{|p{.95\linewidth}|}\hline 
-		WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
-		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  \Delta(X_i,Y_i)^2}/\sqrt{F}$
-			(the denominator normalizes the output to $0 \le d \le 1$).
-			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
-			Otherwise, replace any  missing items with values that maximizes the following.
-			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
-			returns 
-			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
-			same,different (respectively).  
-		Using this measure, WHERE runs as follows:
-		\begin{enumerate}[leftmargin=3mm]
-			\item Find   two   distance cases,  $X,Y$
-			by picking any case $W$ at random, then setting $X$ to its most
-			distant case, then setting $Y$ to the case most distant from
-			$X$
-			(which requires only $O(2N)$ comparisons).
-			\item Project each case $Z$
-			to position $x$ along a    lines that  runs from $X$ to $Y$, If $a,b$ re distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
-			\item Split the data at the median $x$ value of all cases.
-			\item For   splits larger than  $\sqrt{N}$, call this algorithm recursively  from step1.
-		\end{enumerate}\\\hline
-		\end{tabular}
-		\caption{Top-down division of data with WHERE.}\label{fig:where}
-		\end{figure}
 
@@
 
-We divide the $N$ training samples into $\sqrt{N}$ clusters and compute the centroid of these clusters.
-For every test case, we identify a cluster from the training set that most closely resembles it.
-Following this, we find the nearest cluster with a better performance score.
-The differences in the attributes between these two clusters constitute the \textit{"contrast set"}.
-These contrast sets acts as plans that can be used to reflect over the test cases to improve them.
-
-In most SE applications, not all features contribute equally to a problem.
-With this in mind, we opine that it would be beneficial if the above method is extended to include some form of feature weighting, thus enabling the tools to recommend changes to only the most informative features.
-This paper uses a feature weighting scheme similar to that used in C4.5, see \fig{where}d.
-
-Simple though this method may seem, using nearest neighbors presents a couple of significant challenges: 
-\begin{itemize}
-\item[1.] Displacements are seldom limited to small regions, this implies that large changes may be recommended when they are not necessary. This may make the solutions less feasible to implement. 
-\item[2.] The nearness of a test case to a cluster in the training data is subjective.
-This affects the locality of the solutions, thus jeopardizing the relevance of these changes to the test case under consideration.
-\end{itemize}
-
-In view of the above issues, we developed HOW, a tool that combines (a) Feature Selection; (b) Centroid generation from clusters; (c) Contrast generation using gradient between clusters.
-\input{tex/pseudo-code-1.tex}
-
 \subsection{HOW}
@@
 
-\input{tex/dtree.tex}
 \subsection{DECISION TREES}
@@
 
-\section{Experiments}
-
-\begin{figure}[!t]
-\small 
-\[
-\begin{array}{r}
-\mathrm{project}\\
-\mathrm{data}
-\end{array} 
-\left\{\begin{array}{l}\mathit{train}
-        \left\{\begin{array}{l}
-                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
-                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
-              \end{array}\right.
-       \\
-      ~\\
-\mathit{test}  
-    \left\{\begin{array}{l@{~}l}
-           \mathit{actual}& =\mathrm{Defects \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
-           \mathrm{\bf if\;}\mathit{actual} & >  \mathit{0}\\
-           \mathrm{\bf then} &
-           \left\{
-            \begin{array}{l}
-                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
-                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
-                \mathrm{{\bf return}\;} \mathit{actual},\mathit{after}
-            \end{array}
-          \right.
-   \end{array}\right.
-\end{array} \right. 
-\]
-
-\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
-\end{figure}
-
-\subsection{Setup}
-
-In the experiments shown below,  we use the planners to comment on how to reduce
-software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
-tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}.
-
-\fig{work} shows our proposal to verify if the recommendations made by our planners actually work. Accordingly, we divide the
-project data  into two disjoint sets {\em train} and {\em test}
-(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
-Next, from the train set, we build both a {\em planner} as well
-as a {\em  predictor}.
-Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper, our {\em planner} will be from \ref{clust_contrast} and our {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} and Random Forest Regressor taken from the SciKit Learn toolkit~\cite{Pedregosa2012}.
-
-As for the {\em test} data, this is passed to the { predictor}
-to measure the baseline prediction accuracy, false alarm, and recall.
-If our { predictors} fail to perform adequately on the test data,
-then we cannot trust them to comment on our plans. Accordingly,
-if that performance is unsatisfactory, we abort.
-Else, we (1)~apply the { planner} to alter the {\em test} data;
-then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions.
-
-\subsection{Data}
-
-\input{tex/results_jur.tex}
-\input{tex/results_conf.tex}
-\input{tex/prediction_stats.tex}
-
-\begin{figure}[t]
-\begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
-\label{fig:delta_ivy}
-\end{minipage}
-\begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
-\label{fig:delta_apache}
-\end{minipage}
-\caption{Performance Comparison showing the number of times a certain feature requires to be changed.
-X-axis lists the various features that make up a data set.
-The y-axis contains the frequency of change in \%.}
-\end{figure}
-
-The clear advantage of our current work is that it can be readily and automatically applied to domains with historical data.
-However, the obvious drawback with this approach is that our assessment of the plans is only as good as the predictor.
-To mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for {\em satisfactory} thresholds.
-
-For the Jureczko OO data sets, we found:
-\bi
-\item
-Most predictors in this domain were far from perfect;
-\item 
-But several of these predictors could
-be salvaged using some sampling and optimization methods (see the notes on SMOTE and differential evolution, below).
-\ei
-As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository~\cite{promiserepo}.}.
-Given access to $V$ released
-versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
-shown in \fig{j}, this means that we are training on hundreds to thousands
-of classes and testing on smaller test suites).
-Note the three bottom rows marked with $\times$: these contain predominately
-defective classes (two-thirds, or more) and in such data sets, it is hard to distinguish
-good from bad (since there are so many bad examples).
-
-For the  Jureczko   data,
-the plans generated by our planners advise on  how to change static code attributes in order to reduce defects in Java classes. Here, prediction performance of Random Forest can be measured using:
-\bi
-\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
-the test data detected
-by the {\em predictor}.
-\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
-classes that are {\em predicted} to be defective.
-\ei
-
-The ``untuned'' columns of \fig{j} shows Random Forest's {\em pd,pf}
-values generated after learning from the training data and then being applied to the test data.
-If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
-then only two of our data sets ({\em ivy,ant}) are ``good'' enough.
-
-However, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
-Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples.
-Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
-The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
-For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
-However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
-
-Turning now to the Siegmund data, the lessons learnt from the planners advise on how  to change the configurations settings within a Makefile (in order to make resulting compiled code  run faster).
-Here, performance can be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s=\frac{abs(a - p)}{a}$.
-
-This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
-  x264\footnote{Available from the performance prediction section of PROMISE
-  http://openscience.us/repo/performance-predict.}. 
-  After splitting that data into equal sized train:test groups, a Random Forest
-  Regressor trained on one half, then applied to the other, achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\]
-
-That is, we can be very confident that the predictors from the Siegmund data can assess
-our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
-
-In summary, while we cannot trust predictors from some of our defect data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.
-
-\subsection{Assessment}
-\fig{work} showed how we build and test our planners.
-The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
-Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
-Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
-\bi
-\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
-\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
-\ei
-The 40 repeats defined above were repeated for each planner.
-A Scott-Knott procedure was used to test for statistical differences in the result.
-The Scott-Knott procedure was recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{mittas13}.
-
-This method
-sorts a list of $l$ treatments with $ls$ measurements by their median
-score. It then
-splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
- differences  in the observed performances
-before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
- $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
-Scott-Knott  applies an  hypothesis test $H$ to check
-if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.
-
-The hypothesis test $H$ used here is a conjunction of the A12 effect size test and a non-parametric bootstrap sampling.
-In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
-For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
-For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
-These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
-Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
-
-\subsection{Results}
-\input{tex/deltas_avg.tex}
-The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
-\subsection{Discussion}
-\section{Threats to Validity}
-\input{tex/results_models.tex}
-\section{Conclusion}
-
-works for more than defects and tunrimes
-
-\section*{Acknowledgements}
-The work has partially funded by a National Science Foundation CISE CCF award \#1506586.
-\bibliographystyle{plain}
-
-\bibliography{References}
-\end{document}
+\lstnewenvironment{code}[1]{
+\lstset{
+mathescape,
+numbers=left,
+numberstyle=\scriptsize,
+stepnumber=1,
+numbersep=0.5em,
+xleftmargin=1em,
+framextopmargin=2em,
+framexbottommargin=2em,
+showspaces=false,
+showtabs=false,
+showstringspaces=false,
+tabsize=2,
+% Basic
+basicstyle=\ttfamily\scriptsize,
+backgroundcolor=\color{Background},
+language=Python,
+% Comments
+commentstyle=\color{Comments}\slshape,
+% Strings
+stringstyle=\color{Strings},
+morecomment=[s][\color{Strings}]{"""}{"""},
+morecomment=[s][\color{Strings}]{'''}{'''},
+% keywords
+morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
+keywordstyle={[1]\color{Code}\bfseries},
+% additional keywords
+morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, FSel, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
+keywordstyle={[3]\color{Keywords}\bfseries},
+morekeywords={[2]@invari},
+keywordstyle={[2]\color{Decorators}\slshape},
+emph={self},
+emphstyle={\color{self}\slshape},
+firstnumber=last
+%
+}}{}

### 2015-08-23 tex/dtree.tex
Updates from ShareLaTeX

@@
-\begin{figure}[t]
-	\small
+\begin{figure}[!t] 
+	\begin{shaded}
   ~\hrule~
 
-	{\bf Step 1: Distinguish between clusters using  decision trees}
+Using the training data,  divide the data using the decision tree of algorithm of \fig{where}.D.
 
-	Use an entropy-based
-	decision tree (DT) learner to learn what attributes select for a given data. To limit tree size:
+For each item in the test data,
+	  find the {\em current } leaf: take each test instance, run it down to a leaf in the decision tree.  
+After that,	  find the {\em desired} leaf:
+		\begin{itemize}[leftmargin=3mm]
+		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
+		\item Identify {\em sibling} leaves; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
+		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\gamma=0.5$ times the mean score of {\em current}.
 		 \bi
-	\item Only use the top $\alpha=33$\%  of the features, as determined by their information gain~\cite{Irani1993}. 
-	\item Only build the trees down to  max depth of $\beta=10$.
-	\item Only build subtrees if it contains at least $N^{\gamma=0.5}$ examples (where $N$ is the size of the training set).
+		 \item 
+		   If none found, then repeat for $lvl += 1$
+		 \item
+		    Return no plan if the new $lvl$ is above the root.
 		 \ei
-	{\em Score}  DT  leaf nodes  via the mean score of its majority cluster.
-
-	~\hrule~
-
-	{\bf Step 2: Generating contrast sets from DT branches}
-	\begin{itemize}
-		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
-		\item Find the {\em desired} cluster: 
-		\bi
-		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
-		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
-		\item Using the {\em score} defined above, find the {\em better} siblings; i.e.
-those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}.
-If none found, then repeat for $lvl += 1$
 		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
 		\ei
-		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
-		\bi
+	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta:
+		\begin{itemize}[leftmargin=3mm]
 		\item
@@
 		\item
-		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
+		For numerics  discretized into ranges, return a random number selected from the low and high boundaries of the that range.
 		\ei 
-		\ei
+		Finally, return the delta as the plan for improving the test instance.
 		~\hrule~
-		\caption{CROSSTREES. Controlled by the parameters
-			$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
+		\end{shaded}
+		\caption{XTREES}
 		\label{fig:xtrees_bare}

### 2015-08-23 tex/results_models.tex
Updates from ShareLaTeX

@@
 \begin{figure}[htbp]
-\begin{center}
-\begin{minipage}{\linewidth}
+
+~\hrule~
+
 %\noindent\begin{minipage}{0.5\textwidth} 
-{\scriptsize \textbf{Cost}\\[0.1cm]}
-{\scriptsize \begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
+{\small 
+\begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
 \arrayrulecolor{lightgray}
-\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
-  1 &        DTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
-\hline  2 &      kNN+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
-  2 &          HOW &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
-  2 &      kNN &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
-\hline \end{tabular}}
-\end{minipage}\\[0.1cm]
-\begin{minipage}{\linewidth}
-{\scriptsize \textbf{Completion}\\[0.1cm]}
-{\scriptsize \begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
-\arrayrulecolor{lightgray}
-\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
-  1 &      kNN+FS &    0.98  &  0.01 & \quart{10}{12}{22}{22} \\
-\hline  2 &          HOW &    1.0  &  0.02 & \quart{32}{22}{44}{22} \\
-  2 &      kNN &    1.01  &  0.07 & \quart{0}{76}{54}{22} \\
-\hline  3 &        DTREE &    1.04  &  0.03 & \quart{66}{64}{88}{22} \\
-\hline \end{tabular}}
-\end{minipage}
-\end{center}
-\caption{Results: }\label{fig:conf}
+\textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
+  1 &        XTREES &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
+\hline  2 &      CD+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
+  2 &          BIC &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
+  2 &      CD &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
+\hline 
+\end{tabular}}
+
+~\\
+
+In this figure, {\bf 50th}
+is the median value and  {\bf 25-75th} is the second and third quartile range.
+The right-hand-side horizontal lines show the {\bf 50th} median value (as a black dot) in the middle
+of the {\bf 25-75th} percentile range (as lines). All values here are expressed as ratios
+of the median values seen in the untreated data, so {\em smaller} values are {\em better}).
+A median {\bf 50th} value of
+1.0 means ``no change from
+baseline''; numbers less than 1.0 indicate some reduction to the baseline;
+and numbers more than 1.0 indicate ``optimization failure''.
+The {\bf Rank} column on the left-hand-side show the results of a statistical
+analysis and results ranked ``1'' are best. Results change {\bf Rank} if they are significantly different and pass
+ a post-hoc effect size test.
+
+ ~\hrule~
+\caption{Results: Methods 1,2,3,4 applied to some ground-truth data.
+Values collected from   20 repeated runs of each method with different random seeds.
+}\label{fig:conf}
 \end{figure}

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 \usepackage{framed,graphicx,xcolor}
+
+\usepackage{stfloats}
 
@@
 
+ \definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
 \newcommand{\G}{\cellcolor{green}}
@@
 (clump it into related examples) then generate advice specific to each clump.
-
 Accordingly, the rest of this paper defines and evaluates
 automatic methods to find plans from
-  $N$ examples divided  into  clumps containing at most \mbox{ $\alpha=\sqrt{N}$}
-  examples.
-
-\input{tex/prediction_stats.tex}
-
-\section{Test Domains}
-To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
-  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
-  See github.com/ai-se/XTREES\#data for full access to this data.
-
-   The Jurecko data records number of known defects for each class, where the classes are described in terms of
-  nearly two dozen metrics such as number of children, lines of code, etc.
-For details on the Jurecko data, see  \fig{ck} and \fig{jd}.
-
-  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To make that data, Siegmund et al. perturbed
-  the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
-  Berkeley database (one written in ``C'' and one in Java). 
-   \fig{bdbc} shows an example of a feature model defining valid combinations of settings to on the
-   the Siegmund et al. datasets. These feature models were used by Siegmund et al. to ensure all their perturbations are value
-   (we will use the same models to cull invalid plans).
-  Given those valid perturbations, the systems were then compiled and 
-  Siegmund et al. recorded how long each perturbation took to run a test suite.
-
-  Our evaluation strategy (discussed below) divides this data into a training a test set.
-  From the train set we apply a data miner (to learn a quality predictor) and
-  four planning methods (to learn different plans). Next, using the train set,
-  we apply the plans and ask the quality predictor to assess the results.
-
-  Note that this approach depends on having effective predictors for assessing the results.
-  For the Siegmind data, this criteria was   relatively easy to achieve.
-  The data in those data sets have a continuous class (runtime of the compiled system)
-  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
-This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
-  x264. 
-  As a preliminary study, we split that data   into equal sized train:test groups
-  and trained a Random Forest
-  Regressor\footnote{From the SciKit learn kit~\cite{Pedregosa2012}.}   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
-That is, we can be very confident that the predictors from the Siegmund data can assess
-our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
-
- It proved to be somewhat more complicated to commission the Jureczko data sets for this study.
- For the Jureczko OO data sets, we found:
-\bi
-\item
-Most quality predictors built from this data were far from perfect;
-\item 
-But several of these predictors could
-be salvaged using some sampling and optimization methods (see the notes on SMOTE and differential evolution, below).
-\ei
-As shown in \fig{j}, we   explore numerous examples of the Jureczko   object-oriented static code data sets: Ant, Camel, Ivy, Jedit,   Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository~\cite{promiserepo}.}.
-Given access to $V$ released
-versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
-shown in \fig{j}, this means that we are training on hundreds to thousands
-of classes and testing on smaller test suites).
-Note the three bottom rows marked with $\times$: these contain predominately
-defective classes (two-thirds, or more) and in such data sets, it is hard to distinguish
-good from bad (since there are so many bad examples).
-
-For the  Jureczko   data,
-the plans generated by our planners advise on  how to change static code attributes in order to reduce defects in Java classes. Here, prediction performance of Random Forest can be measured using:
-\bi
-\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
-the test data detected
-by the {\em predictor}.
-\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
-classes that are {\em predicted} to be defective.
-\ei
-
-The ``untuned'' columns of \fig{j} shows Random Forest's {\em pd,pf}
-values generated after learning from the training data and then being applied to the test data.
-If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
-then only two of our data sets ({\em ivy,ant}) are ``good'' enough.
-
-However, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
-Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples.
-Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
-The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
-For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
-However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
-
-In summary, while we cannot trust predictors from some of our defect data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.
+  $N$ examples divided  into  many clumps.
 
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \includegraphics[width=1\linewidth]{_figs/BDBC.eps}
@@
 
-	\begin{figure} 
+\input{tex/prediction_stats.tex}
+
+\section{Test Domains}
+To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
+  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
+  See github.com/ai-se/XTREES\#data for full access to this data.
+
+   The Jurecko data records number of known defects for each class, where the classes are described in terms of
+  nearly two dozen metrics such as number of children, lines of code, etc.
+For details on the Jurecko data, see  \fig{ck} and \fig{jd}.
+
+  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To make that data, Siegmund et al. perturbed
+  the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
+  Berkeley database (one written in ``C'' and one in Java). 
+   \fig{bdbc} shows an example of a feature model defining valid combinations of settings to on the
+   the Siegmund et al. datasets. These feature models were used by Siegmund et al. to ensure all their perturbations are value
+   (we will use the same models to cull invalid plans).
+  Given those valid perturbations, the systems were then compiled and 
+  Siegmund et al. recorded how long each perturbation took to run a test suite.
+
+  Our evaluation strategy (discussed below) divides this data into a training a test set.
+  From the train set we apply a data miner (to learn a quality predictor) and
+  four planning methods (to learn different plans). Next, we apply
+  those plans to the  test set and ask the quality predictor to assess the results.
+
+  Note that this approach depends on having effective predictors for assessing the results.
+  For the Siegmind data, this criteria was   relatively easy to achieve.
+  The data in those data sets have a continuous class (runtime of the compiled system)
+  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
+This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
+  x264. 
+  As a preliminary study, we split that data   into equal sized train:test groups
+  and trained a Random Forest
+  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
+That is, we can be very confident that the predictors from the Siegmund data can assess
+our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
+
+ It proved to be  more complicated to commission the Jureczko data sets for this study.
+ For that data, we found:
+\bi
+\item
+ Quality predictors built from this data are far from perfect;
+\item 
+But several of these predictors could
+be salvaged using the techniques discussed in this section.
+\ei
+ \fig{j} shows our preliminary studies with the Jureczko   data.
+Given access to $V$ released
+versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
+shown in \fig{j}, this means that we are training on hundreds to thousands
+of classes and testing on smaller test suites).
+Note the three bottom \colorbox{lavenderpink}{highlighted} rows marked with $\times$: these contain predominately
+defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
+good from bad (since there are so many bad examples).
+
+The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
+For such data, quality predictor   is be measured using:
+\bi
+\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
+the test data detected
+by the {\em predictor}.
+\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
+classes that are {\em predicted} to be defective.
+\ei 
+As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
+Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
+built from the training data, then applied to the test data.
+The ``untuned'' columns of \fig{j} shows those results.
+If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
+then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
+
+Fortunately,
+the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples.
+Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
+
+In summary, while we cannot trust predictors from some of our defect data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+Accordingly, when this study explores the Jureczko data, we will use these five data sets.
+
+(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+
+	\begin{figure}[!t]
 	\begin{shaded}
@@
 
+\input{tex/dtree.tex}
+
 \subsection{Method4 = XTREES = Deltas in Decision Tree Branches}
@@
 The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.
-
-\input{tex/dtree.tex}
 

### 2015-08-23 tex/prediction_stats.tex
Updates from ShareLaTeX

@@
-\begin{figure*}[!t]
-	\scriptsize
+\begin{figure*}[bp]
+	\small
 	\begin{center}
@@
 				poi      & 1.5, 2, 2.5        & 936       & 3.0          & 442          & 64 \\
-				xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
-				log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
-				xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline
+				\rowcolor{lavenderpink}xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
+				\rowcolor{lavenderpink}log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
+				\rowcolor{lavenderpink}xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline
 
@@
 				56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
-				30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
-				32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
-				38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
+			\rowcolor{lavenderpink}	30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
+				\rowcolor{lavenderpink}32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
+				\rowcolor{lavenderpink}38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
 				\hline 
@@
 		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
-		Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
+		The data in  the  \colorbox{lavenderpink}{highlighted} rows, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
 	}\label{fig:j}

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 \section{Introduction}
-We propose a novel {\em planning} approach called XTREES for learning changes to a software system
+Business users are demanding  tools that support    business-level interpretations of their data.
+At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
+Panelists commented  ``prediction is all well and good, but what about decision making?''.
+Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
+So:
+\bi
+\item
+Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
+\item
+Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
+\ei
+In response to this business-level demands for planners,
+we propose a novel {\em planning} approach called XTREES for learning changes to a software system
 such that its performance ``improves'', according to some measure. For  case 
@@
 change(s) required to move from clump to clump.
-
-Our approach is important since business users are demanding analytics tools that support  a  business-level interpretation of their data.
-At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
-Panelists commented  ``prediction is all well and good, but what about decision making?''.
-Note that these panelists are more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
-So:
-\bi
-\item
-Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
-\item
-Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
-\ei
-This main contribution of this paper is to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
-can (a)~select useful planning methods while (b)~culling clearly
-inferior planning methods. 
-We apply this strategy to XTREES and three alternate planning methods to find:
-\bi
-\item Two methods that are clearly inferior;
-\item One method that proposes complex changes, which work as well as XTREES;
-\item And XTREES, that proposes very simple changes.
-\ei
-Of course, there are many more methods for generating plans and
-no   one paper can survey them all.
-But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
-to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
-Our hope is that other   researchers apply this strategy  to repeat and/or improve
-and/or refute our results. To that end, all our materials and data sets are available on-line, see
-github.com/ai-se/XTREES.
 
@@
 
- \begin{figure}
+ \begin{figure}[!t]
  \small
@@
 then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
+Note that, as might have been expected, none of the three bottom rows of \fig{j} were ``good''.
 
@@
 After clustering some training data using the WHERE algorithm of  \fig{where}.C, this method:
-\bi
-\item
- Replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
-continuous/discrete feature.
-\item Finds the closest centroid $C_j$ that has a better
+replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
+continuous/discrete feature. After that, it
+finds the closest centroid $C_j$ that has a better
 performance score. For defect data, ``better'' means fewer defects while for the config data,
 ``better'' means faster runtimes.
-\item Caches the  delta in the independent features between $C_i$ and $C_j$. For continuous
+Next, this method caches the  delta in the independent features between $C_i$ and $C_j$. For continuous
 features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
 in $C_j$. 
-\ei
-Next, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
+Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
 the nearest centroid $C_i$.  It then proposes a plan for improving that test case
@@
 This can be estimated using the Fayyad-Iranni INFOGAIN algorithm~\cite{FayIra93Multi}
-as follows:
-\bi
-\item Ran WHERE to generate   clusters;
-\item Discretize numeric features using the Fayyad-Iranni discretizer
-(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids);
-\item Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows.
-\item Let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
-\item Cull the features as per Papakroni; i.e. just use the $\beta=33\%$ most informative features
+as follows.
+
+Firstly, 
+run  WHERE to generate   clusters/ Next,
+discretize numeric features using the Fayyad-Iranni discretizer
+(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
+Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
+and 
+let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
+Cull the features as per Papakroni; i.e. just use the $\beta=33\%$ most informative features
 where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).
-\ei
+
 After that, Method2 works line Method1 but now the   plans  only
@@
 of the examples that are squashed together to form the centroids. Accordingly,
-our Method3 works like Method2 but some important differences:
-\bi
-\item Clusters are summarized in {\em two} examples: (1)~the centroid found in Method1 and
+our Method3 works like Method2 but some important differences.
+
+In Method3, clusters are summarized in {\em two} examples: (1)~the centroid found in Method1 and
 (2)~the best-in-cluster example; i.e. the  example in that cluster
 with the best performance score.
-\item As before, each centroid $C_i$ is annotated with the nearest centroid $C_j$
+Next, as before, each centroid $C_i$ is annotated with the nearest centroid $C_j$
 with a better performance score.
-\item When generating a plan for a test instance, Method3:
+Howver, after that,  when generating a plan for a test instance, Method3:
 \bi
@@
 and the best in that $C_j$ cluster.
-\ei
 \ei
@@
 \section{Threats to Validity}
+\section{Related Work}
+dashboards
+
+strucftureal learners (bayes)
+
 \section{Conclusion}
@@
 works for more than defects and tunrimes
+
+This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
+can (a)~select useful planning methods while (b)~culling clearly
+inferior planning methods. 
+We apply this strategy to XTREES and three alternate planning methods to find:
+\bi
+\item Two methods that are clearly inferior;
+\item One method that proposes complex changes, which work as well as XTREES;
+\item And XTREES, that proposes very simple changes.
+\ei
+Of course, there are many more methods for generating plans and
+no   one paper can survey them all.
+But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
+to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
+Our hope is that other   researchers apply this strategy  to repeat and/or improve
+and/or refute our results. To that end, all our materials and data sets are available on-line, see
+github.com/ai-se/XTREES.
 

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 \newcommand{\ee}{\end{enumerate}}
-\newcommand{\tion}[1]{\S\ref{sect:#1}}
+\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
 \newcommand{\fig}[1]{Figure~\ref{fig:#1}}
@@
 we propose a novel {\em planning} approach called XTREES for learning changes to a software system
-such that its performance ``improves'', according to some measure. For  case 
-studies, we apply XTREES to  reduce the expected value of:
+such that its performance ``improves'', according to some measure. This paper uses XTREES
+to reduce  the expected value of:
 \bi
@@
 
+The contributions of this paper are two-fold:
+\be
+\item
+The  XTREES planner, which has not been previously pubslished. For demonstration purposes, we show XTREES
+working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
+can determine that some rows are preferred and some are to be avoided).  
+\item An evaluation strategy used to comparatively assess planners.
+As  shown below, that strategy less us identify  inferior planners.
+With this strategy, other researchers could assess improvements or alternatives to the XTREE planner.
+\ei
 The rest of this paper is structured as follows.
-Since XTREES is a somewhat novel algorithm, we present it in tutorial form.
 After describing our test data,
@@
   to implement that   sanity check.
+
+ \begin{figure}[!t]
+ \small
+ \begin{center}
+ \begin{tabular}{r|rr}
+ data set & cases & \% defective\\\hline
+  ant &947& 22\\
+  camel& 1819& 19\\
+ jedit& 1257& 2\\
+ ivy &352& 11\\
+ log4j& 244 &92\\
+ lucebe &442 &59\\
+ poi& 936 &64\\
+ synapse &379 &34\\
+ velocity& 410& 34\\
+ xalan& 2411& 99\\
+ xerces &1055& 74
+ \end{tabular}
+ \end{center}
+ \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
+ \end{figure}
 
@@
  \begin{figure}[!t]
- \small
- \begin{center}
- \begin{tabular}{r|rr}
- data set & cases & \% defective\\\hline
-  ant &947& 22\\
-  camel& 1819& 19\\
- jedit& 1257& 2\\
- ivy &352& 11\\
- log4j& 244 &92\\
- lucebe &442 &59\\
- poi& 936 &64\\
- synapse &379 &34\\
- velocity& 410& 34\\
- xalan& 2411& 99\\
- xerces &1055& 74
- \end{tabular}
- \end{center}
- \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
- \end{figure}
-
- \begin{figure}
 \scriptsize
@@
 
+\section{A Frequently Asked Question}
+
+  Our evaluation strategy (discussed below) divides this data into a training a test set.
+  From the train set we apply a data miner (to learn a quality predictor) and
+  various planning methods (to learn different plans). Next, we try applying
+  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
+  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
+  in the changed examples.
+
+When discussing this strategy, a common concern is as follows. XTREES is evaluated by  comparing
+predicted performance scores before and after a planner makes changes to the feature values of an example:
+After making those
+changes, we may have a new example that has never seen before. Therefore, it should be asked:
+
+{\centerline {\em Can we trust the predictions made on such new examples?}}
+
+Our answer is ``yess'', if those changes select for the {\em region of competency} of the predictor. When we assess
+any data miner, there are two ``clouds'' of data:
+\bi
+\item One cloud is the space of training examples.
+\item Another cloud is the space of test examples.
+\ei
+We should mistrust a model learned from data mining if the test examples  too far removed from the training data; i.e. if
+these two clouds are too far apart. To test for ``too far'', we can run a data mining experiment to test how well
+a model learned from the training data applies to the test data. Such experiments return some performance value, from
+which we can say that if ever any new data falls into the combined region of the training+test cloud, then we should expect
+at least that level of performance.
+
+A corollary of the above is that if we are changing examples, and making predictions about those changed examples,
+then:
+\bi
+\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
+\item More specifically: predictions  about changes that  fall within the space of the training+test data, will be at least
+as accurate as the 
+\ei
+
 \input{tex/prediction_stats.tex}
-
   \section{Test Domains}
@@
    The Jurecko data records number of known defects for each class, where the classes are described in terms of
-  nearly two dozen metrics such as number of children, lines of code, etc.
-For details on the Jurecko data, see  \fig{ck} and \fig{jd}.
+  nearly two dozen metrics such as number of children, lines of code, etc. For details on the Jurecko data, see  \fig{ck} and \fig{jd}. 
+ For the most part, the methods of this paper treat
+  the Jurecko as a discrete class data set, where {\em defects} are true if the raw defect count is greater than zero.
+  The one exception will be the `best-in-cluster'' method that reflects on the numeric value of the raw defect count (see \tion{BIC}).
 
@@
   Siegmund et al. recorded how long each perturbation took to run a test suite.
-
-  Our evaluation strategy (discussed below) divides this data into a training a test set.
-  From the train set we apply a data miner (to learn a quality predictor) and
-  four planning methods (to learn different plans). Next, we apply
-  those plans to the  test set and ask the quality predictor to assess the results.
 
@@
 mention the $\beta=33\%$ most informative features.
-\subsection{Method3 = BIC = Method2 + Best-in-cluster}
+\subsection{Method3 = BIC = Method2 + Best-in-cluster}\label{sect:BIC}
 

### 2015-08-23 tex/jur_char.tex
Updates from ShareLaTeX

@@
 				\rowcolor{lightgray}
-				defect & defect & Boolean: where defects found in post-release bug-tracking systems.
+				nDefects & raw defect counts & Numeric: number of defects found in post-release bug-tracking systems.\\
+				\rowcolor{lightgray}
+				defect & defects present? & Boolean: if {\em nDefects} $>0$ then {\em true} else {\em false}
 			\end{tabular}
@@
 	\end{center}
-	\caption{OO measures used in our defect data sets.  Last line is
-		the dependent attribute (whether a defect is reported to  a
-		post-release bug-tracking system).}\label{fig:ck}
+	\caption{OO measures used in our defect data sets.  Last lines, shown in \textcolor{gray} denote the dependent variables.}\label{fig:ck}
 \end{figure*}

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 
-Our answer is ``yess'', if those changes select for the {\em region of competency} of the predictor. When we assess
+Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
 any data miner, there are two ``clouds'' of data:
@@
 these two clouds are too far apart. To test for ``too far'', we can run a data mining experiment to test how well
-a model learned from the training data applies to the test data. Such experiments return some performance value, from
-which we can say that if ever any new data falls into the combined region of the training+test cloud, then we should expect
-at least that level of performance.
-
-A corollary of the above is that if we are changing examples, and making predictions about those changed examples,
-then:
-\bi
-\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
-\item More specifically: predictions  about changes that  fall within the space of the training+test data, will be at least
-as accurate as the 
-\ei
+a model learned from the training data applies to the test data. Such experiments return some performance value.
+Given this, two things can be said about predictions on changed examples:
+\bi
+\item Predictions  about changes that  fall within the space of the training+test data, will be at least
+as reliable as the performance value found in the data mining experiment.
+\item More generally, predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
+\ei  
+Accodringly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
+training examples.  To see how this works, 
+ \fig{howxy} is from the {\em ivy} data
+set, which is one of the Jureczk data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
+changed  examples displaced after applying one of our planning methods.
+To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
+two  components (using PCA). Note that the  the   changed examples
+cases  (shown in green)  fall closer to the training cases (shown in gray) than
+the test cases (shown in red).  That is, if we believe
+that the gray data is  suitable for assessing
+the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
+that we believe even more in the predictions
+generated from the displaced data).
+
+This pattern of \fig{howxy} (where the displaced test cases are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+paper. Hence,  we can assert that
+predictors learned from these training data have some authority in the regions
+used to generate the plans\footnote{If the reader is concerned that this 2d projection
+losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
+have confirmed that building defect predictors from this 2d space are just as effective
+as predictors learned using all the defect attributes of \fig{ck}.}.
+
+\begin{figure}
+\begin{center}\includegraphics[width=3in]{figs/2d.png}\end{center}
+\caption{Training cases (in gray), test cases (in red) and 
+test cases displaced by HOW (in green). Note how the green (displaced) cases are closer to the  training
+date than the original test data
+i.e. predictors trained on the training cases and
+applied to the test cases should
+work just as well, or better, on the displaced test cases.}\label{fig:howxy}
+\end{figure}
+%XXX end
+
+That said, one thing to be said here
+is that we {\em strongly} recommend that predictors are assessed prior to planning. That is
+issue is explored further in the next section.
 

### 2015-08-23 Manuscript.tex
Updates from ShareLaTeX

@@
 \ei  
-Accodringly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
+Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
 training examples.  To see how this works, 
  \fig{howxy} is from the {\em ivy} data
-set, which is one of the Jureczk data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
+set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
 changed  examples displaced after applying one of our planning methods.
@@
 \begin{figure}
-\begin{center}\includegraphics[width=3in]{figs/2d.png}\end{center}
-\caption{Training cases (in gray), test cases (in red) and 
-test cases displaced by HOW (in green). Note how the green (displaced) cases are closer to the  training
+\includegraphics[width=1in]{_figs/2d.png} 
+\caption{Training examples (in gray), test examples (in red) and 
+changed exampled genreated by a planner (in green). Note how the green (displaced) cases are closer to the  training
 date than the original test data

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

@@
 \item
-The  XTREES planner, which has not been previously pubslished. For demonstration purposes, we show XTREES
+The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
 working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
@@
 
-\input{tex/jur_char.tex}
-
-\section{From Prediction to Planning}
+\section{Preliminaries}
+\subsection{From Prediction to Planning}
 
@@
   to implement that   sanity check.
+
+Another way  to find   changes to a project
+might be to rely
+on the peer review processes used by the 
+SE research community. This approach would propose changes to software
+projects that concur with internationally accept best practices. 
+There are two
+problems with that approach. Firstly, given the rapid pace of change in software
+engineering, we may be asking questions for which there is no currently accept
+``best practice''. For example, only very recently has there been any work
+of predicting software runtimes based on choices within Makefiles. While the work
+of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
+cannot yet represent ``internationally accept best practices'' since that work is so
+recent.
+
+Secondly, given the diversity of SE products and practices
+and personnel, it may well be that the current project being discussed is 
+substantively different to prior work. 
+Numerous recent {\em local learning} results compare (1) single models
+learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
+A repeated result in those studies in that the local models generated the better effort
+and defect predictions (better median results,
+lower variance in the predictions). This paper offers yet another locality result:
+\bi
+\item
+One standard rule in the literature
+is that it is useful to implement modules such that they are internally cohesive (use
+much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
+\item
+While that may be true in general, for particular classes other changes may be more important
+(later in this paper, we show one set of results were that is indeed the case).
+\ei
+In summary, 
+it is useful to have automatic methods to recommend changes. Such
+methods can fill in for human gurus (if such gurus are absent) or 
+to offer a second opinion.
+Also, prior to making automatic recommendations, it is wise to first stratify the data
+(clump it into related examples) then generate advice specific to each clump.
+Accordingly, the rest of this paper defines and evaluates
+automatic methods to find plans from
+  $N$ examples divided  into  many clumps.
+
+\input{tex/jur_char.tex}
+
+\subsection{Trusting the Changes}\label{sect:trust}
+   XTREES is evaluated by  comparing
+predicted performance scores before and after a planner makes changes to the feature values of an example:
+After making those
+changes, we may have a new example that has never seen before. Therefore, it must be asked:
+\bi
+\item
+{\em Can we trust the predictions made on such new examples?}
+\ei
+Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
+any data miner, there are two ``clouds'' of data:
+\bi
+\item One cloud is the space of training examples.
+\item Another cloud is the space of test examples.
+\ei
+For a visualization of these clouds, see \fig{howxy}.
+
+We should mistrust the predictions made by a  model   if it is being applied to examples far outside of its region of competency;
+i.e. to test examples that are too far removed from the training data.   
+To test for ``too far'', we can run a data mining experiment to test how well
+a model learned from the training data applies to the test data. Such experiments return some performance value.
+
+Note that predictions  about changes that  fall within the space of the training+test data, will be at least
+as reliable as the performance value found in the data mining experiment.
+Given this, one thing  can be asserted about predictions on changed examples:
+\bi
+\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
+\ei  
+Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
+training examples.  To see how this works, 
+ \fig{howxy} is from the {\em ivy} data
+set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
+changed  examples displaced after applying one of our planning methods.
+To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
+two  components (using PCA). Note that the  the   changed examples
+cases  (shown in green)  fall closer to the training cases (shown in gray) than
+the test cases (shown in red).  That is, if we believe
+that the gray data is  suitable for assessing
+the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
+that we believe even more in the predictions
+generated from the displaced data).
+
+This pattern of \fig{howxy} (where the displaced test cases are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+paper. Hence,  we can assert that
+predictors learned from these training data have some authority in the regions
+used to generate the plans\footnote{If the reader is concerned that this 2d projection
+losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
+have confirmed that building defect predictors from this 2d space are just as effective
+as predictors learned using all the defect attributes of \fig{ck}.}.
+
+\begin{figure}
+\includegraphics[width=2in]{_figs/twod.eps} 
+\caption{Training examples (in gray), test examples (in red) and 
+changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
+date than the original test data
+i.e. predictors trained on the training examples and
+applied to the test examples should
+work just as well, or better, on the changed test examples.}\label{fig:howxy}
+\end{figure}
+%XXX end
 
@@
 
-Another way  to find   changes to a project
-might be to rely
-on the peer review processes used by the 
-SE research community. This approach would propose changes to software
-projects that concur with internationally accept best practices. 
-There are two
-problems with that approach. Firstly, given the rapid pace of change in software
-engineering, we may be asking questions for which there is no currently accept
-``best practice''. For example, only very recently has there been any work
-of predicting software runtimes based on choices within Makefiles. While the work
-of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
-cannot yet represent ``internationally accept best practices'' since that work is so
-recent.
-
-Secondly, given the diversity of SE products and practices
-and personnel, it may well be that the current project being discussed is 
-substantively different to prior work. 
-Numerous recent {\em local learning} results compare (1) single models
-learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
-A repeated result in those studies in that the local models generated the better effort
-and defect predictions (better median results,
-lower variance in the predictions). This paper offers yet another locality result:
-\bi
-\item
-One standard rule in the literature
-is that it is useful to implement modules such that they are internally cohesive (use
-much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
-\item
-While that may be true in general, for particular classes other changes may be more important
-(later in this paper, we show one set of results were that is indeed the case).
-\ei
-In summary, 
-it is useful to have automatic methods to recommend changes. Such
-methods can fill in for human gurus (if such gurus are absent) or 
-to offer a second opinion.
-Also, prior to making automatic recommendations, it is wise to first stratify the data
-(clump it into related examples) then generate advice specific to each clump.
-Accordingly, the rest of this paper defines and evaluates
-automatic methods to find plans from
-  $N$ examples divided  into  many clumps.
-
  \begin{figure}[!t]
@@
 
-\section{A Frequently Asked Question}
-
-  Our evaluation strategy (discussed below) divides this data into a training a test set.
-  From the train set we apply a data miner (to learn a quality predictor) and
-  various planning methods (to learn different plans). Next, we try applying
-  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
-  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
-  in the changed examples.
-
-When discussing this strategy, a common concern is as follows. XTREES is evaluated by  comparing
-predicted performance scores before and after a planner makes changes to the feature values of an example:
-After making those
-changes, we may have a new example that has never seen before. Therefore, it should be asked:
-
-{\centerline {\em Can we trust the predictions made on such new examples?}}
-
-Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
-any data miner, there are two ``clouds'' of data:
-\bi
-\item One cloud is the space of training examples.
-\item Another cloud is the space of test examples.
-\ei
-We should mistrust a model learned from data mining if the test examples  too far removed from the training data; i.e. if
-these two clouds are too far apart. To test for ``too far'', we can run a data mining experiment to test how well
-a model learned from the training data applies to the test data. Such experiments return some performance value.
-Given this, two things can be said about predictions on changed examples:
-\bi
-\item Predictions  about changes that  fall within the space of the training+test data, will be at least
-as reliable as the performance value found in the data mining experiment.
-\item More generally, predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
-\ei  
-Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
-training examples.  To see how this works, 
- \fig{howxy} is from the {\em ivy} data
-set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
-changed  examples displaced after applying one of our planning methods.
-To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
-two  components (using PCA). Note that the  the   changed examples
-cases  (shown in green)  fall closer to the training cases (shown in gray) than
-the test cases (shown in red).  That is, if we believe
-that the gray data is  suitable for assessing
-the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
-that we believe even more in the predictions
-generated from the displaced data).
-
-This pattern of \fig{howxy} (where the displaced test cases are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
-paper. Hence,  we can assert that
-predictors learned from these training data have some authority in the regions
-used to generate the plans\footnote{If the reader is concerned that this 2d projection
-losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
-have confirmed that building defect predictors from this 2d space are just as effective
-as predictors learned using all the defect attributes of \fig{ck}.}.
-
-\begin{figure}
-\includegraphics[width=1in]{_figs/2d.png} 
-\caption{Training examples (in gray), test examples (in red) and 
-changed exampled genreated by a planner (in green). Note how the green (displaced) cases are closer to the  training
-date than the original test data
-i.e. predictors trained on the training cases and
-applied to the test cases should
-work just as well, or better, on the displaced test cases.}\label{fig:howxy}
-\end{figure}
-%XXX end
-
-That said, one thing to be said here
-is that we {\em strongly} recommend that predictors are assessed prior to planning. That is
+That said, the above comes with two important caveats:
+\bi
+\item 
+We   strongly recommend that predictors are assessed prior to planning. That
 issue is explored further in the next section. 
-
-\input{tex/prediction_stats.tex}
+\item
+Where possible, planners should be assessed via some external
+oracle that can accurately assess new examples. For an example of that kind of analysis,
+see our Validity section, later in this paper.
+\ei
+
   \section{Test Domains}
@@
 
-  Note that this approach depends on having effective predictors for assessing the results.
+\input{tex/prediction_stats.tex}
+
+  Our evaluation strategy (discussed below) divides this data into a training a test set.
+  From the train set we apply a data miner (to learn a quality predictor) and
+  various planning methods (to learn different plans). Next, we try applying
+  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
+  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
+  in the changed examples.
+
+  As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
   For the Siegmind data, this criteria was   relatively easy to achieve.
@@
   \ei
-In the following, note our convention. All variables
+
+  Note that all our methods are {\em trust-increasing} in the manner described in \tion{trust}.
+
+The following descriptions of the methods adopts the following convention. All variables
   set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

@@
 
-\section{Preliminaries}
+\section{Preliminaries}\label{sect:prelim}
 \subsection{From Prediction to Planning}
@@
  \fig{howxy} is from the {\em ivy} data
-set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
+set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
 changed  examples displaced after applying one of our planning methods.
@@
 oracle that can accurately assess new examples. For an example of that kind of analysis,
-see our Validity section, later in this paper.
-\ei
-
-  \section{Test Domains}
+see the Validity section, later in this paper.
+\ei
+
+  \section{Test Data}
 To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
@@
  It proved to be  more complicated to commission the Jureczko data sets for this study.
- For that data, we found:
-\bi
-\item
- Quality predictors built from this data are far from perfect;
-\item 
-But several of these predictors could
+ For that data, we found that the
+ quality predictors built from this data are far from perfect;
+However, for some data sets, the  predictors could
 be salvaged using the techniques discussed in this section.
-\ei
+
  \fig{j} shows our preliminary studies with the Jureczko   data.
@@
 of classes and testing on smaller test suites).
-Note the three bottom \colorbox{lavenderpink}{highlighted} rows marked with $\times$: these contain predominately
+Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately
 defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
@@
 then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
-Note that, as might have been expected, none of the three bottom rows of \fig{j} were ``good''.
+Note that, as might have been expected, none of the \colorbox{lavenderpink}{three bottom rows} of \fig{j} were ``good''.
 
@@
 an over-sampling of minority-class examples.
-Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
 for tree generation, etc.
@@
 Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
-for  $n$ discrete values occurring at frequency
-$f_1,f_2,.. f_n$. That is, \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.
+for  $n$ discrete values  at frequency
+$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.
 
@@
 			\item Project each case $Z$
-			to position $x$ along a    lines that  runs from $X$ to $Y$, If $a,b$ re distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
+			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
 			\item Split the data at the median $x$ value of all cases.
-			\item For   splits larger than  $\alpha=\sqrt{N}$, call this algorithm recursively  from step1.
+			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
 		\end{enumerate}
@@
 
-	Find a split in the values of the independent features that most reduces the variability
-  of the dependent feature\footnote{}. The data is then divided on that feature's split and 
-  the tree learner recurses on each split (stopping when any split is smaller than $\alpha=\sqrt{N}$).
+	Find a split in the values of  independent features that most reduces the variability
+  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
+   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.
 
@@
   \ei
-
-  Note that all our methods are {\em trust-increasing} in the manner described in \tion{trust}.
-
+  Note that all our methods have the two properties proposed in \tion{prelim}:
+  \bi
+  \item They find different plans in different parts of the data;
+  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
+  i.e. they change changing examples such that they move {\em closer} to the training data.
+  \ei
 The following descriptions of the methods adopts the following convention. All variables
@@
   In this paper, we show our current settings to these variables produces useful
-  results. In other work~\cite{krall14}, we are exploring tuning methods to automatically
+  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
   find better settings but  we have nothing definitive yet to report
@@
  {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
-  reporting the delta between the cluster centroids. The advantage of Method1 is its
-  simplicity. However, as shown by our evaluation strategy, it is not an useful
-  method for generating effective plans.
-
-After clustering some training data using the WHERE algorithm of  \fig{where}.C, this method:
+  reporting the delta between the cluster centroids. 
+  After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
 replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
@@
 finds the closest centroid $C_j$ that has a better
-performance score. For defect data, ``better'' means fewer defects while for the config data,
-``better'' means faster runtimes.
-Next, this method caches the  delta in the independent features between $C_i$ and $C_j$. For continuous
+performance score. For defect data, ``better'' means fewer defective examples while for the config data,
+``better'' means lower median runtimes for the examples in that cluster.
+Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
 features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
@@
 
-\subsection{Method2 = CD+FS =  Method1  + \\Feature Selection }
+\subsection{Method2 = \\\CD+FS =  Method1  + Feature Selection }
 

### 2015-08-24 tex/prediction_stats.tex
Updates from ShareLaTeX

@@
-\begin{figure*}[bp]
+\begin{figure*}[!t]
 	\small
@@
 
-	\caption{Training and test {\em data set properties} for the Jureczko data sets,
-		sorted in ascending order of \% defective examples.
+	\caption{Training and test {\em data set properties} for  Jureczko data ,
+		sorted by \% defective examples.
 		On the right-hand-side, we show the {\em results from learning}.
@@
 		(and note that, after tuning, there are more ``good'' than before).
-		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
-		The data in  the  \colorbox{lavenderpink}{highlighted} rows, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
+		Data   marked with ``$\star$'' show large improvements in performance, after tuning
+		(lower {\em pf} or higher {\em pd}).
+		The data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
 	}\label{fig:j}

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

@@
  improve on the predicted values.
-This paper defines an evaluation strategy for evaluating such  planners and applies it to four different planning methods.
-It is shown that our  strategy  can (a) recognize   inferior planning methods and  (b) select useful planning methods.
-Using the best method found by this evaluation strategy,  we can
-reduce
+
+This presents such a tool, called XTREES and evaluates it against three alternatives.
+In our experiments, XTREES reduced
  the expected values of defect counts and  runtimes of software projects to    
@@
 \section{Preliminaries}\label{sect:prelim}
+
 \subsection{From Prediction to Planning}
@@
 a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
-then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''
+then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.
 
@@
 
-This pattern of \fig{howxy} (where the displaced test cases are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
 paper. Hence,  we can assert that
@@
 
-	\begin{figure}[!t]
+	\begin{figure} 
 	\begin{shaded}
@@
 
- {\bf Figure 5.A: Measuring Variability}
+ {\bf Figure 7.A: Measuring Variability}
 
@@
 
-	{\bf Figure 5.B: Measuring distance}
+	{\bf Figure 7.B: Measuring distance}
 
@@
 
-	{\bf Figure 5.C: Top-down Clustering with WHERE}
+	{\bf Figure 7.C: Top-down Clustering with WHERE}
 
@@
 
-	{\bf Figure 5.D: Top-down division with Decision Trees}
+	{\bf Figure 7.D: Top-down division with Decision Trees}
 
@@
 	 ~\hrule~
+
+	 	{\bf Figure 7.E: Finding the most informative rows}
+
+Discretize all numeric features using the Fayyad-Iranni discretizer~\cite{FayIra93Multi}
+(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
+Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
+and 
+let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
+Cull the the features as per Papakroni~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
+where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).
+
+~\hrule~
+
 	  \end{shaded}
@@
   \bi
-  \item They find different plans in different parts of the data;
+  \item They are {\em local learners}; i.e. different test examples
+  will be given plans that are specialized  to their particulars;
   \item They  are {\em trust-increasing} in the manner described in \tion{trust};
@@
 
-\subsection{Method2 = \\\CD+FS =  Method1  + Feature Selection }
-
-{\em Summary:} Method2 works like Method1, but this time, it only mentions the most
-informative features. Hence, Method2's plans are simpler.
+\subsection{Method2 =  CD+FS = \\ Method1  + Feature Selection }
+
+{\em Summary:} Method2 works line Method1 but now the   plans  only
+mention the $\beta=33\%$ most informative features. Hence, Method2's plans are simpler.
 
@@
 This can be estimated using the Fayyad-Iranni INFOGAIN algorithm~\cite{FayIra93Multi}
-as follows.
-
-Firstly, 
-run  WHERE to generate   clusters/ Next,
-discretize numeric features using the Fayyad-Iranni discretizer
-(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
-Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
-and 
-let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
-Cull the features as per Papakroni; i.e. just use the $\beta=33\%$ most informative features
-where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).
-
-After that, Method2 works line Method1 but now the   plans  only
-mention the $\beta=33\%$ most informative features.
-\subsection{Method3 = BIC = Method2 + Best-in-cluster}\label{sect:BIC}
+of \fig{where}.E.
+
+\subsection{Method3 = BIC = \\  Method2 + Best-in-cluster}\label{sect:BIC}
 
@@
 One problem with Method2 is that it makes little use of the detailed knowledge
-of the examples that are squashed together to form the centroids. Accordingly,
-our Method3 works like Method2 but some important differences.
-
-In Method3, clusters are summarized in {\em two} examples: (1)~the centroid found in Method1 and
-(2)~the best-in-cluster example; i.e. the  example in that cluster
+of the examples that were squashed together to form the centroids. 
+Hence,
+in Method3, we collect more information about each cluster.
+
+Method3 summarizes clusters into {\em two} examples: (1)~the centroid $C_x$ found in Method1 and
+(2)~the best-in-cluster $B_x$  example; i.e. the  example in that cluster
 with the best performance score. 
-Next, as before, each centroid $C_i$ is annotated with the nearest centroid $C_j$
-with a better performance score.
-Howver, after that,  when generating a plan for a test instance, Method3:
-\bi
-\item
-Find the test cases's nearest centroid $C_i$;
-\item
-Finds the associated better centroid $C_j$ and its associated best-in-cluster example.
-\item
-Computes a plan to change the test case  from the delta between the test case
-and the best in that $C_j$ cluster.
-\ei
+For the Siegmud data, $B_x$ is the cluster member with the fastest runtime;
+for the Jureczko data, $B_x$  is the cluster member with lowest raw defect count
+(resolving ties at random).
+
+Next,  each centroid is connected by a line to its nearest centroid.
+ In that line, we say $C_i,C_j$ are the centroids associated with worst/best centroid performance scores.
+
+For each test instance, Method3 
+find the nearest line; selects the  best end $C_j$ of that line;  then extracts $B_j$ (which is the
+ best-in-cluster associated with  $C_j$).
+The returned plan is then computed   from the delta between the test case
+and $B_j$.
 Note that, as with Method2, this method's plans only mentions the  $\beta=33\%$ most informative features. 
@@
 
-\subsection{Method4 = XTREES = Deltas in Decision Tree Branches}
+\subsection{Method4 =  XTREES =\\ Deltas in Decision Tree Branches}
 
@@
 the values in the target class. In theory, this extra knowledge can lead to better data mining.
-Accordingly, Method4 uses a supervised method to generate its plans.
-
-In summary, XTREES uses the decision tree algorithm of \fig{where}.D to divide the data.
-Decision trees have branches down to leaves. 
-XTREES builds plans by asking three questions:
+Accordingly, XTREES uses a supervised method to generate its plans.
+Specifically, it  uses the decision tree algorithm of \fig{where}.D to divide the data.
+Next, XTREES builds plans from the branches of the decision trees. It does so by asking three questions:
 \be
@@
 
-\input{tex/results_models.tex}
-
 \section{Experiments}
 
-\begin{figure}[!t]
+\begin{figure}[!b]
 \small 
@@
                 \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
-                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
+                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREES)}
               \end{array}\right.
@@
     \left\{\begin{array}{l@{~}l}
-           \mathit{actual}& =\mathrm{Defects \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
-           \mathrm{\bf if\;}\mathit{actual} & >  \mathit{0}\\
+           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
+           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
            \mathrm{\bf then} &
@@
                 \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
-                \mathrm{{\bf return}\;} \mathit{actual},\mathit{after}
+                \mathrm{{\bf return}\;} R=\frac{\mathit{after}}{\mathit{before}}
             \end{array}
@@
 
-\subsection{Setup}
-
-In the experiments shown below,  we use the planners to comment on how to reduce
-software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
-tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}.
-
-\fig{work} shows our proposal to verify if the recommendations made by our planners actually work. Accordingly, we divide the
+This section describes an experimental design (and results) for evaluating the above four methods.
+
+\subsection{A Strategy for Evaluation Planners}
+
+\subsubsection{Performance Measures}
+
+For a planning system designed for business users, we seek two properties:
+\bi
+\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
+\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
+Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
+\ei
+ In the following, we will show that XTREES and one other method (BIC) are equally effective
+ but XTREES is more succinct (proposes changes to fewer features).
+
+\subsubsection{Collecting Data}
+
+\fig{work} shows our experimental design.  Initially, we divide the
 project data  into two disjoint sets {\em train} and {\em test}
 (so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
-Next, from the train set, we build both a {\em planner} as well
-as a {\em  predictor}.
-Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper, our {\em planner} will be from \ref{clust_contrast} and our {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} and Random Forest Regressor taken from the SciKit Learn toolkit~\cite{Pedregosa2012}.
-
+Next, from the train set, we build both a {\em planner} and
+ a {\em  predictor}.
+
+Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
+\bi
+\item Our {\em planner} will be one of Methods 1,2,3,4;
+\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.
+We use these
+data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
+\ei
 As for the {\em test} data, this is passed to the { predictor}
-to measure the baseline prediction accuracy, false alarm, and recall.
-If our { predictors} fail to perform adequately on the test data,
+to measure performance statistics related to effectiveness.
+
+If our { predictors} fail to perform effectively on the test data,
 then we cannot trust them to comment on our plans. Accordingly,
-if that performance is unsatisfactory, we abort.
+if that performance is unsatisfactory, we abort. Recall from the above discussion that this step indicated
+we should not use some of the  Jureczko data.
+
 Else, we (1)~apply the { planner} to alter the {\em test} data;
 then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions.
-
-\subsection{Data}
+then (3)~return data on the {\em before, after} predictions expressed as the ratio $R=\frac{\mathit{after}}{\mathit{before}}$.
+This is a unit-less ratio with the following properties:
+\bi
+\item If $R= 1$, this means  ``no change from baseline''; 
+\item If $R < 1$, this indicates some reduction to the baseline;
+\item If $R > 1$, this indicates ``optimization failure''.
+\ei
+Also, to measure succinctness, we count the number of features in the plan.
+
+\subsubsection{Studying Repeated Samples}
+Our methods contain some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
+  XTREES' occasional use of a random guess when deciding what part of a discretized range to include in the plan
+  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
+  (with different random number seeds).
+The the value 40 was chosen to be  larger than the 30 samples  required
+to satisfy the central limit theorem.
+
+Our results are presented by the line diagrams like those shown on the right-hand-side of the following example table.
+The black dot shows the median $R$ value and the horizontal likes stretch over the inter-quartile
+range (hereafter, IQR) that is the space from the 25th percentile value to
+the 75th percentile value.
+
+\begin{center}
+
+{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}}\hline
+\arrayrulecolor{lightgray}
+\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
+1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
+2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
+2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
+3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115} \arrayrulecolor{black}\\\hline \end{tabular}}
+\end{center}
+
+In this example, the rows are  sorted on the median values of each method. Note that all the methods
+have   $R<1$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
+while XTREE achieved the greatest reduction (down to 49\% of the original value).
+
+The above table has a  left-hand-side  {\bf Rank} column. This column reports if
+the values for each method are statistically different and are more than trivially different. 
+In this example, table, CD and BIC are ranked together while XTREES and CD+FS are ranked best and worst, respectively.
+
+  \subsubsection{Statistical Methods}
+  We compute {\bf Rank} using a Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}. 
+In that test, using the median values of each method,  
+sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
+Then,
+splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
+ differences  in the observed performances
+before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
+ \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
+Scott-Knott  applies an statistical  hypothesis test $H$ to check
+if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping). 
+If so, Scott-Knott  recurses on the splits.
+In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+
+For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
+For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
+These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
 
@@
 
-The clear advantage of our current work is that it can be readily and automatically applied to domains with historical data.
-However, the obvious drawback with this approach is that our assessment of the plans is only as good as the predictor.
-To mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for {\em satisfactory} thresholds.
-
-\subsection{Assessment}
-\fig{work} showed how we build and test our planners.
+\subsection{Results}
+
+all the smoting and tuning running in train.
+
   The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
 Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
-Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
 \bi
@@
 \ei
-The 40 repeats defined above were repeated for each planner.
-A Scott-Knott procedure was used to test for statistical differences in the result.
-The Scott-Knott procedure was recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{mittas13}.
-
-This method
-sorts a list of $l$ treatments with $ls$ measurements by their median
-score. It then
-splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
- differences  in the observed performances
-before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
- $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
-Scott-Knott  applies an  hypothesis test $H$ to check
-if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.
-
-The hypothesis test $H$ used here is a conjunction of the A12 effect size test and a non-parametric bootstrap sampling.
-In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
-For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
-For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
-These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
-Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
-
-\subsection{Results}
+The 40 repeats defined above were repeated for each planner and the results processed as described in the last section.
+
 \input{tex/deltas_avg.tex}
 The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
-\subsection{Discussion}	whye where?
-Recent results suggest that, when clustering on independent variables of software project data,
+\subsection{Discussion}
+
+\section{Related work}
+results suggest that, when clustering on independent variables of software project data,
 	we are free to select from a wide range of 
@@
 contains signals not associated with the target variable.
-
-\section{Threats to Validity}
+	ats to Validity}
+
+\input{tex/results_models.tex}
+
 \section{Related Work}
@@
 strucftureal learners (bayes)
+
+Many data mining methods offer a succinct summary of project data
+e.g. see the Bayes nets
+favored by Dejaeger et al.~\cite{Dejaeger13}; or   the regression models
+learned by Nagappan~\cite{Nagappan05}.
+Another  technique is to offer a dashboard that visualizes project data.
+For example, the dashboards
+provided by the Hackystat team~\cite{Johnson09} presents extensive and detailed
+summaries of programmer behavior.
+
+Some users do not like being told what to do... they prefer to have the current world explained to them 
+-- but those dashboards typically do not learn threshold points
+(that report where action is needed) nor do they offer specific advice on what actions to take.
 

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

@@
 With this strategy, other researchers could assess improvements or alternatives to the XTREE planner.
-\ei
+\ee
+In all, this paper explores and comparatively assesses four methods for learning planners (and XTREES is Method\#4).
+As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}. As shown below,
+that method is a comparatively far less effective than XTREES. Hence, we   recant that prior work and, based on the results of this
+paper, we now endorse  XTREES.
+
 The rest of this paper is structured as follows.
@@
 
-\begin{figure}
-\includegraphics[width=2in]{_figs/twod.eps} 
+\begin{figure}[!t]
+  \includegraphics[width=2in,bb=0 0 100 100]{figs/aa.eps} 
+ % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
 \caption{Training examples (in gray), test examples (in red) and 
@@
 \end{figure}
+
 %XXX end
@@
 Project & Domain & Lang. & LOC & Features & Config\\\hline
-Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
-Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
+BDBC: Berkeley DB   & Database & C & 219,811 & 18 & 2560\\
+BDBJ: Berkeley DB   & Database & Java & 42,596 & 32  & 400\\
 Apache & Web Server & C & 230,277 & 9 & 192\\
@@
 \begin{figure}[!t]
-\includegraphics[width=1\linewidth]{_figs/BDBC.eps}
+\includegraphics[width=1\linewidth]{figs/BDBC.eps}
 \caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
@@
 
-  \section{Test Data}
+  \section{Test Data}\label{sect:tesd}
 To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
@@
   \ei
-The following descriptions of the methods adopts the following convention. All variables
+
+Our  description of the methods adopts the following convention. All variables
   set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
@@
 of \fig{where}.E.
+
+As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}.
 
@@
 
-\subsubsection{Performance Measures}
-
-For a planning system designed for business users, we seek two properties:
-\bi
-\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
-\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
-Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
-\ei
- In the following, we will show that XTREES and one other method (BIC) are equally effective
- but XTREES is more succinct (proposes changes to fewer features).
-
 \subsubsection{Collecting Data}
@@
 
-{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}}\hline
+{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
 \arrayrulecolor{lightgray}
@@
 2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
-3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115} \arrayrulecolor{black}\\\hline \end{tabular}}
+3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
 \end{center}
@@
 
-\input{tex/results_jur.tex}
-\input{tex/results_conf.tex}
+\subsection{Results}
 
@@
 \begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
+\includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
 \label{fig:delta_ivy}
@@
 \begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
+\includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
 \label{fig:delta_apache}
@@
 X-axis lists the various features that make up a data set.
-The y-axis contains the frequency of change in \%.}
+The y-axis contains the frequency of change in \%.}\label{fig:changed}
 \end{figure}
 
-\subsection{Results}
-
-all the smoting and tuning running in train.
-
-  The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
+\input{tex/results_jur.tex}
+\input{tex/results_conf.tex}
+
+\subsubsection{Effectiveness Results}
+
+For a planning system designed for business users, we seek two properties:
+\bi
+\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
+\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
+Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
+\ei
+
+XTREES performs well in terms of  effectiveness. 
+\fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
+In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
 Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
@@
 \item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
-\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
-\ei
-The 40 repeats defined above were repeated for each planner and the results processed as described in the last section.
+\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
+ all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
+occurred in the {\em train} phase of \fig{work}.
+\ei
+The results show planning can be very effective-- some of the expected value of the reductions are very large:
+\bi
+\item
+Down to 22\% of the original baseline in Ant of \fig{jur};
+\item
+Down to 22\% of the original baseline in X264  of \fig{conf};
+\ei
+In all these results,  with one exception, XTREES always is the top-ranked method and with a median improvement at least 10\% lower (i.e.
+better) than the next ranked method.
+The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
+so XTREES can only reduce the expected values of the runtimes by 1\%.
+
+From these results, measured in terms of effectiveness, we recommend supervised-based division of the data (as used in XTREE) over unsupervised clustering
+(as used in the other methods). As to the very simple methods (CD and CD+FSS), these were proposed in prior publications
+on this work~\cite{me12c}. We now   observe  that that prior work was poorly evaluated so the recommendations
+of that prior paper  must now be rejected.
+
+As to succinctness , that is discussed in the next section.
+
+\subsubsection{Succinctness Results}
 
 \input{tex/deltas_avg.tex}
-The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
-\subsection{Discussion}
+
+\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
+The top plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the bottom
+plot shows a Siegmund data set ({\em BDBJ}).
+
+As predicted in the above discussion on Method 1 (CD), this method always proposes changes  to most features. Method3 (BIC) often proposes changes to many features, which 
+would make its plan tedious to implement. 
+Method 2 (CD+FS)
+proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
+probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREES) rarely fixates
+on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
+all over the space of features.
+
+\fig{changed} reports the changes just for   {\em lucene}) and {\em BDBJ}. \fig{types} shows the mean changes for features found by our
+four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
+Method3 (BIC) make the most changes while XTREES and CD+FS make the fewest.
 
@@
 contains signals not associated with the target variable.
-	ats to Validity}
+	ats to Validity
 

### 2015-08-24 tex/results_models.tex
Updates from ShareLaTeX

@@
 
-~\\
-
-In this figure, {\bf 50th}
-is the median value and  {\bf 25-75th} is the second and third quartile range.
-The right-hand-side horizontal lines show the {\bf 50th} median value (as a black dot) in the middle
-of the {\bf 25-75th} percentile range (as lines). All values here are expressed as ratios
-of the median values seen in the untreated data, so {\em smaller} values are {\em better}).
-A median {\bf 50th} value of
-1.0 means ``no change from
-baseline''; numbers less than 1.0 indicate some reduction to the baseline;
-and numbers more than 1.0 indicate ``optimization failure''.
-The {\bf Rank} column on the left-hand-side show the results of a statistical
-analysis and results ranked ``1'' are best. Results change {\bf Rank} if they are significantly different and pass
- a post-hoc effect size test.
-
  ~\hrule~

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

@@
 
-This presents such a tool, called XTREES and evaluates it against three alternatives.
-In our experiments, XTREES reduced
+This presents such a tool, called XTREE and evaluates it against three alternatives.
+In our experiments, XTREE reduced
  the expected values of defect counts and  runtimes of software projects to    
@@
 In response to this business-level demands for planners,
-we propose a novel {\em planning} approach called XTREES for learning changes to a software system
-such that its performance ``improves'', according to some measure. This paper uses XTREES
-to reduce  the expected value of:
-\bi
-\item  Defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
-\item Runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.
-\ei 
+we propose a novel {\em planning} approach called XTREE for learning changes to a software system
+such that its performance ``improves'', according to some measure. This paper uses XTREE
+to reduce  the expected value of the
+defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
+and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.
+
 The approach discussed here is novel since most data miners used in software analytics
@@
 
-The contributions of this paper are two-fold:
-\be
-\item
-The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
-working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
-can determine that some rows are preferred and some are to be avoided).  
-\item An evaluation strategy used to comparatively assess planners.
+The contributions of this paper are two-fold. Firstly, we offer
+   XTREE, which has not been previously published.
+   XTREE     works  on table of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
+XTREE returns plans that are changes to feature values in the ``bad'' rows to make them more ``good''.
+Secondly, an evaluation strategy used to comparatively assess planners.
 As  shown below, that strategy less us identify  inferior planners.
-With this strategy, other researchers could assess improvements or alternatives to the XTREE planner.
-\ee
-In all, this paper explores and comparatively assesses four methods for learning planners (and XTREES is Method\#4).
-As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}. As shown below,
-that method is a comparatively far less effective than XTREES. Hence, we   recant that prior work and, based on the results of this
-paper, we now endorse  XTREES.
+For example, when exploring software runtimes and software defects,
+we show that XTREE is significantly better than the methods proposed in two of our prior publications~\cite{me12c}.
 
@@
 this paper  walks the reader through four methods that start
-with a standard clustering algorithm, the  progresses on to XTREES. The second
+with a standard clustering algorithm, the  progresses on to XTREE. The second
 half of the paper then presents an extensive evaluation of these methods using 
@@
 \subsection{Trusting the Changes}\label{sect:trust}
-   XTREES is evaluated by  comparing
+   XTREE is evaluated by  comparing
 predicted performance scores before and after a planner makes changes to the feature values of an example:
@@
 For a visualization of these clouds, see \fig{howxy}.
+
+\begin{figure}[!t]
+  \includegraphics[width=\linewidth]{figs/twodee.eps} 
+ % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
+\caption{Training examples (in gray), test examples (in red) and 
+changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
+date than the original test data
+i.e. predictors trained on the training examples and
+applied to the test examples should
+work just as well, or better, on the changed test examples.}\label{fig:howxy}
+\end{figure}
 
@@
 \ei  
-Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
+Accordingly, we should use  {\em trust-increasing} planners. That is, we should use planners that push changde examples {\em towards} the
 training examples.  To see how this works, 
@@
 
+(Aside: If the reader is concerned that this 2d projection
+losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
+have confirmed that building defect predictors from this 2d space are just as effective
+as predictors learned using all the defect attributes of \fig{ck}.)
+
 This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
@@
 predictors learned from these training data have some authority in the regions
-used to generate the plans\footnote{If the reader is concerned that this 2d projection
-losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
-have confirmed that building defect predictors from this 2d space are just as effective
-as predictors learned using all the defect attributes of \fig{ck}.}.
-
-\begin{figure}[!t]
-  \includegraphics[width=2in,bb=0 0 100 100]{figs/aa.eps} 
- % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
-\caption{Training examples (in gray), test examples (in red) and 
-changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
-date than the original test data
-i.e. predictors trained on the training examples and
-applied to the test examples should
-work just as well, or better, on the changed test examples.}\label{fig:howxy}
-\end{figure}
+used to generate the plans.
 
@@
 
-That said, the above comes with two important caveats:
+That said, the above comes with some important caveats:
 \bi
@@
 We   strongly recommend that predictors are assessed prior to planning. That
-issue is explored further in the next section. 
+issue is explored further in the \label{sect:tesd}.
+\item
+Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
 \item
@@
 oracle that can accurately assess new examples. For an example of that kind of analysis,
-see the Validity section, later in this paper.
+see the Validity  in \tion{valid}.
 \ei
@@
   software system   configuration data from by  Siegmund et al.~\cite{sven12}.
-  See github.com/ai-se/XTREES\#data for full access to this data.
+  See github.com/ai-se/XTREE\#data for full access to this data.
 
@@
 
-\section{Four Planning Methods}
+\section{Four Planning Methods}\label{sect:planners}
 In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
 algorithm should be baselined against simpler alternatives. Accordingly,
-this section described XTREES (which we call Method4) against three  
+this section described XTREE (which we call Method4) against three  
 alternatives: 
@@
 
-\subsection{Method4 =  XTREES =\\ Deltas in Decision Tree Branches}
+\subsection{Method4 =  XTREE =\\ Deltas in Decision Tree Branches}
 
@@
 the values in the target class. In theory, this extra knowledge can lead to better data mining.
-Accordingly, XTREES uses a supervised method to generate its plans.
+Accordingly, XTREE uses a supervised method to generate its plans.
 Specifically, it  uses the decision tree algorithm of \fig{where}.D to divide the data.
-Next, XTREES builds plans from the branches of the decision trees. It does so by asking three questions:
+Next, XTREE builds plans from the branches of the decision trees. It does so by asking three questions:
 \be
@@
 \ee
-The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.
+The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{XTREE_bare}.
 
@@
                 \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
-                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREES)}
+                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
               \end{array}\right.
@@
 Our methods contain some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
-  XTREES' occasional use of a random guess when deciding what part of a discretized range to include in the plan
-  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
+  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
+  (see \fig{XTREE_bare}). Hence, we report the $R$ values seen in 40 repeated runs
   (with different random number seeds).
@@
 the values for each method are statistically different and are more than trivially different. 
-In this example, table, CD and BIC are ranked together while XTREES and CD+FS are ranked best and worst, respectively.
+In this example, table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.
 
@@
 
-XTREES performs well in terms of  effectiveness. 
+XTREE performs well in terms of  effectiveness. 
 \fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
@@
 \ei
-In all these results,  with one exception, XTREES always is the top-ranked method and with a median improvement at least 10\% lower (i.e.
+In all these results,  with one exception, XTREE always is the top-ranked method and with a median improvement at least 10\% lower (i.e.
 better) than the next ranked method.
 The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
-so XTREES can only reduce the expected values of the runtimes by 1\%.
+so XTREE can only reduce the expected values of the runtimes by 1\%.
 
@@
 proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
-probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREES) rarely fixates
+probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREE) rarely fixates
 on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
@@
 four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
-Method3 (BIC) make the most changes while XTREES and CD+FS make the fewest.
+Method3 (BIC) make the most changes while XTREE and CD+FS make the fewest.
+
+\section{Validity}\label{sect:valid}
 
@@
 inferior planning methods. 
-We apply this strategy to XTREES and three alternate planning methods to find:
+We apply this strategy to XTREE and three alternate planning methods to find:
 \bi
 \item Two methods that are clearly inferior;
-\item One method that proposes complex changes, which work as well as XTREES;
-\item And XTREES, that proposes very simple changes.
+\item One method that proposes complex changes, which work as well as XTREE;
+\item And XTREE, that proposes very simple changes.
 \ei
@@
 no   one paper can survey them all.
-But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
-to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
+But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
+to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
 Our hope is that other   researchers apply this strategy  to repeat and/or improve
 and/or refute our results. To that end, all our materials and data sets are available on-line, see
-github.com/ai-se/XTREES.
+github.com/ai-se/XTREE.
 
@@
 The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}.
-For full algorithmic details, refer to \fig{xtrees_bare}.
+For full algorithmic details, refer to \fig{XTREE_bare}.
 
@@
 The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity.
-We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).
+We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{XTREE_bare}).
 

### 2015-08-24 Manuscript.tex

@@
  \definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
+ \definecolor{celadon}{rgb}{0.67, 0.88, 0.69}
 \newcommand{\G}{\cellcolor{green}}
@@
 \title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
-\numberofauthors{3} 
+\numberofauthors{4} 
 \author{  
@@
 Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
-       \affaddr{Com. Sci., NC State, USA}\\
+       \affaddr{CS, NC State, USA}\\
        {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
 \alignauthor
-Naveen   Lekkalapudi\\
+Andrian Marcus \\
+       \affaddr{CS, UT Dallas  } \\ 
+       \affaddr{Texas, USA}\\ 
+       {amarcus@utdallas.edu}
+       \alignauthor
+Naveen  \\ Lekkalapudi\\
  \affaddr{Bellhops } \\ 
@@
 \begin{abstract}
- As business users demand more insightful
- analytics, data scientists need to change
- their tools. Instead of just predicting 
+ Business users now demand more insightful
+ analytics. Instead of just predicting 
  some result, they also need tools that generate ``plans'';
  i.e. specific suggestions on  what to change  in order to
- improve on the predicted values.
-
-This presents such a tool, called XTREE and evaluates it against three alternatives.
-In our experiments, XTREE reduced
- the expected values of defect counts and  runtimes of software projects to    
- 45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
+ improve on the predicted values. Hence, we propose the  XTREE planner and  show that
+ it runs better than   three alternative methods. For this paper's    case studies,
+ XTREE reduced
+ the expected values of defects \& runtimes of some software projects to    
+ 45\% \& 61.5\%  (median) and  22\% \& 6\% (best case) of the initial values, respectively.
 \end{abstract}
@@
 
-The contributions of this paper are two-fold. Firstly, we offer
-   XTREE, which has not been previously published.
-   XTREE     works  on table of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
-XTREE returns plans that are changes to feature values in the ``bad'' rows to make them more ``good''.
-Secondly, an evaluation strategy used to comparatively assess planners.
-As  shown below, that strategy less us identify  inferior planners.
-For example, when exploring software runtimes and software defects,
-we show that XTREE is significantly better than the methods proposed in two of our prior publications~\cite{me12c}.
+The contributions of this paper are (1)  
+   XTREE   (which is a new algorithm) and (2) an evaluation strategy for   planners.
+XTREE     works  on tables of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
+In the case studies of this paper (reducing    software runtimes and software defects), our evalautions shows
+ XTREE performing significantly better than   the planners we proposed in  prior   papers~\cite{me12c,krishna15}.
 
@@
 and validity.
+
+To allow for reproducibility, all scripts and data used in this 
+study are available on-line at github.com/ai-se/XTREE.
 
@@
 \ei
-Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
-any data miner, there are two ``clouds'' of data:
-\bi
-\item One cloud is the space of training examples.
-\item Another cloud is the space of test examples.
-\ei
+To answer this question, we note
+that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples.
 For a visualization of these clouds, see \fig{howxy}.
@@
 \caption{Training examples (in gray), test examples (in red) and 
-changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
+changed examples generated by a planner (in green).
+Displayed on the first two components of a PCA analysis of all points. 
+The green (changed) examples are closer to the  training
 date than the original test data
@@
 
-We should mistrust the predictions made by a  model   if it is being applied to examples far outside of its region of competency;
-i.e. to test examples that are too far removed from the training data.   
-To test for ``too far'', we can run a data mining experiment to test how well
+We should mistrust the predictions made by a  model   if it is being applied to examples  that are
+too far away from the
+training cloud.
+To test for ``too far'', we can run a data mining experiment that tests how well
 a model learned from the training data applies to the test data. Such experiments return some performance value.
@@
 Note that predictions  about changes that  fall within the space of the training+test data, will be at least
-as reliable as the performance value found in the data mining experiment.
+as reliable as the performance value found in the above data mining experiment.
 Given this, one thing  can be asserted about predictions on changed examples:
 \bi
-\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
-\ei  
-Accordingly, we should use  {\em trust-increasing} planners. That is, we should use planners that push changde examples {\em towards} the
+\item Predictions for changes that move examples towards/away from the training data can be trusted more/less (respectively).
+\ei  
+Accordingly, we should use  {\em trust-increasing} planners that generate changed examples {\em closer} to the
 training examples.  To see how this works, 
  \fig{howxy} is from the {\em ivy} data
-set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
-changed  examples displaced after applying one of our planning methods.
-To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
-two  components (using PCA). Note that the  the   changed examples
+set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
+changed  examples displaced after applying a plan (in green).
+ Note that the  the   changed examples
 cases  (shown in green)  fall closer to the training cases (shown in gray) than
-the test cases (shown in red).  That is, if we believe
-that the gray data is  suitable for assessing
-the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
-that we believe even more in the predictions
-generated from the displaced data).
-
-(Aside: If the reader is concerned that this 2d projection
-losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
-have confirmed that building defect predictors from this 2d space are just as effective
-as predictors learned using all the defect attributes of \fig{ck}.)
-
+the test cases (shown in red).
+
+In that green region of changed examples, our belief in the value of predictions
+will be as much (or more) as our belief in the value of the predictions in the red region (that
+contains the original test data).
 This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
 paper. Hence,  we can assert that
-predictors learned from these training data have some authority in the regions
-used to generate the plans.
+predictors learned from these training examples have some authority in the regions
+contain the changes examples.
+
+That said, the above comes with some important caveats:
+\bi
+\item 
+We   strongly recommend that predictors are assessed prior to planning. That
+issue is explored further in the \tion{tesd}.
+\item
+Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
+\item
+Where possible, planners should be assessed via some external
+oracle that can accurately assess new examples. For an example of that kind of analysis,
+see  \tion{valid}.
+\ei
 
@@
 
-That said, the above comes with some important caveats:
-\bi
-\item 
-We   strongly recommend that predictors are assessed prior to planning. That
-issue is explored further in the \label{sect:tesd}.
-\item
-Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
-\item
-Where possible, planners should be assessed via some external
-oracle that can accurately assess new examples. For an example of that kind of analysis,
-see the Validity  in \tion{valid}.
-\ei
-
   \section{Test Data}\label{sect:tesd}
@@
   and trained a Random Forest
-  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
+  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$.
 That is, we can be very confident that the predictors from the Siegmund data can assess
@@
 The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
-For such data, quality predictor   is be measured using:
-\bi
-\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
+For such data, quality predictor   is be measured using
+(1) the  probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
 the test data detected
-by the {\em predictor}.
-\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
+by the {\em predictor}; and (2) the 
+probability of false alarm (a.k.a. ``pf''): the percent of non-fault
 classes that are {\em predicted} to be defective.
-\ei 
+
 As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
@@
 for tree generation, etc.
-The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
 For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
@@
 
-	\begin{figure} 
-	\begin{shaded}
-  ~\hrule~
-
- {\bf Figure 7.A: Measuring Variability}
-
-	For  continuous and discrete values,
-the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
-Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
-where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
-Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
-for  $n$ discrete values  at frequency
-$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.
-
-  ~\hrule~
-
-	{\bf Figure 7.B: Measuring distance}
-
-	We use
-		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  \Delta(X_i,Y_i)^2}/\sqrt{F}$
-			(the denominator normalizes the output to $0 \le d \le 1$).
-			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
-			Otherwise, replace any  missing items with values that maximizes the following.
-			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
-			returns 
-			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
-			same,different (respectively).
-
- ~\hrule~
-
-	{\bf Figure 7.C: Top-down Clustering with WHERE}
-
-				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
-		Using this measure, WHERE runs as follows:
-		\begin{enumerate}[leftmargin=3mm]
-			\item Find   two   distance cases,  $X,Y$
-			by picking any case $W$ at random, then setting $X$ to its most
-			distant case, then setting $Y$ to the case most distant from
-			$X$
-			(which requires only $O(2N)$ comparisons).
-			\item Project each case $Z$
-			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
-			\item Split the data at the median $x$ value of all cases.
-			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
-		\end{enumerate}
- In terms of related work,
-	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
-	  at each recursive level to find the PCA principle component. Our method, on the other hand,
-	  performs the same task with only $O(2N)$ distance calculations 	using the 
-	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.
-
-	   ~\hrule~
-
-	{\bf Figure 7.D: Top-down division with Decision Trees}
-
-	Find a split in the values of  independent features that most reduces the variability
-  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
-   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.
-
-	 ~\hrule~
-
-	 	{\bf Figure 7.E: Finding the most informative rows}
-
-Discretize all numeric features using the Fayyad-Iranni discretizer~\cite{FayIra93Multi}
-(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
-Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
-and 
-let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
-Cull the the features as per Papakroni~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
-where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).
-
-~\hrule~
-
-	  \end{shaded}
-		\caption{Some algorithms used in this paper.}\label{fig:where}
-		\end{figure}
-
 \section{Four Planning Methods}\label{sect:planners}
+
 In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
@@
 this section described XTREE (which we call Method4) against three  
-alternatives: 
-  \bi
-  \item  
-  Method1, Method2 and Method3 use the   top-down
+alternatives.
+  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
+  The other methods use the   top-down
 	bi-clustering method described in \fig{where}.C  which recursively divides the
 	data in two  using a dimension that captures the greatest variability in the data. 
-	\item
- Method4 uses the decision tree learner of \fig{where}.D. 
-  \ei
-  Note that all our methods have the two properties proposed in \tion{prelim}:
+	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}.
+
+ \subsection{  Method Properties}
+  Note that all   methods have the  properties proposed in \tion{prelim}:
   \bi
@@
   will be given plans that are specialized  to their particulars;
-  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
-  i.e. they change changing examples such that they move {\em closer} to the training data.
-  \ei
+  \item They  are {\em trust-increasing};
+  i.e. they change  examples such that they move {\em closer} to the training data.
+  \ei
+
+ \subsection{  Methods}
 
@@
 
-\subsection{Method1 = CD =   Centroid Deltas}
-
- {\em Summary:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and
+\subsubsection{Method1= CD=   Centroid Deltas}
+
+ {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and
 where you want to be  (which we will call $C_j$).
 
- {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
-  reporting the delta between the cluster centroids. 
+{\em Assumption1: } Method1 assumes that large data sets can be adequately represented by a few dozen (or so) 
+centroids.
+
+ {\em Details1:} Method1 clusters project data by reflecting   on the independent variables, then
+  reports the delta between the cluster centroids. 
   After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
@@
 
-\subsection{Method2 =  CD+FS = \\ Method1  + Feature Selection }
-
-{\em Summary:} Method2 works line Method1 but now the   plans  only
+\subsubsection{Method2=CD+FS=Method1+Feature Selection }
+
+{\em Summary2:} Method2 works line Method1 but now the   plans  only
 mention the $\beta=33\%$ most informative features. Hence, Method2's plans are simpler.
 
-{\em Details:} In most SE applications, not all features contribute equally to a problem.  Hence, one problem
-with the above method is that its plans comment on all features, even if those values are due
-to noise or other effects not connected to better performance.
-
-This is a standard issue in data mining, with a standard solution.
-A common result is that the signal in a table of data is mostly contained in a handful of features~\cite{hall03,kohavi97}.
+{\em Assumption2: } Method2 assumes that, when reasoning about centroids, we can
+just use   features
+that   best distinguish    centroids; i.e. whose values appear in just a few centroids.
+
+{\em Details2:} A common result is that the signal in a table of data is mostly contained in a handful of features~\cite{hall03,kohavi97}.
 Papakroni~\cite{papa13} has tested for this effect in the Jureczko data sets.
@@
 
-Based on the above, it might be possible to simplify Method1's plans by pruning back the features used to make
-the plans. Following on from Papakroni, our Method2 returns plans
+Based on the above, it might be possible to simplify the plans found by Method1  by pruning back the features in those
+ plans. Following on from Papakroni, our Method2 returns plans
 containing just the top $\beta=33\%$ most informative features. Here, ``informative'' means
@@
 
-As an historical note,  Method\#2 was previously proposed in one of our prior publications~\cite{me12c}.
-
-\subsection{Method3 = BIC = \\  Method2 + Best-in-cluster}\label{sect:BIC}
-
-{\em Summary:} Method3 is like Method2, but it uses more knowledge about the training data.
-
-{\em Details:}
-One problem with Method2 is that it makes little use of the detailed knowledge
-of the examples that were squashed together to form the centroids. 
-Hence,
-in Method3, we collect more information about each cluster.
-
+ \input{tex/algos}
+\subsubsection{Method3= BIC=   Method2 + Best-in-cluster}\label{sect:BIC}
+
+{\em Summary3:} Method3 is like Method2, but it uses more knowledge about the training data.
+
+{\em Assumption3: } Method3 assumes that there exists ``gradients'' between and
+within clusters which, if used, will better guide us to finding better plans.
+
+{\em Details3:}
 Method3 summarizes clusters into {\em two} examples: (1)~the centroid $C_x$ found in Method1 and
@@
 
-Next,  each centroid is connected by a line to its nearest centroid.
- In that line, we say $C_i,C_j$ are the centroids associated with worst/best centroid performance scores.
-
+Method3 connects  each centroid to a nearest neighbor
+by {\em gradient}.
+Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
 For each test instance, Method3 
-find the nearest line; selects the  best end $C_j$ of that line;  then extracts $B_j$ (which is the
+find the nearest gradient, the runs up to the top  best end $C_j$, then extracts $B_j$ (which is the
  best-in-cluster associated with  $C_j$).
@@
 and $B_j$.
-Note that, as with Method2, this method's plans only mentions the  $\beta=33\%$ most informative features. 
+
 %\input{tex/pseudo-code-1.tex}
@@
 
-\subsection{Method4 =  XTREE =\\ Deltas in Decision Tree Branches}
-
-{\em Summary:} Method4 builds a decision tree,  then generates
+\subsubsection{Method4=XTREE=  Deltas in Decision   Branches}
+
+{\em Summary4:} Method4 builds a decision tree,  then generates
 plans from the difference between two branches:
@@
 
-{\em Details:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
+{\em Assumption4:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
 the algorithm used for the initial clustering. Unsupervised methods like WHERE
-execute without knowledge of the target class.
-{\em Supervised} methods, on the other hand, are allowed to reflect on more knowledge including
-the values in the target class. In theory, this extra knowledge can lead to better data mining.
-Accordingly, XTREE uses a supervised method to generate its plans.
-Specifically, it  uses the decision tree algorithm of \fig{where}.D to divide the data.
+execute without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.
+
+{\em Details4:} 
+XTREE uses a supervised   decision tree algorithm of \fig{where}.D to divide the data.
 Next, XTREE builds plans from the branches of the decision trees. It does so by asking three questions:
@@
 \ee
-The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{XTREE_bare}.
+The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.
 
@@
 
-\begin{figure}[!b]
-\small 
+This section describes an experimental design (and results) for evaluating the above four methods. 
+\subsection{Experimental Design}
+
+\subsubsection{A Strategy for Evaluating Planners}
+
+ Here is our experimental design:
+
+ \begin{figure}[!h]
+{\small 
 \[
@@
 \end{array} \right. 
-\]
-
-\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
+\]}
+ \caption{Experimental .}\label{fig:design}
  \end{figure}
 
-This section describes an experimental design (and results) for evaluating the above four methods.
-
-\subsection{A Strategy for Evaluation Planners}
-
-\subsubsection{Collecting Data}
-
-\fig{work} shows our experimental design.  Initially, we divide the
+As shown in \fig{design}, we divide the
 project data  into two disjoint sets {\em train} and {\em test}
@@
 then we cannot trust them to comment on our plans. Accordingly,
-if that performance is unsatisfactory, we abort. Recall from the above discussion that this step indicated
+if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
 we should not use some of the  Jureczko data.
@@
 \item If $R= 1$, this means  ``no change from baseline''; 
-\item If $R < 1$, this indicates some reduction to the baseline;
+\item If $R < 1$, this indicates ``some reduction to the baseline'';
 \item If $R > 1$, this indicates ``optimization failure''.
 \ei
-Also, to measure succinctness, we count the number of features in the plan.
-
-\subsubsection{Studying Repeated Samples}
-Our methods contain some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
+ \subsubsection{Statistical Methods}
+ Our methods use some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
   XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
-  (see \fig{XTREE_bare}). Hence, we report the $R$ values seen in 40 repeated runs
+  (see \fig{xtree_bare}). Hence, we report the $R$ values seen in 40 repeated runs
   (with different random number seeds).
@@
 
-Our results are presented by the line diagrams like those shown on the right-hand-side of the following example table.
-The black dot shows the median $R$ value and the horizontal likes stretch over the inter-quartile
-range (hereafter, IQR) that is the space from the 25th percentile value to
-the 75th percentile value.
-
-\begin{center}
-
-{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
-\arrayrulecolor{lightgray}
-\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
-1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
-2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
-2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
-3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
-\end{center}
-
-In this example, the rows are  sorted on the median values of each method. Note that all the methods
-have   $R<1$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
-while XTREE achieved the greatest reduction (down to 49\% of the original value).
-
-The above table has a  left-hand-side  {\bf Rank} column. This column reports if
-the values for each method are statistically different and are more than trivially different. 
-In this example, table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.
-
-  \subsubsection{Statistical Methods}
-  We compute {\bf Rank} using a Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}. 
+  To rank our methods using the results from these 40
+  repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}. 
 In that test, using the median values of each method,  
@@
 
-\subsection{Results}
-
-\begin{figure}[t]
-\begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
-\label{fig:delta_ivy}
-\end{minipage}
-\begin{minipage}{0.5\textwidth}
-\includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
-\label{fig:delta_apache}
-\end{minipage}
+\subsubsection{Report Format}
+
+Our results are presented by the line diagrams like those shown on the right-hand-side of the following example table.
+The black dot shows the median $R$ value and the horizontal likes stretch over the inter-quartile
+range (hereafter, IQR) that is the space from the 25th percentile value to
+the 75th percentile value.
+
+\begin{center}
+
+{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
+\arrayrulecolor{lightgray}
+\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
+1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
+2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
+2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
+3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
+\end{center}
+
+In this example table, the rows are  sorted on the median values of each method. Note that all the methods
+have   $R<1$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
+while XTREE achieved the greatest reduction (down to 49\% of the original value).
+
+The above eample table has a  left-hand-side  {\bf Rank} column, computed using the
+Scott-Knott test described above. This column reports if
+the values for each method are statistically different and are more than trivially different. 
+In this example table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.
+
+\subsubsection{Other Details}
+
+\fig{jur} and \fig{conf1} show the effectiveness of our methods seen in 40 repeats with each data set.
+In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
+Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
+\bi
+\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
+\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
+ all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
+occurred in the {\em train} phase of \fig{design}.
+\ei
+\newpage\subsection{Experimental Results}
+
+% \begin{figure}[t]
+% \begin{minipage}{0.5\textwidth}
+% \includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
+% \label{fig:delta_ivy}
+% \end{minipage}
+% \begin{minipage}{0.5\textwidth}
+% \includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
+% \label{fig:delta_apache}
+% \end{minipage}
+% \caption{Performance Comparison showing the number of times a certain feature requires to be changed.
+X-axis lists the various features that make up a data set.
+The y-axis contains the frequency of change in \%.}\label{fig:changed}
+% \end{figure}
+
+\input{tex/results_jur.tex}
+\input{tex/results_conf.tex}
+
+\begin{figure*}[bp]
+\centering
+\includegraphics[width=\linewidth]{figs/Deltas-both.eps}
 \caption{Performance Comparison showing the number of times a certain feature requires to be changed.
@@
 The y-axis contains the frequency of change in \%.}\label{fig:changed}
-\end{figure}
-
-\input{tex/results_jur.tex}
-\input{tex/results_conf.tex}
+\end{figure*}
 
@@
 \ei 
-
-XTREE performs well in terms of  effectiveness. 
-\fig{jur} and \fig{conf} show the effectiveness of our methods seen in 40 repeats with each data set.
-In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively.
-Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
-\bi
-\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
-\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
- all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
-occurred in the {\em train} phase of \fig{work}.
-\ei
-The results show planning can be very effective-- some of the expected value of the reductions are very large:
+Measured in terms of effectiveness,
+some data sets were harder to optimize that others.  SQL (in \fig{conf1}) defied all 
+our methods for reducing runtimes. Also, BDBJ (in \fig{conf1}) was hard to optimize
+for all methods except XTREE.  However, in other data sets,
+large reductions were observed:
 \bi
@@
 \item
-Down to 22\% of the original baseline in X264  of \fig{conf};
-\ei
-In all these results,  with one exception, XTREE always is the top-ranked method and with a median improvement at least 10\% lower (i.e.
-better) than the next ranked method.
-The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
-so XTREE can only reduce the expected values of the runtimes by 1\%.
-
-From these results, measured in terms of effectiveness, we recommend supervised-based division of the data (as used in XTREE) over unsupervised clustering
-(as used in the other methods). As to the very simple methods (CD and CD+FSS), these were proposed in prior publications
-on this work~\cite{me12c}. We now   observe  that that prior work was poorly evaluated so the recommendations
-of that prior paper  must now be rejected.
-
-As to succinctness , that is discussed in the next section.
+Down to 6\% of the original baseline in BDBD  of \fig{conf1};
+\ei
+Overall, XTREE was most effective. It was always  the top-ranked method and (with 
+the exception of SQL), had significant reductions in the median performance
+values: median improvement of least 10\% lower (i.e. better) than the next ranked method
 
@@
 
+\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
+The left-hand-side plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the right-hand-side
+  shows a Siegmund data set ({\em BDBJ}).
+
+  In these plots, the {\em more} succinct a planning method, the {\em less} percent
+  of the runs
+  where 
+  it   recommends changing a particular feature (i.e. the vertical bars in that plot
+  are {\em lower}). For example,
+ Method1 (CD) was the least succinct  since it  wanted to change all features 
+ (observe the change frequencies as 
+high as 100\% for all features). 
+Method1's policy of ``change everything''  might be acceptable
+if this approach lead 
+to the most effective changes. However, looking at 
+\fig{jur} and  \fig{conf1}, there is no evidence for this.
+
+An  interesting feature of \fig{changed} was that fewer things 
+were changed in the config data sets {\em BDBJ} than in the defect data set
+{\em lucene}. 
+In turns out that this holds true across nearly all our data sets.
+\fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
+there are fewer changed features in the config data than in the defect prediction data. 
+One explanation for that is the nature of the features: the
+config data has binary independent features while the defect data sets has continuous
+features. When exploring these different data types, it is possible to find more
+``gentle slopes'' the lead to small changes
+in the continuous features and more ``sharp cliffs'' that lead to major change in the discrete
+data. Hence, our planners make fewer larger changes in the config data.
+offer
+
+Looing across the columns of \fig{types}, we see that Method3 (BIC) is the most
+succinct for data sets with deiscrete features (
+
 \input{tex/deltas_avg.tex}
 
-\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
-The top plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the bottom
-plot shows a Siegmund data set ({\em BDBJ}).
+Method2 (CD+FSS) was more succinct than Method1 
+since the feature selection mechanism, removed most of
+the feature (evidence: the blue bars for CD+FSS appear least often in both plots). One
+repeated pattern in these results was that 
+when CD+FSS decided that a feature needs to be changed, it always changed it in all repoeats (see how the blue bars for CD+FSS are usually at 100\%). We conjecture that this is a side-effect
+of feature selection-- if you discard most of the controllables, then you always have to change the
+remaining.
+
+In terms of succinctness,  Method3 (BIC)  had different  nearly as unsuccinct as Method1 (see in those results
+that most features have black bars that extend up beyonf 50\% and often to near 100\%). 
+Method3 was more succinct for {\em BDBJ}
 
@@
 
+\subsection{Discussion}
+
+These results let us comment on the assumptions of the various planning methods:
+\bi
+\item Assumption1: 
+\item Assumption2: 
+\item Assumption3:
+\item Assumption4:
+In all these results,  with one exception,  The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
+so XTREE can only reduce the expected values of the runtimes by 1\%.
+\ei
+
 \section{Validity}\label{sect:valid}
@@
 \section{Related work}
+
+mockus
+
+planning
+
+harman and change
+
+cbr
+
 results suggest that, when clustering on independent variables of software project data,

### 2015-08-24 tex/algos.tex

	\begin{figure} 
	\begin{shaded}
  ~\hrule~

 {\bf Figure 7.A: Measuring Variability}

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values  at frequency
$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

  ~\hrule~

	{\bf Figure 7.B: Measuring distance}

	We use
		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  w_i\Delta(X_i,Y_i)^2}$. Here, $w_i$ is a weight term for each feature
		 (usually set to 1).
			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
			Otherwise, replace any  missing items with values that maximizes the following.
			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
			returns 
			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
			same,different (respectively).

 ~\hrule~

	{\bf Figure 7.C: Top-down Clustering with WHERE}

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

	   ~\hrule~

	{\bf Figure 7.D: Top-down division with Decision Trees}

	Find a split in the values of  independent features that most reduces the variability
  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.

	 ~\hrule~

	 	{\bf Figure 7.E: Finding the most informative rows}

Discretize all numeric features using the Fayyad-Iranni discretizer~\cite{FayIra93Multi}
(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
and 
let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
Cull the the features as per Papakroni~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).

~\hrule~

	  \end{shaded}
		\caption{Some algorithms used in this paper.}\label{fig:where}
		\end{figure}

### 2015-08-24 tex/prediction_stats.tex

--- 2015-08-14 tex/prediction_stats.tex
@@
 \begin{figure*}[!t]
-	\scriptsize
+	\small
 	\begin{center}
@@
 				pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
-				55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
-				65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
+				\rowcolor{celadon}55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
+				\rowcolor{celadon}	65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
 				49 & 31 &   & 56 & 37 &   & 5 & 6\\
-				49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
+				\rowcolor{celadon}	49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
 				45 & 19 &   & 47 & 15 &   & 2 & -4\\
@@
 				56 & 25 &   & 60 & 25 & y & 4 & 0\\
-				56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
-				30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
-				32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
-				38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
+				\rowcolor{celadon}	56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
+			\rowcolor{lavenderpink}	30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
+				\rowcolor{lavenderpink}32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
+				\rowcolor{lavenderpink}38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
 				\hline 
@@
 		(and note that, after tuning, there are more ``good'' than before).
-		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
-		Data   marked with ``$\times$'' are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
+	Results  	\colorbox{celadon}{ marked with ``$\star$''} show large improvements in performance, after tuning
+		(lower {\em pf} or higher {\em pd}).
+		The data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
 	}\label{fig:j}

### 2015-08-25 Manuscript.tex
Updates from ShareLaTeX

@@
 XTREE     works  on tables of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
-In the case studies of this paper (reducing    software runtimes and software defects), our evalautions shows
- XTREE performing significantly better than   the planners we proposed in  prior   papers~\cite{me12c,krishna15}.
-
-The rest of this paper is structured as follows.
-After describing our test data,
-this paper  walks the reader through four methods that start
-with a standard clustering algorithm, the  progresses on to XTREE. The second
-half of the paper then presents an extensive evaluation of these methods using 
-the Jureczko  and Siegmund data sets. This is followed by notes on related work
+In the case studies of this paper  our evaluations shows
+ XTREE performing significantly ``better'' than   the planners we proposed in  prior   papers~\cite{me12c,krishna15},
+ where ``better'' means:
+\bi
+\item {\em Effective:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  in expected values of the results should  be  observed. 
+\item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
+Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
+\item {\em Surprising}; i.e. we want the planners to (occasionally) tell us things we do not know already (otherwise,
+there would be no point to using the planner).
+\ei 
+The rest of this paper is describes our data, our planners, and the experiment that
+ranks XTREE against alternate approaches.  This is followed by notes on related work
 and validity.
-
-To allow for reproducibility, all scripts and data used in this 
+Note that, to allow for reproducibility, all scripts and data used in this 
 study are available on-line at github.com/ai-se/XTREE.
@@
 and defect predictions (better median results,
-lower variance in the predictions). This paper offers yet another locality result:
+lower variance in the predictions). \tion{surprise} of this  paper offers yet another locality result:
 \bi
@@
 After making those
-changes, we may have a new example that has never seen before. Therefore, it must be asked:
-\bi
-\item
-{\em Can we trust the predictions made on such new examples?}
-\ei
+changes, we may have a new example that has never seen before. Therefore, it must be asked
+{\em ``can we trust the predictions made on such new examples?''}
+
 To answer this question, we note
@@
 \begin{figure}[!t]
-\includegraphics[width=1\linewidth]{figs/BDBC.eps}
+\includegraphics[width=0.9\linewidth]{figs/BDBC.eps}
 \caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
@@
 
- \subsection{  Method Properties}
-  Note that all   methods have the  properties proposed in \tion{prelim}:
-  \bi
-  \item They are {\em local learners}; i.e. different test examples
-  will be given plans that are specialized  to their particulars;
-  \item They  are {\em trust-increasing};
+  Note that all   methods have the  properties proposed in \tion{prelim}.
+  That is, they are {\em local learners}; i.e. different test examples
+  will be given plans that are specialized  to their particulars. Also,
+  they  are {\em trust-increasing}; 
   i.e. they change  examples such that they move {\em closer} to the training data.
-  \ei
 
@@
   XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
-  (see \fig{xtree_bare}). Hence, we report the $R$ values seen in 40 repeated runs
+  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
   (with different random number seeds).
@@
 \ei
-\newpage\subsection{Experimental Results}
+ \subsection{Experimental Results}
+
+Recall from our introduction that we are assessing planners on three criteria:
+{\em effectiveness}, which is how much they reduce the expected value of the changed examples;
+{\em succinctness}, which is how many things we need to change to achieve a plan;
+and {\em surprise}, which is how different are the plans from standard truisms.
 
@@
 
-\begin{figure*}[bp]
+\begin{figure*}[!t]
 \centering
 \includegraphics[width=\linewidth]{figs/Deltas-both.eps}
-\caption{Performance Comparison showing the number of times a certain feature requires to be changed.
-X-axis lists the various features that make up a data set.
-The y-axis contains the frequency of change in \%.}\label{fig:changed}
+\caption{Percent frequency for how often certain feature was changed bu a plan.}\label{fig:changed}
 \end{figure*}
@@
 
-For a planning system designed for business users, we seek two properties:
-\bi
-\item {\em Effectiveness:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  should  be  observed. 
-\item {\em Succinctness:} given two plans with the same effectiveness, we prefer the shorter one.
-Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
-\ei 
 Measured in terms of effectiveness,
@@
 the exception of SQL), had significant reductions in the median performance
-values: median improvement of least 10\% lower (i.e. better) than the next ranked method
+values: median improvement of least 10\% lower (i.e. better) than the next ranked method.
 
@@
 One explanation for that is the nature of the features: the
-config data has binary independent features while the defect data sets has continuous
-features. When exploring these different data types, it is possible to find more
-``gentle slopes'' the lead to small changes
-in the continuous features and more ``sharp cliffs'' that lead to major change in the discrete
-data. Hence, our planners make fewer larger changes in the config data.
-offer
-
-Looing across the columns of \fig{types}, we see that Method3 (BIC) is the most
-succinct for data sets with deiscrete features (
+defect data sets has continuous
+features while the 
+config data has binary independent features  (where so setting was turned ``on'' or ``off''). When exploring these different data types, it is possible to find more
+\bi
+\item
+``Fentle slopes'' the lead to small changes
+in  continuous space;
+\item
+ ``Sharp cliffs'' that lead to major change in the discrete
+space. 
+\ei
+Hence, our planners can make fewer larger changes in the config data.
+ For example, Method3 (BIC) and Method1 (CD) make far fewer changes in   data sets with discrete features (the Siegmund data) than in data with continuous features (the Jureczko data).
+
+ As to the method that lead to best effectiveness in \fig{jur} and \fig{conf1}, we note from \fig{types}
+ that XTREE
+ has a consistent behavior in both the discrete and continuous data sets.  Specifically,  in all our data sets,
+ XTREE changes 
+ usually changes around a fifth of the features.
 
@@
 
-Method2 (CD+FSS) was more succinct than Method1 
-since the feature selection mechanism, removed most of
-the feature (evidence: the blue bars for CD+FSS appear least often in both plots). One
-repeated pattern in these results was that 
-when CD+FSS decided that a feature needs to be changed, it always changed it in all repoeats (see how the blue bars for CD+FSS are usually at 100\%). We conjecture that this is a side-effect
-of feature selection-- if you discard most of the controllables, then you always have to change the
-remaining.
-
-In terms of succinctness,  Method3 (BIC)  had different  nearly as unsuccinct as Method1 (see in those results
-that most features have black bars that extend up beyonf 50\% and often to near 100\%). 
-Method3 was more succinct for {\em BDBJ}
-
-As predicted in the above discussion on Method 1 (CD), this method always proposes changes  to most features. Method3 (BIC) often proposes changes to many features, which 
-would make its plan tedious to implement. 
-Method 2 (CD+FS)
-proposes less, due to its use of feature selection and, in the case of {\em BDBJ}, it fixates on changing just one feature (which
-probably explains why CD+FS performs so poorly for {\em BDBJ} in \fig{conf}). Finally, Method4 (XTREE) rarely fixates
-on changing one feature (only on {\em LUCENE}'s lines of code feature) and seems to prefer making many small changes,
-all over the space of features.
-
-\fig{changed} reports the changes just for   {\em lucene}) and {\em BDBJ}. \fig{types} shows the mean changes for features found by our
-four methods (here, mean change is the average height of the histogram bars seen in, say, \fig{changed}). Overall, Method1 (CD) and
-Method3 (BIC) make the most changes while XTREE and CD+FS make the fewest.
-
-\subsection{Discussion}
-
-These results let us comment on the assumptions of the various planning methods:
-\bi
-\item Assumption1: 
-\item Assumption2: 
-\item Assumption3:
-\item Assumption4:
-In all these results,  with one exception,  The  exception is SQL which (as shown in \fig{conf}), seems to defy all methods for optimization
-so XTREE can only reduce the expected values of the runtimes by 1\%.
-\ei
-
-\section{Validity}\label{sect:valid}
-
+\subsubsection{Surprising Results}\label{sect:surprise}
+
+If a planner only ever reports conclusions that were already known, then that planner offers
+little value-added over ``just use established wisdom''. Accordingly, we studied our results
+for plans that were somewhat counter-intuitive.
+
+Such a plan was seen in the {\em lucene} results of \fig{changed}. Recall the standard advice for
+OO systems: build classes that are internally cohesive with low coupling to other parts of the system.  
+\subsubsection{Summary}
+In summary, we  have two reasons to endorse XTREE. Firstly, it is usually far more effective than other
+methods. Secondly, even when we jump from continuous to discrete features, XTREE's plans consistently
+use just a small portion of the features.
+
+In terms of the assumptions that motivated our different methods, we note that 
+{\em Assumption4} seems most powerful; i.e. for the purposes of planning,
+it is better to reflect on the target class.
+
+\section{Threats to Validity}\label{sect:valid}
+
+As with any empirical study, biases can affect the final results. Therefore, any
+conclusions made from this work must be considered with the following issues in
+mind.
+
+\subsection{ Learner Bias}
+For building the defect predictors in this study, we elected
+to use  Random Forest  and Random Forest Regressors .
+We chose this approach,  based on its reputation for having the better  performance of 
+21 other learners for defect prediction~\cite{lessmann}.
+Data mining is a
+large and active field and any single study can only use a small
+subset of the known classification algorithms.
+
+That said, we have taken care to document in this paper the decisions made by engineering
+judgement that could effect our conclusions. The above code used a set of variables which future
+work should vary in order to test the internal validity of our conclusions:
+\bi
+\item All our planners divide data into groups of size $\alpha=\sqrt{N}$
+\item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);
+\item Method4 (XTREE) assumed that another sibling was useful if  had 
+ a  score less than $\gamma=0.5$ times the mean score of the current leaf.
+ \ei
+
+\subsection{  Sampling bias} threatens any data mining experiment; i.e., what matters
+there may not be true here. For example, the data sets used here comes from two sources
+(Seigmund et al. and Jureczko et al.) and any biases in their selection procedures
+threaten the validity of these results. 
+That said,
+the best we can do is define our methods and publicize our data and code so that other researchers can
+try to repeat our results and, perhaps, point out a previously unknown bias
+in our analysis. Hopefully, other researchers will emulate our methods in
+order to repeat, refute, or improve our results.
+
+\subsection{  Evaluation bias} Another threat to validity of this work is our use
+of predictors learned on the training data to assess the impact of our planners.
+This issues was discussed in detail in \tion{trust}.
+
+To those notes, we add a few more details. If possible, planners should be assessed via some external oracle that can accurately assess new examples. For example, in search-based software engineering,
+examples can be assigned objective scores via  some model. In this approach, a changed example can be assessed by
+generating actual objective scores from the model.
+
+\input{tex/results_models.tex}
+
+For example, the COCOMO model from University of 
+Southern California estimates software development time using a combination of industrial data and some domain expertise
+from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
+as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
+inputs. This generated an development time estimate for each example which we passed to our four methods.
+40 times, we let those methods propose changes to those projects. 
+For assessment purposes, the changes projects were then feed back to the COCOMO
+oracle.
+
+As shown in \fig{coc}, those results endorsed the conclusions of the rest
+of this paper; i.e. that XTREE's supervised methods are best for generating plans on how to change
+example projects.
+
+\newpage
 \section{Related work}
 
-mockus
-
-planning
-
-harman and change
-
-cbr
-
-results suggest that, when clustering on independent variables of software project data,
-	we are free to select from a wide range of 
-	clustering methods.  Ganesan~\cite{div14} explored 
-	different clustering methods for SE data using   effort and defect data from
-	the PROMISE repository~\cite{promiserepo}.
-	That study explored
-	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
-	below.
-	Clusters were assessed via the performance of prediction 
-	learned from the clusters by Random Forests (for defect prediction)
-	and M5prime or linear regression (for effect estimation).  Ganesan found
-	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
-	Accordingly, for clustering, we select a clustering method that generates a large
-	number of small cluster
-	($\sqrt{N}$ clusters
-	from $N$ examples) that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e.
-contains signals not associated with the target variable.
-	ats to Validity
-
-\input{tex/results_models.tex}
-
-\section{Related Work}
-dashboards
-
-strucftureal learners (bayes)
-
-Many data mining methods offer a succinct summary of project data
-e.g. see the Bayes nets
-favored by Dejaeger et al.~\cite{Dejaeger13}; or   the regression models
-learned by Nagappan~\cite{Nagappan05}.
-Another  technique is to offer a dashboard that visualizes project data.
-For example, the dashboards
-provided by the Hackystat team~\cite{Johnson09} presents extensive and detailed
-summaries of programmer behavior.
-
-Some users do not like being told what to do... they prefer to have the current world explained to them 
--- but those dashboards typically do not learn threshold points
-(that report where action is needed) nor do they offer specific advice on what actions to take.
+\subsection{Planning in AI}
+
+The XTREE planner is somewhat different to the logical-based planners explored by 
+classical AI. 
+That kind of planning is a logical procedure~\cite{Fikes1971}
+that seeks an ordering on {\em operators} to take some domain
+{\em state} from a {\em start} state to a {\em  goal} state.
+This classical logical approach is known to suffer from
+computational bottlenecks~\cite{Bylander1994} while tools like XTREE will scale to any domains
+that can generate decision trees.
+
+\subsection{Evaluating Changes}
+
+Some organizations have the resources to 
+run repeated trials to assess  different treatments.
+For example, in one remarkable recent study, Bente et al. report results
+were the same for specification that was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
+to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
+then check in 12 months time
+ which teams have fewer defects than the others.
+
+Very few industrial or research groups have access
+to the kinds of resources needed for this kind study  (evidence: in the six years since the
+publication of that work, we know of only one   similar study to Bente et al.). Also, given the
+diversity of modern software projects, it might be unreasonable to demand that all
+proposed changes for all projects are always evaluated by something like the Bente et al. study.
+Hence, 
+this paper has used data miners to build an oracle that can assess changed examples. The advantage
+of this approach is that it required far less resources to assess the effectiveness of proposed
+changes to a project.
+
+\subsection{Search-based SE}
+
+Another way way to propose changes to software artifacts
+is   via some search-based method~\cite{Harman2009,Harman2011}. Such SBSE methods are   evolutionary programs that 
+make
+ extensive changes to  some initial sample of project data
+ (perhaps 
+100s to 100,000s of mutations). Each of these mutations
+is reassessed using some domain model.
+Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
+deb14,Cui2005a,zhang07:TEC}.
+
+One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
+of \tion{trust}, these methods may not be {\em trust-increasing} those algorithms make no attempt
+to mutate new examples away from the kinds of data used to commission the model (in which case, we would
+start doubting the model's output).
+
+Another issue with standard search-based SE methods is that they require ready access to 
+trustworthy domain model that can offer an assessment
+of newly generated examples. While some domains have such models (e.g. see the COCOMO effort estimation model
+used in the last section), our experience is that many others do not.  For example, 
+consider software defect prediction and all the intricate issues that may lead to defects in a product. A model that includes {\em all} those
+potential issues would be very large and complex. Further,
+the empirical data required to validate any/all parts
+of that model can be hard to find.
+
+What we would recommend is a two-pronged policy.
+In domains with ready access to trusted models, we recommend
+the kinds of tools that are widely used in the search-based
+software engineering community such as GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
+deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like XTREE.
 
@@
 
+\balance
 \bibliography{References}
@@
 The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}.
-For full algorithmic details, refer to \fig{XTREE_bare}.
+For full algorithmic details, refer to \fig{xtrees_bare}.
 
@@
 The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity.
-We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{XTREE_bare}).
+We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).
 

### 2015-08-25 tex/dtree.tex
Updates from ShareLaTeX

@@
-\begin{figure}[!t] 
+\begin{figure}[!b] 
 	\begin{shaded}
@@
 
-Using the training data,  divide the data using the decision tree of algorithm of \fig{where}.D.
+Using the training data,  divide the data using the decision tree of algorithm of \fig{where}.D into groups of
+size $\alpha=\sqrt{N}$.
 
@@
 		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\gamma=0.5$ times the mean score of {\em current}. 
-		 \bi
-		 \item 
-		   If none found, then repeat for $lvl += 1$
-		 \item
-		    Return no plan if the new $lvl$ is above the root.
-		 \ei
+		   If none found, then repeat for $lvl += 1$. Also,
+		    return no plan if the new $lvl$ is above the root. 
 		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
@@
 		\end{shaded}
-		\caption{XTREES}
-		\label{fig:xtrees_bare}
+		\caption{XTREE}		\label{fig:xtrees_bare}
 	\end{figure}

### 2015-08-25 tex/results_models.tex
Updates from ShareLaTeX

@@
-\begin{figure}[htbp]
+\begin{figure}[!b]
 
@@
 \textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
-  1 &        XTREES &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
+  1 &        XTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
 \hline  2 &      CD+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
@@
 \caption{Results: Methods 1,2,3,4 applied to some ground-truth data.
-Values collected from   20 repeated runs of each method with different random seeds.
-}\label{fig:conf}
+Values collected from   40 repeated runs of each method with different random seeds.
+Results show ratios of (1) sum of development effort (in the changes examples) to (2)
+the of the development effort in the original test data.
+}\label{fig:coc}
 \end{figure}

### 2015-08-25 Manuscript.tex
Updates from ShareLaTeX

@@
 In response to this business-level demands for planners,
-we propose a novel {\em planning} approach called XTREE for learning changes to a software system
+we propose a novel {\em planning} method called XTREE for learning changes to a software system
 such that its performance ``improves'', according to some measure. This paper uses XTREE
@@
 
-The approach discussed here is novel since most data miners used in software analytics
-comment on ``what is'' and not ``what to do''. For example, a clustering
-algorithm might find clumps of similar examples, but does not comment on the minimal
-change(s) required to move from clump to clump.
-
 The contributions of this paper are (1)  
-   XTREE   (which is a new algorithm) and (2) an evaluation strategy for   planners.
-XTREE     works  on tables of data with weighted class values indicating what rows are ``good'' and what rows are ``bad''.
+    the new XTREE    algorithm and (2) an evaluation strategy for   planners.
 In the case studies of this paper  our evaluations shows
- XTREE performing significantly ``better'' than   the planners we proposed in  prior   papers~\cite{me12c,krishna15},
+ XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
  where ``better'' means:
 \bi
-\item {\em Effective:}    i.e.  if  its recommendations  are  applied  then  some  statistically significant  change  in expected values of the results should  be  observed. 
+\item {\em Effective:}    i.e.  if  a plan is   applied  then  some  statistically significant  change  in expected values of the results will  be  observed. 
 \item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
@@
 \section{Preliminaries}\label{sect:prelim}
+
+\subsection{What is a ``Plan''?}
+
+Our planners   use tables of data with independent features and one dependent feature, called the 
+``class''. Classes are weighted such that they indicate  what rows are ``good'' and what rows are ``bad''. Plans  change   a row such that it is more like they 
+will be more similar to ``good'' than ``bad''.
+Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
+  adjust   feature $Z_j$:
+\[
+\forall \delta_j \in \Delta :  Z_j = Z_j + \delta_j
+\]
+For example, to simplify a  large bug-prone  method, our planners might suggest
+to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
+two simpler functions).
+
+Note that we make no assumption that a plan mentions every features
+(so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than
+plan2).
 
@@
  % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
-\caption{Training examples (in gray), test examples (in red) and 
-changed examples generated by a planner (in green).
-Displayed on the first two components of a PCA analysis of all points. 
-The green (changed) examples are closer to the  training
-date than the original test data
-i.e. predictors trained on the training examples and
-applied to the test examples should
-work just as well, or better, on the changed test examples.}\label{fig:howxy}
+\caption{Gray, red, green show (1) training examples, (2) test examples and 
+  (3) tests that have been altered by planners.
+  This figure uses axes generated from the first two components of a PCA analysis of all points. 
+}\label{fig:howxy}
 \end{figure}
@@
 {\em Assumption4:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
-the algorithm used for the initial clustering. Unsupervised methods like WHERE
-execute without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.
+the clustering algorithm (WHERE) that  
+executes without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.
 
@@
 XTREE uses a supervised   decision tree algorithm of \fig{where}.D to divide the data.
-Next, XTREE builds plans from the branches of the decision trees. It does so by asking three questions:
+Next, XTREE builds plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
+That code asks three questions, the last of which returns the plan:
 \be
@@
 \ee
-The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.
 
@@
 \item
-``Fentle slopes'' the lead to small changes
+``Gentle slopes'' the lead to small changes
 in  continuous space;
@@
 
-Such a plan was seen in the {\em lucene} results of \fig{changed}. Recall the standard advice for
-OO systems: build classes that are internally cohesive with low coupling to other parts of the system.  
-\subsubsection{Summary}
-In summary, we  have two reasons to endorse XTREE. Firstly, it is usually far more effective than other
-methods. Secondly, even when we jump from continuous to discrete features, XTREE's plans consistently
-use just a small portion of the features.
-
-In terms of the assumptions that motivated our different methods, we note that 
-{\em Assumption4} seems most powerful; i.e. for the purposes of planning,
-it is better to reflect on the target class.
+Such a surprising plan can be  seen in the {\em lucene} results. Recall the standard advice for
+OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
+to specific projects by checking how often a planner changes the
+ coupling-related features:
+\bi
+\item {\em ca}:   afferent couplings  =  how many other classes use the specific
+				class;
+\item {\em  	ce}:  efferent couplings =  how many other classes is used by the
+				specific class. 
+\item {\em cbm}: coupling between methods =  total number of new/redefined methods
+				to which all the inherited methods are coupled
+\item	{\em cbo}:  coupling between objects = a value that increases when the methods of one
+				class access services of another.
+
+\item {\em ic}:   inheritance coupling =  number of parent classes to which a given
+				class is coupled (includes counts of methods and variables inherited)
+\ei
+For the  {\em lucene} XTREE results  of  \fig{changed},
+the most frequent change was to alter the lines of code in a class (see the tallest
+red historgram in that figure on the {\em loc}, or lines of code). 
+Looking at the logs of our planner, we can see
+that XTREE's proposed changed is to {\em reduce} the size of a class.
+The only way to do that,  while keeping the  same functionality, it is create a network
+of smaller classes that interact to produce that functionality.
+That is, we would need to {\em increase} the coupling of those classes to achieve XTREE's plan.
+
+In theory, increasing coupling between classes complicates and confuses a class design.
+But the {\em lucene} XTREE results  of  \fig{changed} rarely proposes changing   the
+coupling features {\em ca, ce, cbm, ic} (in fact, XTREE never proposes any change to {\em ic}).
+
+The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
+in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
+XTREE is warning against, say, some factory class generating a large
+community of agents, all of the same class, who
+co-ordinate on some task. This is a different issue to the class redesign issue that would
+be triggered by altering {\em loc}}.
+
+In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
+At least for the {\em lucense} data set, we can see advice that recommends {\em increasing coupling}
+to reduce defects.
 
@@
 
-\subsection{  Sampling bias} threatens any data mining experiment; i.e., what matters
+\subsection{  Sampling bias} 
+Sampling bias threatens any data mining experiment; i.e., what matters
 there may not be true here. For example, the data sets used here comes from two sources
@@
 
-\subsection{  Evaluation bias} Another threat to validity of this work is our use
+\subsection{  Evaluation Bias} Another threat to validity of this work is our use
 of predictors learned on the training data to assess the impact of our planners.
@@
 as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
-inputs. This generated an development time estimate for each example which we passed to our four methods.
+inputs. This generated an development time estimate for each example, which we passed to our four methods.
 40 times, we let those methods propose changes to those projects. 
@@
 
-As shown in \fig{coc}, those results endorsed the conclusions of the rest
-of this paper; i.e. that XTREE's supervised methods are best for generating plans on how to change
+Using this approach, it is possible to assess the value of a plan by measuring its
+effectiveness with respect to some ground truth (in this case, the COCOMO model).
+As shown in \fig{coc}, XTREE passes this assessment.
+Those results from the COCOMO model endorsed the conclusions of the rest
+of this paper; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change
 example projects.
 
-\newpage
 \section{Related work}
@@
 
-\section{Conclusion}
-
-works for more than defects and tunrimes
-
-This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
-can (a)~select useful planning methods while (b)~culling clearly
-inferior planning methods. 
-We apply this strategy to XTREE and three alternate planning methods to find:
-\bi
-\item Two methods that are clearly inferior;
-\item One method that proposes complex changes, which work as well as XTREE;
-\item And XTREE, that proposes very simple changes.
-\ei
+\section{Summary}
+
+The planner proposed in this paper propose changes to software project details in order to improve the expected
+value of the performance scores of that part of the project.
+To evaluate these planners,
+data miners can be used to build oracles to assess planners.
+Such planners should be {\em trust-increasing}; i.e. they propose changes that generate
+changed examples that are closer to the training data of the data miner.
+One caveat here is that the evaluations we can make on the planner are only as good as the predictive
+performance of the data miner. Hence, if domain data does not support satisfactory predictors, then
+planning in that domain cannot be evaluated.
+
+Four planners were assessed here for the tasks of reducing defects and runtimes. 
+Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
+paper is that a novel fourth method clearly out-performs the other three
+(measured in terms of {\em effectiveness, succinctness, and surprise}).
+The reason for the superior performance of our new method is clear: when planning how to change
+to different class methods, it is best to use a learner that reflects over those classes
+as it divides the data.
+
 Of course, there are many more methods for generating plans and
@@
 Our hope is that other   researchers apply this strategy  to repeat and/or improve
-and/or refute our results. To that end, all our materials and data sets are available on-line, see
-github.com/ai-se/XTREE.
+and/or refute our results.
 

### 2015-08-26 Manuscript.tex
changing naveen's location
Updates from ShareLaTeX

@@
 
-\newcommand{\quart}[4]{\begin{picture}(100,4)%1
+\newcommand{\quart}[4]{\begin{picture}(80,4)%1
 {\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}
@@
  \affaddr{Bellhops } \\ 
-       \affaddr{Washington D.C.}\\ 
+       \affaddr{Tennessee, USA}\\ 
        {navek91@gmail.com}
@@
  i.e. specific suggestions on  what to change  in order to
- improve on the predicted values. Hence, we propose the  XTREE planner and  show that
- it runs better than   three alternative methods. For this paper's    case studies,
- XTREE reduced
- the expected values of defects \& runtimes of some software projects to    
- 45\% \& 61.5\%  (median) and  22\% \& 6\% (best case) of the initial values, respectively.
+ improve on the predicted values.
+
+ This paper proposes
+ and evaluates the  XTREE planner for software projects. 
+XTREE inputs tables of data with independent features plus a   weighted class
+that indicate how much each row is  ``bad'' or  ``better''. Plans are edits proposed for each 
+row such that the changed row is more likely to be ``better''.  XTREE
+learns those plans by building a decision tree across the data, then reporting
+the differences in the branches from some current branch to another desired branch.
+
+Using data from 11 software projects,  XTREE found   larger reductions that three
+alternate methods.  Those reductions had 
+median size of (55\%,38.5\%) and   largest size of (78\%,94\%) in
+(defects,runtimes), respectively.
 \end{abstract}
@@
  XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
- where ``better'' means:
-\bi
-\item {\em Effective:}    i.e.  if  a plan is   applied  then  some  statistically significant  change  in expected values of the results will  be  observed. 
-\item {\em Succinct:} i.e. given two plans with the same effectiveness, we prefer the shorter one.
-Such shorter plans are easier to explain (less to show) and faster to implement (less to change).
-\item {\em Surprising}; i.e. we want the planners to (occasionally) tell us things we do not know already (otherwise,
-there would be no point to using the planner).
-\ei 
-The rest of this paper is describes our data, our planners, and the experiment that
+ where ``better'' means 
+{\em effective} (plans change the  expected values of the class);
+{\em succinct} (it is inconvenient if  plans   always require changing  everything);
+and {\em surprising} (a planner should sometimes tell us things we do not expect
+since, otherwise,
+there is no value added from  using the planner).
+
+The rest of this paper   describes our data, our planners, and the experiment that
 ranks XTREE against alternate approaches.  This is followed by notes on related work
@@
 Our planners   use tables of data with independent features and one dependent feature, called the 
-``class''. Classes are weighted such that they indicate  what rows are ``good'' and what rows are ``bad''. Plans  change   a row such that it is more like they 
-will be more similar to ``good'' than ``bad''.
+``class''. Classes are weighted such that they indicate  what rows are ``bad'' and what rows are ``better''. Plans  change   a row such that it is more likely to be ``better''. 
 Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
@@
 This paper is about the next step {\em after} prediction. Suppose
-a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
+a business user is presented with  a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
 then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.
@@
 To answer this question, we note
-that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples.
-For a visualization of these clouds, see \fig{howxy}.
-
+that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples
+(for a visualization of these clouds, see \fig{howxy}).
 \begin{figure}[!t]
-  \includegraphics[width=\linewidth]{figs/twodee.eps} 
+  \includegraphics[width=.9\linewidth]{figs/twodee.eps} 
  % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
@@
 \end{figure}
-
 We should mistrust the predictions made by a  model   if it is being applied to examples  that are
@@
 oracle that can accurately assess new examples. For an example of that kind of analysis,
-see  \tion{valid}.
+see  \tion{coc}.
 \ei
@@
 
-\caption{Siegmund data.
-For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
+\caption{Siegmund data.}\label{fig:cpm}
 \end{figure}
@@
 
-In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
-algorithm should be baselined against simpler alternatives. Accordingly,
-this section described XTREE (which we call Method4) against three  
-alternatives.
+This section described XTREE (which we call Method4) and  three  
+alternate methods for learning plans.
   XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
@@
 Method3 connects  each centroid to a nearest neighbor
-by {\em gradient}.
+by a {\em gradient}.
 Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
@@
 
-\section{Experiments}
+\newpage \section{Experiments}
 
@@
 \item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
- all the SMOTE-ing and Random Forst tunings (discussed in \tion{tesd})
+ all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd})
 occurred in the {\em train} phase of \fig{design}.
@@
 
-% \begin{figure}[t]
-% \begin{minipage}{0.5\textwidth}
-% \includegraphics[width=0.9\linewidth]{figs/Deltas-Lucene.eps}
-% \label{fig:delta_ivy}
-% \end{minipage}
-% \begin{minipage}{0.5\textwidth}
-% \includegraphics[width=0.9\linewidth]{figs/Deltas-BDBJ.eps}
-% \label{fig:delta_apache}
-% \end{minipage}
-% \caption{Performance Comparison showing the number of times a certain feature requires to be changed.
-X-axis lists the various features that make up a data set.
-The y-axis contains the frequency of change in \%.}\label{fig:changed}
-% \end{figure}
-
 \input{tex/results_jur.tex}
@@
 \includegraphics[width=\linewidth]{figs/Deltas-both.eps}
-\caption{Percent frequency for how often certain feature was changed bu a plan.}\label{fig:changed}
+\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
 \end{figure*}
@@
 co-ordinate on some task. This is a different issue to the class redesign issue that would
-be triggered by altering {\em loc}}.
+be triggered by altering {\em loc}.
 
@@
 
-\subsection{  Evaluation Bias} Another threat to validity of this work is our use
+\subsection{  Evaluation Bias}\label{sect:coc}
+Another threat to validity of this work is our use
 of predictors learned on the training data to assess the impact of our planners.
@@
 example projects.
+
+\subsection{  Feasibility}\label{sect:feasibility}
+It is important to also explore the general practicality of the plans.
+To this end, we are currently undertaking is a \textit{feasibility study} collaboration with industrial partners to assess the general feasibility of our changes.
+With this study, we would like to understand the nature of changes suggested by XTREE.
+Specifically: are the changes proposed by XTREE small enough to be easily implemented on software projects?
+Also, are the changes large enough to be feasible?
+In other words, we are asking if users can sufficient fine-grained control to make the changes proposed in \fig{changed}.
 
@@
 One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
-of \tion{trust}, these methods may not be {\em trust-increasing} those algorithms make no attempt
+of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
 to mutate new examples away from the kinds of data used to commission the model (in which case, we would
@@
 (measured in terms of {\em effectiveness, succinctness, and surprise}).
-The reason for the superior performance of our new method is clear: when planning how to change
-to different class methods, it is best to use a learner that reflects over those classes
-as it divides the data.
-
-Of course, there are many more methods for generating plans and
-no   one paper can survey them all.
-But the goal of this paper is not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
-to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
-Our hope is that other   researchers apply this strategy  to repeat and/or improve
-and/or refute our results.
+We conjecture that XTREE worked better than the rest since:
+\bi
+\item
+It uses a supervised method to divide the data and;
+\item
+When planning how to move examples to better classes, it is   best to use a learner that reflects over 
+differences between the classes.
+\ei
+
+\section{Questions for Future Work}
+
+XTREE, as used here, sought improvements in a one goal (the   class variable). Does XTREE work for multi-goal reasoning?
+
+ The XTREE algorithm seems quite general to any data of table with rows containing
+weighted classes (so we can distinguish ``bad'' rows from ``better'' ones). 
+Does   XTREE works  on   domains (other than the  defect/runtime data explored here)?
+
+As an example of a domain that might benefit from XTREE, recent results raise doubts about
+the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``really bad smell''
+detector to select the subset of possible refactorings that have the most   potential benefit?
+
+As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory,  XTREE should work on
+any domain where data miners can generate decision trees. Given the current state of the art in Big Data,
+can   XTREE  be applied to  very large data sets?
+
+There are many more methods for generating plans and
+no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta,\gamma$
+parameters that controlled XTREE. Would we get better results if we varied those parameters?
+
+That said, the goal of this paper was not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
+was to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
+Will other  researchers apply this strategy  to repeat and/or improve
+(or even refute) our results?  Perhaps.
 

### 2015-08-27 Manuscript.tex
Updates from ShareLaTeX

@@
  Business users now demand more insightful
- analytics. Instead of just predicting 
- some result, they also need tools that generate ``plans'';
- i.e. specific suggestions on  what to change  in order to
- improve on the predicted values.
-
- This paper proposes
- and evaluates the  XTREE planner for software projects. 
-XTREE inputs tables of data with independent features plus a   weighted class
-that indicate how much each row is  ``bad'' or  ``better''. Plans are edits proposed for each 
-row such that the changed row is more likely to be ``better''.  XTREE
+ analytics; specifically, tools that generate ``plans''--
+  specific suggestions on  what to change  in order to
+ improve  the predicted values.
+
+ This paper proposes  the  XTREE planner for software projects. 
+XTREE inputs tables of data of independent features plus a   weighted class
+  indicating how much each row is  ``bad'' or  ``better''. 
+  Plans are edits to  
+rows such that the changed row is more likely to be ``better''.  XTREE
 learns those plans by building a decision tree across the data, then reporting
@@
 The contributions of this paper are (1)  
-    the new XTREE    algorithm and (2) an evaluation strategy for   planners.
-In the case studies of this paper  our evaluations shows
+    the new XTREE    algorithm and (2) an evaluation strategy that shows
  XTREE performing significantly ``better'' than    planners  proposed in our prior   work~\cite{me12c,krishna15},
@@
 and validity.
-Note that, to allow for reproducibility, all scripts and data used in this 
+To enable reproducibility, all scripts and data used in this 
 study are available on-line at github.com/ai-se/XTREE.
@@
 
-Our planners   use tables of data with independent features and one dependent feature, called the 
-``class''. Classes are weighted such that they indicate  what rows are ``bad'' and what rows are ``better''. Plans  change   a row such that it is more likely to be ``better''. 
+Our planners   use tables of data with independent features and a dependent class
+feature. Classes have weights  that  indicate  what rows are ``bad'' or ``better''. Plans  change   a row such that it is more likely to be ``better''. 
 Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
   adjust   feature $Z_j$:
-\[
-\forall \delta_j \in \Delta :  Z_j = Z_j + \delta_j
-\]
+
+{\small\[
+\forall \delta_j \in \Delta :  Z_j =  
+\begin{cases}
+     Z_j + \delta_j& \text{if $Z_j$ is numeric}\\
+    \delta_j              & \text{otherwise}
+\end{cases}
+\]}
 For example, to simplify a  large bug-prone  method, our planners might suggest
@@
 \begin{figure}[!t]
-  \includegraphics[width=.9\linewidth]{figs/twodee.eps} 
+  \includegraphics[width=\linewidth]{figs/twodee.eps} 
  % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
@@
 \item 
-We   strongly recommend that predictors are assessed prior to planning. That
-issue is explored further in the \tion{tesd}.
+The quality of the prediction depends on the nature of the training data.
+Thus, we strongly recommend that both data set and the predictor be assessed prior to planning.
+This ensures that the predictor's performance is adequate for a data set.
+We tackle this issue in detail in \tion{tesd}.
 \item
@@
 
-   The Jurecko data records number of known defects for each class, where the classes are described in terms of
+   The Jurecko data records number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
   nearly two dozen metrics such as number of children, lines of code, etc. For details on the Jurecko data, see  \fig{ck} and \fig{jd}. 
- For the most part, the methods of this paper treat
-  the Jurecko as a discrete class data set, where {\em defects} are true if the raw defect count is greater than zero.
-  The one exception will be the `best-in-cluster'' method that reflects on the numeric value of the raw defect count (see \tion{BIC}).
+ All the planning methods of this paper reflect on the numeric value of the raw defect counts.
+The predictor considers defects to be a discrete class data set,where defects are \textit{\texttt{TRUE}} if the numeric defect count is greater than zero and \textit{\texttt{FALSE}} otherwise.
 
@@
 of classes and testing on smaller test suites).
-Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately
-defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
-good from bad (since there are so many bad examples).
+Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
+In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
 
@@
 
-As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
+As a preliminary study, we split the Jureczko  data into train and test groups.
 Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
@@
 The ``untuned'' columns of \fig{j} shows those results.
-If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
-then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
-Note that, as might have been expected, none of the \colorbox{lavenderpink}{three bottom rows} of \fig{j} were ``good''.
-
-Fortunately,
-the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
+For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select these data sets.
+With this threshold, however, none of our data sets were be suitable for this study.
+
+Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples.
+an over-sampling of minority-class examples and under-sampling of majority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
@@
 
-\newpage \section{Experiments}
+\section{Experiments}
 
@@
 \]}
- \caption{Experimental .}\label{fig:design}
+ \caption{Experimental design .}\label{fig:design}
  \end{figure}
@@
 
+Overall,  XTREE's plans were usually succinct--  in all our data sets,
+ XTREE changes 
+ usually changes around a fifth of the features (see \fig{types}.
+
 An  interesting feature of \fig{changed} was that fewer things 
@@
 
- As to the method that lead to best effectiveness in \fig{jur} and \fig{conf1}, we note from \fig{types}
- that XTREE
- has a consistent behavior in both the discrete and continuous data sets.  Specifically,  in all our data sets,
- XTREE changes 
- usually changes around a fifth of the features.
-
 \input{tex/deltas_avg.tex}
@@
 \bi
-\item {\em ca}:   afferent couplings  =  how many other classes use the specific
+\item {\em ca}:   afferent couplings  =  \# classes using this
 				class;
-\item {\em  	ce}:  efferent couplings =  how many other classes is used by the
-				specific class. 
-\item {\em cbm}: coupling between methods =  total number of new/redefined methods
+\item {\em  	ce}:  efferent couplings =  \# classes  used by this
+				  class. 
+\item {\em cbm}: coupling between methods =  \# new/redefined methods
 				to which all the inherited methods are coupled
@@
 
-\item {\em ic}:   inheritance coupling =  number of parent classes to which a given
-				class is coupled (includes counts of methods and variables inherited)
+\item {\em ic}:   inheritance coupling =  \#parent classes  which a given
+				class is coupled (includeing methods and variables inherited)
 \ei
@@
 In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
-At least for the {\em lucense} data set, we can see advice that recommends {\em increasing coupling}
+At least for the {\em lucene} data set, we can see advice that recommends {\em increasing coupling}
 to reduce defects.
@@
 
-For example, the COCOMO model from University of 
-Southern California estimates software development time using a combination of industrial data and some domain expertise
-from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
-as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
-inputs. This generated an development time estimate for each example, which we passed to our four methods.
+% For example, the COCOMO model from University of 
+% Southern California estimates software development time using a combination of industrial data and some domain expertise
+% from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
+% as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
+% inputs. This generated an development time estimate for each example, which we passed to our four methods.
+% 40 times, we let those methods propose changes to those projects. 
+% For assessment purposes, the changes projects were then feed back to the COCOMO
+% oracle.
+
+The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring that management challenge.
+POM3 implements the Boehm and Turner model of agile programming~\cite{boehm2003balancing} where teams select tasks as they appear in the scrum backlog.
+POM3 can study the implications of different ways to adjust task lists in the face of shifting priorities.
+The model outputs estimates task completion rates; programmer idle rates; and total overall cost.
+POM3 represents requirements as a set of trees.
+POM3 models requirements as a dependency tree.
+A single requirement in the tree of a prioritization value and a cost, along with a list of child-requirements and dependencies.
+Before any requirement can be satisfied, its children and dependencies must first be satisfied.
+POM3 simulates changing priorities by making teams aware of random items in the requirements tree at random intervals, thus forcing teams to constantly readjust their "to do" lists.
+For further details on this model see~\cite{boehm2003balancing}, ~\cite{boehm2003using}, and ~\cite{port2008}.
+\fig{coc} shows results from using POM3 as an oracle to assess our planning methods and their ability to reduce the project cost.
+In this experiment, we generated a training and testing set with 1000 randomly generated instances, which we passed to our four methods.
 40 times, we let those methods propose changes to those projects. 
-For assessment purposes, the changes projects were then feed back to the COCOMO
+For assessment purposes, the changes projects were then feed back to the POM3
 oracle.
@@
 Using this approach, it is possible to assess the value of a plan by measuring its
-effectiveness with respect to some ground truth (in this case, the COCOMO model).
-As shown in \fig{coc}, XTREE passes this assessment.
-Those results from the COCOMO model endorsed the conclusions of the rest
-of this paper; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change
-example projects.
+effectiveness with respect to some ground truth (in this case, the POM3 model).
+As shown in \fig{coc}, XTREE reduces the  passes this assessment. These results from the POM3 model further endorsed the conclusions of the rest
+of this paper; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.
 
@@
 It is important to also explore the general practicality of the plans.
-To this end, we are currently undertaking is a \textit{feasibility study} collaboration with industrial partners to assess the general feasibility of our changes.
+To this end, we are currently undertaking a \textit{feasibility study} in collaboration with industrial partners to assess the general feasibility of our changes.
 With this study, we would like to understand the nature of changes suggested by XTREE.
-Specifically: are the changes proposed by XTREE small enough to be easily implemented on software projects?
-Also, are the changes large enough to be feasible?
-In other words, we are asking if users can sufficient fine-grained control to make the changes proposed in \fig{changed}.
+For example, we want to check if programmers are willing to implement the changes proposed by XTREE or are there other forces (e.g.
+social, psychological) that might block a team from accepting the advice of an outside oracle like XTREE.
+Note this a concern not just issues for XTREE, but for any tool proposing refactorings.
 
@@
 Some organizations have the resources to 
-run repeated trials to assess  different treatments.
-For example, in one remarkable recent study, Bente et al. report results
-were the same for specification that was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
+run repeated trials to assess  project changes.
+For example, in one   study, Bente et al. report results
+were the same  specification that was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
 to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
@@
 \item
-When planning how to move examples to better classes, it is   best to use a learner that reflects over 
-differences between the classes.
+When planning how to move examples to better classes, it is   best to  reflects over 
+differences between those classes.
 \ei
@@
 
-As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory,  XTREE should work on
-any domain where data miners can generate decision trees. Given the current state of the art in Big Data,
+As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory,  XTREE   works on
+  domain where data miners can generate decision trees. Given the current state of the art in Big Data,
 can   XTREE  be applied to  very large data sets?
@@
 \section*{Acknowledgements}
-The work has partially funded by a National Science Foundation CISE CCF award \#1506586.
+The work has partially funded by NSF  award \#1506586.
 \bibliographystyle{plain}

### 2015-08-27 tex/results_models.tex
Updates from ShareLaTeX

@@
-\begin{figure}[!b]
+\begin{figure}[!t]
 
@@
  ~\hrule~
-\caption{Results: Methods 1,2,3,4 applied to some ground-truth data.
+\caption{Results: Methods 1,2,3,4 applied to some ground-truth data (in this case, the POM3 model).
 Values collected from  40 repeated runs of each method with different random seeds.
-Results show ratios of (1) sum of development effort (in the changes examples) to (2)
-the of the development effort in the original test data.
+Results show the efficacy of XTREE in reducing the total overall cost in the original test data.
 }\label{fig:coc}

### 2015-08-27 Manuscript.tex
Updates from ShareLaTeX

@@
   Berkeley database (one written in ``C'' and one in Java).
-   \fig{bdbc} shows an example of a feature model defining valid combinations of settings to on the
-   the Siegmund et al. datasets. These feature models were used by Siegmund et al. to ensure all their perturbations are value
+Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref})~\cite{sven12}.
+It's worth noting that in each of the above data set, some configurations are interdependent, this is expressed using a \textit{feature model}.
+   \fig{bdbc} shows an example of such a feature model defining valid combinations of settings to on the Berkley database (``C'' version).
+These feature models were used by Siegmund et al.
+to ensure all their perturbations are valid
    (we will use the same models to cull invalid plans).
-  Given those valid perturbations, the systems were then compiled and 
-  Siegmund et al. recorded how long each perturbation took to run a test suite.
 
@@
  \input{tex/algos}
-\subsubsection{Method3= BIC=   Method2 + Best-in-cluster}\label{sect:BIC}
+\subsubsection{Method3= GD=   Method2 + Gradient}\label{sect:BIC}
 
@@
  best-in-cluster associated with  $C_j$).
-The returned plan is then computed  from the delta between the test case
+The returned plan is then computed as a fraction of the gradient between the test case
 and $B_j$.
@@
                 \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
-                \mathrm{{\bf return}\;} R=\frac{\mathit{after}}{\mathit{before}}
+                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
             \end{array}
@@
 then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions expressed as the ratio $R=\frac{\mathit{after}}{\mathit{before}}$.
-This is a unit-less ratio with the following properties:
-\bi
-\item If $R= 1$, this means  ``no change from baseline''; 
-\item If $R < 1$, this indicates ``some reduction to the baseline'';
-\item If $R > 1$, this indicates ``optimization failure''.
+then (3)~return data on the {\em before, after} predictions expressed as percent improvement $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$ with the following following properties:
+\bi
+\item If $R= 0\%$, this means  ``no change from baseline''; 
+\item If $R < 0\%$, this indicates ``some reduction to the baseline'';
+\item If $R > 0\%$, this indicates ``optimization failure''.
 \ei
@@
 \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
-1 &         XTREE &    0.49  &  0.13 & \quart{7}{25}{17}{115} \\\hline  
-2 &      CD &    0.59  &  0.18 & \quart{15}{34}{36}{115} \\
-2 &          BIC &    0.60  &  0.12 & \quart{24}{24}{38}{115} \\\hline  
-3 &      CD+FS &    0.62  &  0.06 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
+1 &         CD+FS &    49  &  13 & \quart{7}{25}{17}{115} \\\hline  
+2 &      CD &    59  &  18 & \quart{15}{34}{36}{115} \\
+2 &          BIC &    60  &  12 & \quart{24}{24}{38}{115} \\\hline  
+3 &      XTREE &    62  &  6 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
 \end{center}
@@
 In this example table, the rows are  sorted on the median values of each method. Note that all the methods
-have   $R<1$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
-while XTREE achieved the greatest reduction (down to 49\% of the original value).
+have   $R<0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
+while XTREE achieved the greatest reduction (of 60\% from the original value).
 
@@
 \bi
-\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data. 
+\item For Siegmund data, we performed a k-fold cross validation.
+Given the relatively smaller sample sizes in Apache and BDBJ, we chose $k=2$.
+This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably.
+However, in 2-fold cross validation, the division of data tends to affect the outcome significantly.
+To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data, repeating the process 40 times as mentioned above.
 \item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
@@
 \item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);
+\item Method3 uses a fraction $F=0.5$ to compute the plan from the delta;
 \item Method4 (XTREE) assumed that another sibling was useful if  had 

### 2015-08-27 tex/results_models.tex
Updates from ShareLaTeX

@@
 \begin{figure}[!t]
-
-~\hrule~
-
 %\noindent\begin{minipage}{0.5\textwidth} 
 {\small 
-\begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
+\begin{tabular}{llrrc}
 \arrayrulecolor{lightgray}
 \textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
-  1 &        XTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
-\hline  2 &      CD+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
-  2 &          BIC &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
-  2 &      CD &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
-\hline 
+  1 &          HOW &    -3.3  &  2.2 & \quart{26}{2}{27}{32} \\
+  1 &      baseln0 &    -2.9  &  11.7 & \quart{17}{14}{28}{32} \\
+  1 &      baseln1 &    -2.75  &  65.5 & \quart{0}{78}{28}{32} \\
+\hline  2 &        dtree &    51.5  &  14.84 & \quart{82}{17}{93}{32} \\
 \end{tabular}}
-
- ~\hrule~
-\caption{Results: Methods 1,2,3,4 applied to some ground-truth data (in this case, the POM3 model).
+\caption{Methods 1,2,3,4 applied to some ground-truth data (in this case, the POM3 model).
 Values collected from  40 repeated runs of each method with different random seeds.
-Results show the efficacy of XTREE in reducing the total overall cost in the original test data.
+Results show the efficacy of XTREE in reducing the total overall cost in the original test data, when other planners fail to do so.
 }\label{fig:coc}

### 2015-08-28 Manuscript.tex
Updates from ShareLaTeX

@@
   Berkeley database (one written in ``C'' and one in Java).
-Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref})~\cite{sven12}.
+Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}.
 It's worth noting that in each of the above data set, some configurations are interdependent, this is expressed using a \textit{feature model}.

### 2015-08-28 tex/results_models.tex
Updates from ShareLaTeX

@@
 \textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
-  1 &          HOW &    -3.3  &  2.2 & \quart{26}{2}{27}{32} \\
-  1 &      baseln0 &    -2.9  &  11.7 & \quart{17}{14}{28}{32} \\
-  1 &      baseln1 &    -2.75  &  65.5 & \quart{0}{78}{28}{32} \\
-\hline  2 &        dtree &    51.5  &  14.84 & \quart{82}{17}{93}{32} \\
+  1 &          BIC &    -3.3  &  2.2 & \quart{26}{2}{27}{32} \\
+  1 &      CD &    -2.9  &  11.7 & \quart{17}{14}{28}{32} \\
+  1 &      CD+FS &    -2.75  &  65.5 & \quart{0}{78}{28}{32} \\
+\hline  2 &        XTREE &    51.5  &  14.84 & \quart{82}{17}{93}{32} \\
 \end{tabular}}

### 2015-08-28 Manuscript.tex
Updates from ShareLaTeX

@@
 \definecolor{Background}{rgb}{1,1,1}
-
+\bibliographystyle{unsrt}
  \definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
@@
 
-Using data from 11 software projects,  XTREE found   larger reductions that three
+Using data from 11 software projects,  XTREE found   larger reductions than three
 alternate methods.  Those reductions had 
-median size of (55\%,38.5\%) and   largest size of (78\%,94\%) in
+median size of (56\%,28\%) and   largest size of (60\%,77\%) in
 (defects,runtimes), respectively.
@@
 
-Note that we make no assumption that a plan mentions every features
+Note that we make no assumption that a plan mentions every feature
 (so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than
@@
 This paper is about the next step {\em after} prediction. Suppose
-a business user is presented with  a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
+a business user is presented with  a prediction and they do not like what they see; e.g. the runtimes are too long or the number of defects is too high. This user may
 then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.
@@
 all their future projects. They comment ``past experiences were taken into account without 
-much consideration for their context~\cite{passos11}.  
+much consideration for their context''~\cite{passos11}.  
  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that 
-  supposed software engineering    ``gurus'' rarely use lessons
+  supposed software engineering    ``Gurus'' rarely use lessons
   from past projects to improve their future reasoning and that such poor
@@
 problems with that approach. Firstly, given the rapid pace of change in software
-engineering, we may be asking questions for which there is no currently accept
+engineering, we may be asking questions for which there is no currently acceptable
 ``best practice''. For example, only very recently has there been any work
 of predicting software runtimes based on choices within Makefiles. While the work
-of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
-cannot yet represent ``internationally accept best practices'' since that work is so
+of Siegmund et al.~\cite{sven12} is certainly state of the art, it does not yet represent ``internationally accept best practices'' since that work is so
 recent.
@@
 learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
-A repeated result in those studies in that the local models generated the better effort
+A repeated result in those studies is that the local models generated better effort
 and defect predictions (better median results,
@@
 While that may be true in general, for particular classes other changes may be more important
-(later in this paper, we show one set of results were that is indeed the case).
+(later in this paper, we show one set of results where that is indeed the case).
 \ei
@@
 it is useful to have automatic methods to recommend changes. Such
-methods can fill in for human gurus (if such gurus are absent) or 
+methods can fill in for human gurus (if such Gurus are absent) or 
 to offer a second opinion.
@@
 \begin{figure}[!t]
-  \includegraphics[width=\linewidth]{figs/twodee.eps} 
+  \includegraphics[width=\linewidth]{figs/2d.eps} 
  % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
@@
 Note that predictions  about changes that  fall within the space of the training+test data, will be at least
-as reliable as the performance value found in the above data mining experiment.
-Given this, one thing  can be asserted about predictions on changed examples:
-\bi
-\item Predictions for changes that move examples towards/away from the training data can be trusted more/less (respectively).
-\ei  
+as accurate as the predictions of the original test data.
+With this, we assert that the predictions for changes that move examples towards/away from the training data can be trusted more/less (respectively).
+
 Accordingly, we should use {\em trust-increasing} planners that generate changed examples {\em closer} to the
@@
 In that green region of changed examples, our belief in the value of predictions
-will be as much (or more) as our belief in the value of the predictions in the red region (that
+will be as much, if not more than our belief in the value of the predictions in the red region (that
 contains the original test data).
-This pattern of \fig{howxy} (where the changes examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+This pattern of \fig{howxy} (where the examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
 paper. Hence,  we can assert that
 predictors learned from these training examples have some authority in the regions
-contain the changes examples.
+containing the changes examples.
 
@@
 The quality of the prediction depends on the nature of the training data.
-Thus, we strongly recommend that both data set and the predictor be assessed prior to planning.
+Thus, we strongly recommend that both the data set and the predictor be assessed prior to planning.
 This ensures that the predictor's performance is adequate for a data set.
@@
 To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
-  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
-  See github.com/ai-se/XTREE\#data for full access to this data.
-
-   The Jurecko data records number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
-  nearly two dozen metrics such as number of children, lines of code, etc. For details on the Jurecko data, see  \fig{ck} and \fig{jd}. 
+  software system   configuration data from by  Siegmund et al.~\cite{sven12}\footnote{ See \url{http://git.io/vGYxc} for full access to this data.}.
+
+   The Jureczko data records number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
+  nearly two dozen metrics such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. 
  All the planning methods of this paper reflect on the numeric value of the raw defect counts.
-The predictor considers defects to be a discrete class data set,where defects are \textit{\texttt{TRUE}} if the numeric defect count is greater than zero and \textit{\texttt{FALSE}} otherwise.
-
-  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To make that data, Siegmund et al. perturbed
+The predictor considers defects to a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.
+
+  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To obtain the  data, Siegmund et al. perturbed
   the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
@@
 Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}.
-It's worth noting that in each of the above data set, some configurations are interdependent, this is expressed using a \textit{feature model}.
-   \fig{bdbc} shows an example of such a feature model defining valid combinations of settings to on the Berkley database (``C'' version).
+It's worth noting that in each of the above data set, several configurations are interdependent, this is expressed using a \textit{feature model}.
+   \fig{bdbc} shows an example of such a feature model defining valid combinations of settings in the Berkley database (``C'' version).
 These feature models were used by Siegmund et al.
@@
 
-  Our evaluation strategy (discussed below) divides this data into a training a test set.
+  Our evaluation strategy (discussed below) divides this data into a training and a test set.
   From the train set we apply a data miner (to learn a quality predictor) and
@@
   As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
-  For the Siegmind data, this criteria was   relatively easy to achieve.
+  For the Siegmund data, this criteria was   relatively easy to achieve.
   The data in those data sets have a continuous class (runtime of the compiled system)
-  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
+  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
 This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
@@
   and trained a Random Forest
-  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$.
-That is, we can be very confident that the predictors from the Siegmund data can assess
+  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other.
+After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$.
+Therefore, we can be very confident that the predictors from the Siegmund data can assess
 our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
@@
 of classes and testing on smaller test suites).
-Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
+Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
 In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
 
-The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
+In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
 For such data, quality predictor   is be measured using
@@
 We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
-For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select these data sets.
+For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
 With this threshold, however, none of our data sets were be suitable for this study.
@@
 Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples and under-sampling of majority-class examples.
+an over-sampling of minority-class examples and an under-sampling of majority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
@@
 \item Method3 uses a fraction $F=0.5$ to compute the plan from the delta;
-\item Method4 (XTREE) assumed that another sibling was useful if  had 
- a  score less than $\gamma=0.5$ times the mean score of the current leaf.
+\item XTREE considered a sibling useful if it's score was less than $\gamma=0.5$ times the mean score of the current leaf.
  \ei

### 2015-08-28 Manuscript.tex
Updates from ShareLaTeX

@@
 the differences in the branches from some current branch to another desired branch.
-
-Using data from 11 software projects,  XTREE found   larger reductions than three
+Using data from 11 software projects,  XTREE can find  larger reductions than three
 alternate methods.  Those reductions had 
@@
 To enable reproducibility, all scripts and data used in this 
-study are available on-line at github.com/ai-se/XTREE.
+study are available online at \url{http://git.io/vG3DG}.
 
@@
 
-\subsection{From Prediction to Planning}
+\subsection{From Prediction to Planning}\label{sect:p2p}
 
@@
 much consideration for their context''~\cite{passos11}.  
- J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that 
+ J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that the 
   supposed software engineering  ``Gurus'' rarely use lessons
@@
 problems with that approach. Firstly, given the rapid pace of change in software
-engineering, we may be asking questions for which there is no currently acceptable
-``best practice''. For example, only very recently has there been any work
-of predicting software runtimes based on choices within Makefiles. While the work
-of Siegmund et al.~\cite{sven12} is certainly state of the art, it does not yet represent ``internationally accept best practices'' since that work is so
-recent.
+engineering, we may be asking questions for which there is no current widely
+accepted concept of
+``best practice''.
 
@@
 After making those
-changes, we may have a new example that has never seen before. Therefore, it must be asked
+changes, we may have a new example that has never been seen before. Therefore, it must be asked
 {\em ``can we trust the predictions made on such new examples?''}
@@
 
-Accordingly, we should use {\em trust-increasing} planners that generate changed examples {\em closer} to the
+Accordingly, we need {\em trust-increasing} planners to generate changed examples {\em closer} to the
 training examples.  To see how this works, 
  \fig{howxy} is from the {\em ivy} data
-set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
+set (one of the Jureczko data sets used in this paper). It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
 changed  examples displaced after applying a plan (in green).
@@
  \section{Test Data}\label{sect:tesd}
-To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
-  software system   configuration data from by  Siegmund et al.~\cite{sven12}\footnote{ See \url{http://git.io/vGYxc} for full access to this data.}.
-
-   The Jureczko data records number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
+To assess our planning methods, our data (see git.io/vGYxc) comes from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
+  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
+
+   The Jureczko data records the number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
   nearly two dozen metrics such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. 
@@
 
-  Our evaluation strategy (discussed below) divides this data into a training and a test set.
-  From the train set we apply a data miner (to learn a quality predictor) and
+  Our evaluation strategy (discussed below) divides this data into a training and a testing set.
+  From the training set we apply a data miner (to learn a quality predictor) and
   various planning methods (to learn different plans). Next, we try applying
@@
   Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other.
-After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$.
-Therefore, we can be very confident that the predictors from the Siegmund data can assess
+After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$;
+  i.e.  we are confident that the predictors from the Siegmund data can assess
 our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
@@
 As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
+Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
 built from the training data, then applied to the test data.
@@
 
-In summary, while we cannot trust predictors from some of our defect data sets,
+In summary, while we cannot trust predictors from some of our Jureczko data sets,
 we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
@@
 
-This section described XTREE (which we call Method4) and  three  
+This section describes XTREE (which we call Method4) and  three  
 alternate methods for learning plans.
@@
 	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}. 
-
-  Note that all   methods have the  properties proposed in \tion{prelim}.
-  That is, they are {\em local learners}; i.e. different test examples
-  will be given plans that are specialized  to their particulars. Also,
-  they  are {\em trust-increasing}; 
+All  methods have the  properties proposed in \tion{prelim}: (1)~they are {\em local learners}
+that find   plans   particular to each test case; and (2)~they  are {\em trust-increasing}; 
   i.e. they change  examples such that they move {\em closer} to the training data.
@@
 Our  description of the methods adopts the following convention. All variables
-  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
+  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$.
   In this paper, we show our current settings to these variables produces useful
@@
 
-{\em Assumption1: } Method1 assumes that large data sets can be adequately represented by a few dozen (or so) 
+{\em Assumption1: } Large data sets can be adequately represented by a few dozen (or so) 
 centroids.
@@
   After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
-replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
+replaces all clusters with a  centroid $C_i$ computed from the median value of each
 continuous/discrete feature. After that, it
 finds the closest centroid $C_j$ that has a better
-performance score. For defect data, ``better'' means fewer defective examples while for the config data,
+performance score. For Jureczko data, ``better'' means fewer defective examples while for the Seigmund data,
 ``better'' means lower median runtimes for the examples in that cluster.
@@
 
-{\em Summary2:} Method2 works line Method1 but now the   plans  only
+{\em Summary2:} Method2 works like Method1 but now the   plans  only
 mention the $\beta=33\%$ most informative features. Hence, Method2's plans are simpler.
 
-{\em Assumption2: } Method2 assumes that, when reasoning about centroids, we can
-just use   features
-that   best distinguish    centroids; i.e. whose values appear in just a few centroids.
+{\em Assumption2: }  When reasoning about centroids, we can
+just use features that best distinguish centroids; i.e. whose values appear in just a few centroids.
 
@@
  \input{tex/algos}
-\subsubsection{Method3= GD=   Method2 + Gradient}\label{sect:BIC}
-
-{\em Summary3:} Method3 is like Method2, but it uses more knowledge about the training data.
-
-{\em Assumption3: } Method3 assumes that there exists ``gradients'' between and
-within clusters which, if used, will better guide us to finding better plans.
+\subsubsection{Method3= Best In Cluster (BIC)=   Method1 + Gradient}\label{sect:BIC}
+
+{\em Summary3:} Method3 is like Method1, but it uses more knowledge about the training data.
+
+{\em Assumption3: } There exists ``gradients'' between clusters which, if used, will better guide us to finding better plans.
 
 {\em Details3:}
-Method3 summarizes clusters into {\em two} examples: (1)~the centroid $C_x$ found in Method1 and
-(2)~the best-in-cluster $B_x$  example; i.e. the  example in that cluster
-with the best performance score. 
-For the Siegmud data, $B_x$ is the cluster member with the fastest runtime;
-for the Jureczko data, $B_x$  is the cluster member with lowest raw defect count
-(resolving ties at random).
-
-Method3 connects  each centroid to a nearest neighbor
-by a {\em gradient}.
-Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
-For each test instance, Method3 
-find the nearest gradient, the runs up to the top  best end $C_j$, then extracts $B_j$ (which is the
- best-in-cluster associated with  $C_j$).
-The returned plan is then computed as a fraction of the gradient between the test case
-and $B_j$.
+Method3 generates clusters from the training data just like Method1.
+Following this, it summarizes the clusters $C_X$ using ``Best-In-Cluster'' $B_X$, which is (1) the centroid of the cluster for the Jureczko data; (2) the cluster's member with the fastest runtime for the Seigmund data.
+Following this, Method3 connects  each cluster $C_i$ to a nearest neighbor $C_j$ by a {\em gradient}.
+Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the (worst,best) performance scores, respectively.
+
+For each test instance, Method3 finds the nearest gradient, the runs up to the top (best) end $C_j$, then extracts $B_j$ (which is the best-in-cluster associated with  $C_j$).
+The returned plan is then computed as a fraction ($\omega=0.5$) of the distance between the test case and it's corresponding $B_j$.
 
@@
 
- \begin{figure}[!h]
+ \begin{figure}[!ht]
 {\small 
@@
 then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions expressed as percent improvement $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$ with the following following properties:
+then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
 \bi
-\item If $R= 0\%$, this means  ``no change from baseline''; 
-\item If $R < 0\%$, this indicates ``some reduction to the baseline'';
-\item If $R > 0\%$, this indicates ``optimization failure''.
+\item If $R  = 0\%$, this means  ``no change from baseline''; 
+\item If $R \gt 0\%$, this indicates ``some reduction to the baseline'';
+\item If $R \lt 0\%$, this indicates ``optimization failure''.
 \ei
@@
   XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
-  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
-  (with different random number seeds).
-The the value 40 was chosen to be  larger than the 30 samples  required
-to satisfy the central limit theorem.
+  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs,
+  with different random number seeds
+(we use 40 since that is  more than the 30 samples  needed
+to satisfy the central limit theorem).
 
@@
   repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}.
-In that test, using the median values of each method,  
+
+In accordance to that test, using the median values of each method, we
 sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
-Then,
-splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
+Then, we
+split $l$ into sub-lists $m,n$ in order to maximize the expected value of
  differences  in the observed performances
@@
  \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
-Scott-Knott  applies an statistical  hypothesis test $H$ to check
+
+We then apply a apply a statistical hypothesis test $H$ to check
 if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping). 
 If so, Scott-Knott  recurses on the splits.
-In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
 
@@
 
-Our results are presented by the line diagrams like those shown on the right-hand-side of the following example table.
-The black dot shows the median $R$ value and the horizontal likes stretch over the inter-quartile
-range (hereafter, IQR) that is the space from the 25th percentile value to
-the 75th percentile value.
+Our results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
+The black dot shows the median $R$ value and the horizontal likes stretch 
+from the 25th percentile to the 75th percentile (a region called the inter-quartile
+range, or IQR).
 
@@
 
-{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
+{\small Example \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
 \arrayrulecolor{lightgray}
-\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
-1 &         CD+FS &    49  &  13 & \quart{7}{25}{17}{115} \\\hline  
-2 &      CD &    59  &  18 & \quart{15}{34}{36}{115} \\
-2 &          BIC &    60  &  12 & \quart{24}{24}{38}{115} \\\hline  
-3 &      XTREE &    62  &  6 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
+\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
+1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
+\hline 2 &          BIC &    50  &  12 & \quart{34.2}{9.6}{40}{115} \\ 
+\hline 3 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
+3 &         CD+FS &    43  &  13 & \quart{30}{10.4}{35}{115} \\  
+\hline \end{tabular}}
 \end{center}
@@
 In this example table, the rows are  sorted on the median values of each method. Note that all the methods
-have   $R<0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
-while XTREE achieved the greatest reduction (of 60\% from the original value).
-
-The above eample table has a  left-hand-side  {\bf Rank} column, computed using the
-Scott-Knott test described above. This column reports if
-the values for each method are statistically different and are more than trivially different. 
-In this example table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.
+have   $R\gt0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
+while XTREE achieved the greatest reduction (of 62\% from the original value).
+
+The above example table has a  left-hand-side  {\bf Rank} column, computed using the
+Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and CD+FS together are ranked the worst.
 
@@
 Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
-\bi
-\item For Siegmund data, we performed a k-fold cross validation.
-Given the relatively smaller sample sizes in Apache and BDBJ, we chose $k=2$.
+
+For Siegmund data, we performed a k-fold cross validation.
+Given the relatively small sample sizes in Apache and BDBJ, we chose $k=2$.
 This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably.
 However, in 2-fold cross validation, the division of data tends to affect the outcome significantly.
-To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data, repeating the process 40 times as mentioned above.
-\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
- all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd})
-occurred in the {\em train} phase of \fig{design}.
-\ei
+To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the other, repeating the process 40 times as mentioned above.
+
+The Jureczko data, being temporal in nature, allows us
+to implement a validation procedure that ensures only past data is ever used to
+predict future values. Hence, in that data, we used the train/test sets shown in \fig{j}. 
+(Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).
+
 \subsection{Experimental Results}
@@
 some data sets were harder to optimize that others.  SQL (in \fig{conf1}) defied all 
-our methods for reducing runtimes. Also, BDBJ (in \fig{conf1}) was hard to optimize
-for all methods except XTREE.  However, in other data sets,
+our methods for reducing runtimes. While all methods performed equally well for Jedit, XTREE was the only method that could optimize BDBJ (in \fig{conf1}). In general, in most data sets,
 large reductions were observed:
@@
 \item
-Down to 22\% of the original baseline in Ant of \fig{jur};
-\item
-Down to 6\% of the original baseline in BDBD  of \fig{conf1};
+An improvement of 60\%, as compared to the original baseline in Ant and Lucene of \fig{jur};
+\item
+An improvement of 77\% as compared to the original baseline in BDBC  of \fig{conf1};
 \ei
-Overall, XTREE was most effective. It was always  the top-ranked method and (with 
-the exception of SQL), had significant reductions in the median performance
-values: median improvement of least 10\% lower (i.e. better) than the next ranked method.
+
+Overall, XTREE was most the effective. It was always the top-ranked method with 
+the exception of SQL and Jedit.
+Where it ranked better, it had significant improvements in the median performance values (1) In the Jureczko data sets, there was a median improvement of around 10\% larger than the next ranked method; (2) In the Seigmund data sets, improvements as large as 36\% greater than the next ranked method were observed.
 
@@
   are {\em lower}). For example,
- Method1 (CD) was the least succinct  since it  wanted to change all features 
+  XTREE's plans were usually succinct--  in all  data sets,
+ XTREE   changes around a fifth of the features (see \fig{types}).
+On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features 
  (observe the change frequencies as 
@@
 if this approach lead 
-to the most effective changes. However, looking at 
+to the most effective changes. However, in
 \fig{jur} and  \fig{conf1}, there is no evidence for this.
 
-Overall,  XTREE's plans were usually succinct--  in all our data sets,
- XTREE changes 
- usually changes around a fifth of the features (see \fig{types}.
-
 An  interesting feature of \fig{changed} was that fewer things 
-were changed in the config data sets {\em BDBJ} than in the defect data set
+were changed in the Seigmund data sets like {\em BDBJ} than in the Jureczko data set like
 {\em lucene}. 
@@
 \fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
-there are fewer changed features in the config data than in the defect prediction data. 
+there are fewer changed features in the Seigmund data than in the defect prediction data. 
 One explanation for that is the nature of the features: the
-defect data sets has continuous
+Jureczko data sets have continuous
 features while the 
-config data has binary independent features  (where so setting was turned ``on'' or ``off''). When exploring these different data types, it is possible to find more
-\bi
-\item
-``Gentle slopes'' the lead to small changes
-in  continuous space;
-\item
- ``Sharp cliffs'' that lead to major change in the discrete
-space. 
-\ei
-Hence, our planners can make fewer larger changes in the config data.
- For example, Method3 (BIC) and Method1 (CD) make far fewer changes in   data sets with discrete features (the Siegmund data) than in data with continuous features (the Jureczko data).
+Seigmund data has binary independent features (where settings can be turned ``ON'' or ``OFF'').
+While the features of the Jureczko data sets are not constrained, those of Seigmund data sets are subject to several constrains (dictated by their feature models, see \fig{bdbc}).
+It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores.
 
@@
 
-\item {\em ic}:   inheritance coupling =  \#parent classes  which a given
+\item {\em ic}:   inheritance coupling =  \# parent classes  which a given
 				class is coupled (includeing methods and variables inherited)
 \ei
-For the  {\em lucene} XTREE results  of  \fig{changed},
+In many of our results with the Jureczko data, it was indeed true that the changes
+proposed by XTREE lead to lower coupling. However, the {\em lucene} results were quite
+different and quite surprising.
+
+In the {\em lucene} XTREE results  of  \fig{changed},
 the most frequent change was to alter the lines of code in a class (see the tallest
@@
 Looking at the logs of our planner, we can see
-that XTREE's proposed changed is to {\em reduce} the size of a class.
+that XTREE's proposed change is to {\em reduce} the size of a class.
 The only way to do that,  while keeping the  same functionality, it is create a network
@@
 \item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);
-\item Method3 uses a fraction $F=0.5$ to compute the plan from the delta;
+\item Method3 uses  $\omega=0.5$ to find its deltas;
 \item XTREE considered a sibling useful if it's score was less than $\gamma=0.5$ times the mean score of the current leaf.
@@
 of predictors learned on the training data to assess the impact of our planners.
-This issues was discussed in detail in \tion{trust}.
+This issue was discussed in detail in \tion{trust}.
 
@@
 
-The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring that management challenge.
+The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring management challenges.
 POM3 implements the Boehm and Turner model of agile programming~\cite{boehm2003balancing} where teams select tasks as they appear in the scrum backlog.
@@
 The model outputs estimates task completion rates; programmer idle rates; and total overall cost.
-POM3 represents requirements as a set of trees.
 POM3 models requirements as a dependency tree.
-A single requirement in the tree of a prioritization value and a cost, along with a list of child-requirements and dependencies.
+A single requirement in the tree of a prioritization value has a cost, along with a list of child-requirements and dependencies.
 Before any requirement can be satisfied, its children and dependencies must first be satisfied.
@@
 For further details on this model see~\cite{boehm2003balancing}, ~\cite{boehm2003using}, and ~\cite{port2008}.
+
 \fig{coc} shows results from using POM3 as an oracle to assess our planning methods and their ability to reduce the project cost.
@@
 40 times, we let those methods propose changes to those projects. 
-For assessment purposes, the changes projects were then feed back to the POM3
+For assessment purposes, the changed projects were then feed back to the POM3
 oracle.
@@
 effectiveness with respect to some ground truth (in this case, the POM3 model).
-As shown in \fig{coc}, XTREE reduces the  passes this assessment. These results from the POM3 model further endorsed the conclusions of the rest
-of this paper; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.
-
-\subsection{  Feasibility}\label{sect:feasibility}
-It is important to also explore the general practicality of the plans.
-To this end, we are currently undertaking a \textit{feasibility study} in collaboration with industrial partners to assess the general feasibility of our changes.
-With this study, we would like to understand the nature of changes suggested by XTREE.
-For example, we want to check if programmers are willing to implement the changes proposed by XTREE or are there other forces (e.g.
-social, psychological) that might block a team from accepting the advice of an outside oracle like XTREE.
-Note this a concern not just issues for XTREE, but for any tool proposing refactorings.
+As shown in \fig{coc}, XTREE reduces the cost by 59\% thereby passing this assessment.
+These results from the POM3 model further endorse our previous conclusions; i.e.
+compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.
 
@@
 classical AI. 
-That kind of planning is a logical procedure~\cite{Fikes1971}
+Those kinds of planners employ a logical procedure~\cite{Fikes1971}
 that seeks an ordering on {\em operators} to take some domain
@@
 This classical logical approach is known to suffer from
-computational bottlenecks~\cite{Bylander1994} while tools like XTREE will scale to any domains
+computational bottlenecks~\cite{Bylander1994} while tools like XTREE will scale to any domain
 that can generate decision trees.
@@
 run repeated trials to assess  project changes.
-For example, in one   study, Bente et al. report results
-were the same  specification that was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
+For example, in one   study, Bente et al. reported results
+where the same  specification was developed by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
 to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
@@
 this paper has used data miners to build an oracle that can assess changed examples. The advantage
-of this approach is that it required far less resources to assess the effectiveness of proposed
+of our approach is that it required far less resources to assess the effectiveness of proposed
 changes to a project.
@@
 of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
-to mutate new examples away from the kinds of data used to commission the model (in which case, we would
+to prevent new examples from mutating away from the kinds of data used to commission the model (in which case, we would
 start doubting the model's output).
@@
 
-\section{Summary}
-
-The planner proposed in this paper propose changes to software project details in order to improve the expected
+\section{Conclusions}
+
+The planner proposed in this paper proposes changes to software project details in order to improve the expected
 value of the performance scores of that part of the project.
@@
 data miners can be used to build oracles to assess planners.
-Such planners should be {\em trust-increasing}; i.e. they propose changes that generate
+Such planners should be {\em trust-increasing}; i.e. they should propose changes that generate
 changed examples that are closer to the training data of the data miner.
@@
 \item
-When planning how to move examples to better classes, it is   best to  reflects over 
+When planning how to move examples to better classes, it is best to reflect over 
 differences between those classes.
 \ei
-
-\section{Questions for Future Work}
-
+\newpage
+\section{Future Work}
+Future work in this area could explore numerous questions. For example, 
 XTREE, as used here, sought improvements in a one goal (the class variable). Does XTREE work for multi-goal reasoning?
 
- The XTREE algorithm seems quite general to any data of table with rows containing
+Also, the XTREE algorithm seems quite general to any data of table with rows containing
 weighted classes (so we can distinguish ``bad'' rows from ``better'' ones). 
@@
 As an example of a domain that might benefit from XTREE, recent results raise doubts about
-the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``really bad smell''
-detector to select the subset of possible refactorings that have the most   potential benefit?
-
-As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory,  XTREE   works on
-  domain where data miners can generate decision trees. Given the current state of the art in Big Data,
-can   XTREE  be applied to  very large data sets?
+the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``bad smell'' detector to select the subset of possible refactorings that have the most potential benefit?
+
+As to scalability, XTREE is a post-processor to a decision tree algorithm.
+Hence, in theory, XTREE   works on domain where data miners can generate decision trees.
+Given the current state of the art in Big Data, can XTREE  be applied to  very large data sets?
+
+We discussed above in \tion{p2p} the general  conclusions of Passos, J{\o}rgensen  et.al~\cite{passos11,jorgensen09};
+i.e. software developers are reluctant to surrender their old biases
+in the face of new data. 
+Accordingly, it must be asked if  the {\em mental resistance} of developers
+will prevent them applying XTREE's automatically generated recommendations of tools?
+Note this this issue is not just a concern for XTREE, but also for any 
+automatic tool proposing refactorings.
 
 There are many more methods for generating plans and
-no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta,\gamma$
+no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta$, and $\gamma$
 parameters that controlled XTREE. Would we get better results if we varied those parameters?
@@
 was to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
-Will other  researchers apply this strategy  to repeat and/or improve
+The authors of this paper would actively support other teams exploring this
+method (with or without using our current code base). So our last questions is this:
+will other  researchers try to repeat and/or improve
 (or even refute) our results?  Perhaps.
@@
 There are a wide variety of clustering methods to choose from.
-A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and defect data from the PROMISE repository~\cite{promise}.
+A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and Jureczko data from the PROMISE repository~\cite{promise}.
 In that study methods such as WHERE, K-Means, mini-batch-K-Means, DBScan, EM, and Ward were investigated.

### 2015-08-28 tex/algos.tex
Updates from ShareLaTeX

@@
 the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
-Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
+Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{(x_i - \bar{x})^2}{n-1}\right)}$}
 where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
@@
 
-				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
+				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$. 
 		Using this measure, WHERE runs as follows:
@@
 	Find a split in the values of  independent features that most reduces the variability
-  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
-   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.
+  of the dependent feature (measured using \fig{where}.A). Contruct a tree using these splits.
 

### 2015-08-28 tex/prediction_stats.tex
Updates from ShareLaTeX

@@
 		On the right-hand-side, we show the {\em results from learning}.
-		Data is ``good'' if it has recall over 60\% and false alarm under 40\%
-		(and note that, after tuning, there are more ``good'' than before).
+		Data is usable if it has a recall of 60\% or more and false alarm of 30\% or less (and note that, after tuning, there are more usable data sets than before).
 Results  	\colorbox{celadon}{ marked with ``$\star$''} show large improvements in performance, after tuning
 		(lower {\em pf} or higher {\em pd}).
-		The data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
+		Data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are  performing
+		poorly-- that data so few non-defective examples  that it  is hard for
+		our learners to distinguish between classes.
 	}\label{fig:j}

### 2015-08-28 tex/results_models.tex
Updates from ShareLaTeX

@@
 \arrayrulecolor{lightgray}
-\textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
-  1 &          BIC &    -3.3  &  2.2 & \quart{26}{2}{27}{32} \\
-  1 &      CD &    -2.9  &  11.7 & \quart{17}{14}{28}{32} \\
-  1 &      CD+FS &    -2.75  &  65.5 & \quart{0}{78}{28}{32} \\
-\hline  2 &        XTREE &    51.5  &  14.84 & \quart{82}{17}{93}{32} \\
-\end{tabular}}
+\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
+  1 &        XTREE &   59   &  9 & \quart{75}{4}{78}{99} \\
+\hline   2 &      BIC &    5  & 1  & \quart{44}{0}{44}{99} \\
+2 &          CD+FS & -7  & 100  & \quart{0}{53}{43}{99} \\   
+2 &      CD &    -9  &  77 & \quart{21}{41}{42}{99} \\
+\hline \end{tabular}}
 \caption{Methods 1,2,3,4 applied to some ground-truth data (in this case, the POM3 model).
 Values collected from  40 repeated runs of each method with different random seeds.
-Results show the efficacy of XTREE in reducing the total overall cost in the original test data, when other planners fail to do so.
-}\label{fig:coc}
+Results show the efficacy of XTREE in reducing the total overall cost in the original test data, when other planners fail to do so.}\label{fig:coc}
 \end{figure}

### 2015-08-28 Manuscript.tex
Updates from ShareLaTeX

@@
 For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
-These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+These researchers warn that even if a hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
 Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
@@
 Measured in terms of effectiveness,
-some data sets were harder to optimize that others.  SQL (in \fig{conf1}) defied all 
+some data sets were harder to optimize than others.  SQL (in \fig{conf1}) defied all 
 our methods for reducing runtimes. While all methods performed equally well for Jedit, XTREE was the only method that could optimize BDBJ (in \fig{conf1}). In general, in most data sets,
@@
   are {\em lower}). For example,
-  XTREE's plans were usually succinct--  in all  data sets,
+  XTREE's plans were usually succinct --  in all  data sets,
  XTREE   changes around a fifth of the features (see \fig{types}).
@@
 If a planner only ever reports conclusions that were already known, then that planner offers
-little value-added over ``just use established wisdom''. Accordingly, we studied our results
+little value over ``just use established wisdom''. Accordingly, we studied our results
 for plans that were somewhat counter-intuitive.
@@
 that XTREE's proposed change is to {\em reduce} the size of a class.
-The only way to do that,  while keeping the  same functionality, it is create a network
+The only way to do that,  while keeping the  same functionality, is to create a network
 of smaller classes that interact to produce that functionality.
@@
 POM3 can study the implications of different ways to adjust task lists in the face of shifting priorities.
-The model outputs estimates task completion rates; programmer idle rates; and total overall cost.
+The model outputs estimated task completion rates; programmer idle rates; and total overall cost.
 POM3 models requirements as a dependency tree.
@@
 40 times, we let those methods propose changes to those projects. 
-For assessment purposes, the changed projects were then feed back to the POM3
+For assessment purposes, the changed projects were then fed back to the POM3
 oracle.
@@
 
-The XTREE planner is somewhat different to the logical-based planners explored by 
+The XTREE planner is somewhat different to the logic-based planners explored by 
 classical AI. 

### 2015-08-29 Manuscript.tex
Updates from ShareLaTeX

@@
 \begin{abstract}
- Business users now demand more insightful
- analytics; specifically, tools that generate ``plans''--
-  specific suggestions on  what to change  in order to
- improve  the predicted values.
-
- This paper proposes  the  XTREE planner for software projects. 
-XTREE inputs tables of data of independent features plus a   weighted class
-  indicating how much each row is  ``bad'' or  ``better''. 
-  Plans are edits to  
-rows such that the changed row is more likely to be ``better''.  XTREE
-learns those plans by building a decision tree across the data, then reporting
-the differences in the branches from some current branch to another desired branch.
-Using data from 11 software projects,  XTREE can find  larger reductions than three
-alternate methods.  Those reductions had 
-median size of (56\%,28\%) and   largest size of (60\%,77\%) in
-(defects,runtimes), respectively.
+
+Business users now demand more insightful analytics; specifically, tools that generate ``plans''-- specific suggestions on  what to change  in order to improve  the predicted values.
+
+This paper proposes XTREE, a planner for software projects.
+XTREE receives tables of data with independent features and a corresponding weighted class which indicates the quality (``bad'' or ``better'') of each row in the table.
+Plans are edits to  the rows which ensures the changed row is more likely to be of a ``better'' quality.
+XTREE learns those plans by building a decision tree across the data, then reporting the differences in the branches from some current branch to another desired branch.
+Using data from 11 software projects,  XTREE can find better plans compared to three
+alternate methods. Those plans have lead to improvements with a median size of (56\%, 28\%) and   largest size of (60\%, 77\%) in (defect counts, runtimes), respectively.
 \end{abstract}
+
+\keywords{Defect prediction; Configuration; Planning; Decision trees.}
 \section{Introduction}
@@
 it is useful to have automatic methods to recommend changes. Such
-methods can fill in for human gurus (if such Gurus are absent) or 
+methods can fill in for human expertise (if such experts are absent) or 
 to offer a second opinion.
@@
 
-Accordingly, we need {\em trust-increasing} planners to generate changed examples {\em closer} to the
+Accordingly, we need {\em trust-increasing} planners to generate new examples {\em closer} to the
 training examples.  To see how this works, 
@@
 In that green region of changed examples, our belief in the value of predictions
-will be as much, if not more than our belief in the value of the predictions in the red region (that
+will be just as much as, if not more than, our belief in the value of the predictions in the red region (that
 contains the original test data).
-This pattern of \fig{howxy} (where the examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+This pattern of \fig{howxy} (where the new examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
 paper. Hence,  we can assert that
@@
  All the planning methods of this paper reflect on the numeric value of the raw defect counts.
-The predictor considers defects to a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.
+The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.
 
@@
 Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}.
-It's worth noting that in each of the above data set, several configurations are interdependent, this is expressed using a \textit{feature model}.
+It's worth noting that in each of the above data sets, several features are interdependent, this is expressed using a \textit{feature model}.
    \fig{bdbc} shows an example of such a feature model defining valid combinations of settings in the Berkley database (``C'' version).
@@
   Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other.
-After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$;
+After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.8, 99.1, 96.1, \}\%$;
   i.e.  we are confident that the predictors from the Siegmund data can assess
@@
 In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
-For such data, quality predictor   is be measured using
+For such data, quality of the predictor can be measured using
 (1) the  probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
@@
 As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
+Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was
 built from the training data, then applied to the test data.
@@
 alternate methods for learning plans.
-  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
+  XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
   The other methods use the   top-down
@@
 
- {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and
-where you want to be  (which we will call $C_j$).
+ {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be  (which we will call $C_j$).
 
@@
 
-{\em Assumption3: } There exists ``gradients'' between clusters which, if used, will better guide us to finding better plans.
+{\em Assumption3: } There exists ``gradients'' between clusters which, if used, will better guide us to finding beneficial plans.
 
@@
 the clustering algorithm (WHERE) that  
-executes without knowledge of the target class.  {\em Supervised} methods, on the other hand, assume that it is useful to also reflect on the target class.
+executes without knowledge of the target class.  {\em Supervised} methods assume that it is useful to also reflect on the target class.
 
@@
 \subsubsection{A Strategy for Evaluating Planners}
-
- Here is our experimental design:
 
@@
 
-As shown in \fig{design}, we divide the
+Our experimental design is shown in \fig{design}. We divide the
 project data  into two disjoint sets {\em train} and {\em test}
@@
 \item If $R  = 0\%$, this means  ``no change from baseline''; 
-\item If $R \gt 0\%$, this indicates ``some reduction to the baseline'';
+\item If $R \gt 0\%$, this indicates ``
+
+improvement over the baseline'';
 \item If $R \lt 0\%$, this indicates ``optimization failure''.
@@
 
-{\small Example \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
+{\small Example \begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
 \arrayrulecolor{lightgray}
-\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
+\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
 1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
@@
 In this example table, the rows are  sorted on the median values of each method. Note that all the methods
-have   $R\gt0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
+have   $R\gt0\%$; i.e. all these methods reduced the expected value of the performance score in that experiment
 while XTREE achieved the greatest reduction (of 62\% from the original value).
@@
 some data sets were harder to optimize than others.  SQL (in \fig{conf1}) defied all 
-our methods for reducing runtimes. While all methods performed equally well for Jedit, XTREE was the only method that could optimize BDBJ (in \fig{conf1}). In general, in most data sets,
+our methods for reducing runtimes. XTREE was the only method that could optimize BDBJ (in \fig{conf1}). In general, in most data sets,
 large reductions were observed:
@@
 Overall, XTREE was most the effective. It was always the top-ranked method with 
-the exception of SQL and Jedit.
+the exception of SQL.
 Where it ranked better, it had significant improvements in the median performance values (1) In the Jureczko data sets, there was a median improvement of around 10\% larger than the next ranked method; (2) In the Seigmund data sets, improvements as large as 36\% greater than the next ranked method were observed.
@@
 
-  In these plots, the {\em more} succinct a planning method, the {\em less} percent
-  of the runs
-  where 
-  it   recommends changing a particular feature (i.e. the vertical bars in that plot
+  In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e. the vertical bars in that plot
   are {\em lower}). For example,
@@
 While the features of the Jureczko data sets are not constrained, those of Seigmund data sets are subject to several constrains (dictated by their feature models, see \fig{bdbc}).
-It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores.
+It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores, thereby making insignificant changes.
+As evidenced by the results in \fig{}
 
@@
 \ei
-\newpage
 \section{Future Work}

### 2015-08-29 tex/algos.tex
Updates from ShareLaTeX

@@
 			Otherwise, replace any  missing items with values that maximizes the following.
-			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
+			For numerics, $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
 			returns 
@@
 	Find a split in the values of  independent features that most reduces the variability
-  of the dependent feature (measured using \fig{where}.A). Contruct a tree using these splits.
+  of the dependent feature (measured using \fig{where}.A). Construct a standard decision tree using these splits.
 
@@
 let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
-Cull the the features as per Papakroni~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
+Cull the the features as per~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
 where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).

### 2015-08-29 Manuscript.tex
Updates from ShareLaTeX

@@
 It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores, thereby making insignificant changes.
-As evidenced by the results in \fig{}
+As a result all methods, except XTREE, fail to optimize the runtime, as evidenced by the results in \fig{conf1}.
 
@@
 
-If a planner only ever reports conclusions that were already known, then that planner offers
+If a planner only ever reported conclusions that were already known, then that planner offers
 little value over ``just use established wisdom''. Accordingly, we studied our results
@@
 
-Such a surprising plan can be  seen in the {\em lucene} results. Recall the standard advice for
+Such a surprising plan can be  seen in {\em lucene}. Recall the standard advice for
 OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
@@
 proposed by XTREE lead to lower coupling. However, the {\em lucene} results were quite
-different and quite surprising.
+different and rather surprising.
 
@@
 the most frequent change was to alter the lines of code in a class (see the tallest
-red historgram in that figure on the {\em loc}, or lines of code). 
+red histogram in that figure on the {\em loc}, or lines of code). 
 Looking at the logs of our planner, we can see
@@
 \bi
-\item All our planners divide data into groups of size $\alpha=\sqrt{N}$
+\item All our planners divide data into groups of size $\alpha=\sqrt{N}$;
 \item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);

### 2015-08-29 Manuscript.tex
Updates from ShareLaTeX

@@
 
-% For example, the COCOMO model from University of 
-% Southern California estimates software development time using a combination of industrial data and some domain expertise
-% from its author (Barry Boehm). \fig{coc} shows results from using COCOMO 
-% as an oracle to assess our planning methods. In this experiment, random projects were used as COCOMO
-% inputs. This generated an development time estimate for each example, which we passed to our four methods.
-% 40 times, we let those methods propose changes to those projects. 
-% For assessment purposes, the changes projects were then feed back to the COCOMO
-% oracle.
-
 The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring management challenges.
@@
 POM3 simulates changing priorities by making teams aware of random items in the requirements tree at random intervals, thus forcing teams to constantly readjust their "to do" lists.
-For further details on this model see~\cite{boehm2003balancing}, ~\cite{boehm2003using}, and ~\cite{port2008}.
+For further details on this model see~\cite{boehm2003using},~\cite{port2008}, and~\cite{boehm2003balancing}.
 
@@
 This classical logical approach is known to suffer from
-computational bottlenecks~\cite{Bylander1994} while tools like XTREE will scale to any domain
+computational bottlenecks~\cite{Bylander1994}. On the other hand, tools like XTREE will scale to any domain
 that can generate decision trees.
@@
 
-Very few industrial or research groups have access
+Seldom do industrial or research groups have access
 to the kinds of resources needed for this kind study  (evidence: in the six years since the
@@
 proposed changes for all projects are always evaluated by something like the Bente et al. study.
-Hence, 
-this paper has used data miners to build an oracle that can assess changed examples. The advantage
+Hence, this paper has used data miners to build an oracle that can assess changed examples. The advantage
 of our approach is that it required far less resources to assess the effectiveness of proposed
